### 2023-05-02 Data & Scripts/MSC R Script.Rmd
First Push
Massive overhaul and reorganization

---
title: "R Script V4"
author: "Alex Popescu"
date: "2023-01-23"
output: github_document
editor_options: 
  chunk_output_type: console
---

V4 - Made some revisions and added plots. All colors are colorblind friendly.

# MSC Thesis Script

> I used the advanced filter tool in BORIS to extract all instances where a behavior occurred in the presence and absence of a sentinel. This generated 6 tables.
> I then manually inputted the ID, generalized environment, number of **recorded** crows and whether or not the site was baited in the recording. I additionally removed observations in agricultural settings, and observations past September 29th.

### Reasoning

> The number of crows in the area is not a good estimate of group size. The number of foraging members during the foraging event should be better.
> Whether or not the site was baited will likely affect the duration of movement, since crows will need to look for food in non-baited recordings, while the crows have easy access to a large concentration of food in baited recordings. Therefore, the baited crows are expected to move less than non-baited crows.
> #KMG: I can imagine a reviewer would want to know hwo the baiting was standardized - something to keep in mind when you are writing. I removed the observations in agricultural settings because they had no sentinels present. This could be talked about in the discussion: agricultural fields may lack perches for individuals to act as sentinels. Vineyards may have some structures, but the visibility is limited when the vines are laden with fruit.\
> #KMG: This is something that can go in the methods. You had three sites, but one was removed due to lack of sentinals, and then you can have a bit in the discussion. I removed observations past Sept 29 because I was told behaviors can change wildly in the winter. The number of crows spotted decreased past that date, suggesting the migrants have left or were leaving. To not have to include time of year (I already have part of the breeding season), I decided to simplify.\

### Data Management

> After combining the 6 tables into a long one, labeling the behaviors and the presence of a sentinel on the way, I made sure that the headers were factors.\

```{r Packages, include=F}
library("tidyverse")
library("psych")
library("car")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

```{r Open Crow and manage, include =F}
setwd(getwd())
Crow<-read.csv("Crow.csv", stringsAsFactors = T)
#Some management #
Crow$ID<-as.factor(Crow$ID)
Crow$Number.of.Crows<-as.numeric(Crow$Number.of.Crows)
Crow$LDur<-log(Crow$Duration)
```

```{r Crow Double checking, echo = F}
str(Crow)
```

> Good news! Observation ID, bait, presence of a sentinel and behavior are indeed recognized as factors! Number of crows and bout duration as a numeric variable.

# Bout Duration

## Figures

### Histogram

> I first decided to look at the distribution of my durations. I set up a histogram with the distibution of durations across all behaviors. I had to log-transform the duration to get a better distribution.\

```{r Duration Histograms Individual, include = F}
Untransformed.Dur<-hist(Crow$Duration, breaks = 25)
Transformed.Dur<-hist(Crow$LDur, breaks = 25)
```

```{r Duration Histograms Combined, echo=F}
#KMG: good practice when changing par() things. I've added the code in. It will reset the parameters within the session
# par(mfrow = c(1,2))
op <- par(mfrow = c(1,2))
plot(Untransformed.Dur, main="Untransformed", ylab="Frequency", xlab="Duration (s)")
plot(Transformed.Dur, main="Transformed", ylab="", xlab="Log Duration (s)")
par(op)
```

```{r Crow Histogram, echo=F}
Hist<-ggplot(data = Crow) +
  geom_histogram(aes(fill = Behavior, x = LDur)
                 , alpha = 1
                 , bins = 50)+
  scale_fill_manual(values=cbPalette
                    , labels = c("Foraging", "Alert", "Moving"))+
  xlab("Log Duration (s)")+
  ylab("Count")
Hist
```

> I should use a robust linear mixed model. LMMs by themselves are robust against violations of their assumptions (I found an article about this), but I will nonetheless use 'robustlmm' to make my model, since their computations for weights are apparently better.

### Boxplot

```{r Crow boxplot, echo = F}
Box<-ggplot(data = Crow)+
  geom_boxplot_pattern(aes(x = G.Env, y = LDur, fill = Behavior, pattern = Sentinel)
                       , pattern_angle = 45
                       , pattern_size = 0.05
                       , pattern_spacing = 0.02
                       , color = "black")+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = cbPalette
                    , labels = c("Foraging", "Alert", "Moving"))+
  scale_pattern_manual(values = c("crosshatch", "none")
                       , labels = c("Sentinel Absent", "Sentinel Present"))+
  guides(pattern = guide_legend(title = "Presence of a sentinel"
                                , override.aes = list(fill = c("white")))
         , fill = guide_legend(override.aes = list(pattern = "none")))+
  xlab("Generalized Environment")+
  ylab("Log Duration(s)") +
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())
Box
#KMG suggestion: I'd make the hatch marks the one where sentinal is absent. They remind me of X's which makes me think there is no sentinal present. Just me, though, so not a big deal. Also a reminder to check if these are colour blind firendly-I don't think the blue and green are.

#AP: I swapped the crosshatch in the figure and used a colorblind-friendly palette.
```

### Violin plot

```{r Crow Violin plot, echo = F}
Violin<-ggplot(data = Crow
               , aes(x = Behavior, y = LDur, fill = Sentinel))+
  geom_violin()+
  theme_bw()+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = cbPalette
                    , labels = c("Sentinel Absent", "Sentinel Present"))+
  scale_x_discrete(labels=c("Foraging", "Alert", "Moving"))+
  xlab("Generalized Environment")+
  ylab("Log Duration(s)")+
  facet_grid(~G.Env)+
  stat_summary(fun = "mean"
               , geom = "crossbar"
               , position = "dodge"
               , color = "red")
Violin
```

> #KMG: Is there a way to group the violin plot like the boxplot? So the violins for foraging sentinal present/absent are next to each other? Will be easier to see your conclusions. Also, have you looked at plotting means and SE? Here's a plot #Kiyoko plot of means and SE. KMG: I've put the function to generate means, SE, SD, and CI in the github. With this plot, it's easier to see the differences in the commercial area where the sentinal presence increases the behaviours. It's much messier in the green areas. I wonder if it's a refuge or visibility thing?

```{r Base R Plot, include = F}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
MeanDur <- summarySE(data = Crow, measurevar = "Duration"
                     , groupvars = c("G.Env", "Behavior", "Sentinel"))
#switching the order of groupvars will change how they are grouped.
plot(x = c(1:12)
     , y = MeanDur$Duration
     , xlim = c(0,12)
     , ylim = c(1,3)
     , type = "n"
     , ylab = "Duration (seconds)", xlab = "Generalized Environment")
arrows(c(1:12), MeanDur$Duration - MeanDur$se
       , c(1:12), MeanDur$Duration + MeanDur$se
       , length = 0.1, angle = 90, code = 3, lwd = 2)
points(x = c(1:12), y = MeanDur$Duration
       , pch = c(rep(c(16, 17), 3)), cex = 2
       , col = c("orange", "orange", "blue", "blue", "green", "green"))
legend(x = "bottomright", c("Sentinel Absent", "Sentinal Present", "Foraging", "Alert", "Moving")
       , bty = "o", cex = 1.05, bg = "white"
       , pch = c(16, 17, 15, 15, 15), col = c("black", "black", "orange", "blue", "green"))
#KMG: I'm not going to fix the X-axis labels and those kinds of things right now. I can do that if you like this plot or things it's informative. I think it's easier to see the differences. left side is commercial, right side is green. Legend should be correct.
```

```{r Dot Plot V2, echo = F}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
MeanDur <- summarySE(data = Crow, measurevar = "Duration"
                     , groupvars = c("G.Env", "Behavior", "Sentinel"))
DurDot<-ggplot(data = MeanDur
               , aes(x = G.Env
                     , y = Duration
                     , color = Behavior
                     , shape = Sentinel))+
  geom_point(position = position_dodge2(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(Duration-se)
                    , ymax=(Duration+se))
                , width = 0.2
                , position = position_dodge(width=0.9))+
  theme_bw() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , labels = c("Foraging", "Alert", "Moving")) +
  scale_shape_manual(values = c(16,17)
                     , labels = c("Sentinel Absent"
                                  , "Sentinel Present")) +
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())

DurDot
```

> Very cool! I modified the violin and boxplot, but I think the last plot is the most informative. We can see that the presence of a sentinel increases the duration of bouts in commercial areas. In green areas, however, this effect is reduced apart from in alert behavior.

## The Tests

### Assumption testing first

> I used robustlmm::rlmer to create the models. It takes longer to run than the regular lmer test because of the computations it makes to weigh the model. I then plotted the models to test the assumptions of the model. The model I used was "log(Duration) \~ Behavior + (Sentinel\* G.Env) + Number of crows + Baited + (1\|ID)". KMG: what is the backslash?
>
> Behavior is expected to be affected by both the presence of a sentinel and the environment the crows are in. I am therefore curious to see the interactions between these factors. The presence of bait and number of crows are not linked with the presence of a sentinel or the environment. Arguably, larger groups are more likely to have sentinels than not, but I don't believe the interaction to be important here.

### RLMM

```{r Duration RLMM, echo=F}
RLMM<-rlmer(LDur~Behavior+Sentinel*G.Env+Number.of.Crows + Baited + (1|ID), data = Crow)
plot(RLMM)
```

> That took AGES to run. There's a big "cliff" on the normal Q-Q vs residuals plot. This means that there are more extreme values than would be expected if they came from a truly normal population.

### Bout Results

```{r Duration RLMM Summary, echo = F}
summary(RLMM)
```

> A whole lot of numbers appear and no p-values. I want it to look better.

```{r Duration Better Summary, echo = F}
sjPlot::tab_model(RLMM
                  , show.re.var = T
                  , title = "Effects of generalized environment, bait, number of crows and the presence of a sentinel on durations of behaviors"
                  , dv.labels = " Effects of on duration of behaviors")
#KMG: is baiting not in here as well?
```

> Like this! This is much better! Lets try a simpler output. Curious... Removing the three-way interaction caused many factors to become significant. Number of crows remains non-significant. #KMG: Can we visualize the the baited effects on duration as well?

```{r Simpler output, echo = F}
Simple.Dur<-lmer(LDur~Behavior+Sentinel*G.Env+Number.of.Crows + Baited + (1|ID), data = Crow)
Anova(Simple.Dur)
ranova(Simple.Dur)
```

> Simple model shows significant effects of Behavior, and the interaction between generalized environment and Behavior, and generalized environment and the presence of a sentinel.

### Bout Summary

> I generated a plot to better see what the interaction between Generalized Environment and the presence of a sentinel. First, let's make a summary of the results.

```{r Duration Summary, echo=F}
MeanDur
```

> There are too many combinations to effectively see any trends, let's look at the significant interactions from the models.

### Bait

```{r Bait effect on bout, echo = F}
BaitPlot<- Crow %>%
  summarySE(measurevar = "Duration"
            , groupvars = "Baited") %>%
  ggplot(aes(x = Baited, y = Duration))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_text(aes(label = signif(Duration))
            , size = 3
            , position=position_dodge2(width=0.9, preserve='single')
            , hjust = -0.5
            , vjust = 0.1)+
  geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
                , width = 0.2
                , position = position_dodge(width=.9))+
  theme_bw()+
  theme(text=element_text(size = 14))+
  xlab("")+
  ylab("Mean Bout Duration (s)")+
  scale_x_discrete(labels=c("Not Baited", "Baited"))+
  theme(legend.position = 'bottom'
        , axis.title.x = element_blank())
BaitPlot
```

### Plot the Interactions

<!-- --- -->

<!-- ```{r Behavior x G.Env Plot, echo = F} -->

<!-- BGInteraction<-Crow %>% -->

<!--   group_by(Behavior, G.Env) %>% -->

<!--   summarize(mean = mean(log(Duration)) -->

<!--             , se = sd(log(Duration))/sqrt(length(log(Duration))) -->

<!--             , upper = mean(log(Duration))+sd(log(Duration))/sqrt(length(log(Duration))) -->

<!--             , lower = mean(log(Duration))-sd(log(Duration))/sqrt(length(log(Duration))) -->

<!--             , .groups = 'keep') %>% -->

<!--   ggplot(aes(x = G.Env, y = mean, colour = Behavior))+ -->

<!--   geom_point(position = position_dodge(width = 0.9)) + -->

<!--   geom_text(aes(label = signif(mean)) -->

<!--             , size = 3 -->

<!--             , position=position_dodge2(width=0.9, preserve='single') -->

<!--             , hjust = 1.1 -->

<!--             , vjust = -.5)+ -->

<!--   geom_errorbar(aes(ymin=lower, ymax=upper) -->

<!--                 , width = 0.2 -->

<!--                 , position = position_dodge(width=.9))+ -->

<!--   theme_bw()+ -->

<!--   theme(text=element_text(size = 14))+ -->

<!--   labs(colour = "Behavior")+ -->

<!--   xlab("")+ -->

<!--   ylab("Mean log Bout Duration")+ -->

<!--   scale_colour_manual(values = c("orange", "blue", "green") -->

<!--                     , labels = c("Foraging", "Alert", "Moving"))+ -->

<!--   scale_x_discrete(labels=c("Commercial", "Green Area"))+ -->

<!--   theme(legend.position = 'bottom') -->

<!-- BGInteraction -->

<!-- ``` -->

<!-- > The error bars here represent the standard error. It looks like in both environments, the duration of bouts of alertness are smaller than the other two behaviors. Interestingly, in Green areas, individuals decrease the durations of bouts of movement and increase the duration of bouts of foraging. Let's next check out the interaction between the presence of a sentinel and generalized environment. -->

```{r Sentinel x G.Env Plot, echo = F}
SGInteraction<- Crow %>%
  summarySE(measurevar = "Duration"
            , groupvars = c("G.Env", "Sentinel")) %>%
  ggplot(aes(x = G.Env, y = Duration, colour = Sentinel))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_text(aes(label = signif(Duration))
            , size = 3
            , position=position_dodge2(width=0.9, preserve='single')
            , hjust = -0.5
            , vjust = 0.1)+
  geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
                , width = 0.2
                , position = position_dodge(width=.9))+
  theme_bw()+
  theme(text=element_text(size = 14))+
  labs(colour = "Sentinel")+
  xlab("")+
  ylab("Mean Bout Duration (s)")+
  scale_colour_manual(values = cbPalette
                    , labels = c("Sentinel Absent", "Sentinel Present"))+
  scale_x_discrete(labels=c("Commercial", "Green Area"))+
  theme(legend.position = 'bottom'
        , axis.title.x = element_blank())
SGInteraction
```

> Very cool! I believe I'm getting better at this (or I make good annotations on other scripts...). There is a difference in bout duration between foragers in the presence and absence of a sentinel. The difference is reduced in green areas. Again, the error bars are the standard error. KMG: Cool. You'll want to think about what exactly do you want to present. You can generate all sorts of figures. I personally think these means and SE plots are better. You can add the raw data in as a cloud of points, or you can add histograms sideways next to the plots as well.

### Effect Size

> I next plotted the effect size for each effect.

```{r Duration Effect Size, echo = F}
sjPlot::plot_model(RLMM
                   , show.values=T
                   , show.p=T
                   , value.offset = 0.25
                   , value.size = 4
                   , title = "Effects of generalized environment, number of crows and the presence of a sentinel on the duration of behaviors")
```

> Interesting... According to my model, Alert behavior had a significant negative effect on duration, while the presence of a sentinel and Green Areas have a significant positive effect on duration. The interaction between Moving and Green areas also had a significant negative effect on bout duration. I'm still relatively green when it comes to LMMs, so this may be evident, but I'll make my assumptions clear. From the previous figure, I see that "Head Down", G.Env[Commercial], Baited[N], etc are missing. Therefore I assume that the comparisons are made with those. #KMG: THat is correct. Therefore, I can conclude that: - Crows spend significantly less time with their head's up than down. - Crows spend significantly less time moving than with their heads down. - The duration of behaviors decreases significantly in the absence of a sentinel. - The duration of behaviors decreases significantly in Commercial areas. - The duration of behaviors decreases significantly in baited sites. KMG: When you present these kinds of things, its helpful to write in the same direction. So instead of significantly less and increases, perhaps significantly more and increases, or significantly less and decreases. - There was an interaction between the presence of a sentinel and the generalized environment. That being said, many papers start removing variables to fine tune their model. This is something I am considering in order to better identify the effects of the presence of a sentinel and the type of environment in which the individual forages in.

### What about the group size?

> From the previous results, the number of crows had little effect on the individual vigilance of foragers. This is slightly unexpected, since we would assume there is safety in numbers. By that logic, and hypotheses like the 'many eyes' hypothesis, we would expect to observe a decrease in the duration of some behaviors, namely "Alert". To test this, I decided to plot the model's estimates on the duration of behavior "Alert" to see if there is any trend.

```{r Effects of Group Size, echo = F}
##### Plot estimates with data - Number of crows #####
effects_Num<-effects::effect(term="Number.of.Crows", mod = RLMM)
summary(effects_Num)
Num<-as.data.frame(effects_Num)
Num.crows<-ggplot() +
  geom_point(data = subset(Crow, Behavior == "HU")
             , aes(x = Number.of.Crows, y = log(Duration))
                   , color = "black") + 
  geom_point(data = Num
             , aes(x = Number.of.Crows, y = fit)
             , color = "blue") +
  geom_line(data = Num
            , aes( x = Number.of.Crows, y = fit, group = 1)
            , color = "blue") +
  geom_ribbon(data = Num
              , aes( x = Number.of.Crows, ymin = lower, ymax = upper, group = 1)
              , alpha = 0.3
              , fill = "blue") + 
  labs(x="Number of crows (centered and scaled)"
       , y = "log Duration (s)")
Num.crows
```

> Looks like the data does not follow the predictions from the model. I therefore conclude that the group size had no effect on the duration of alert behavior, at any group size (that I tested).

# Proportion Data

```{r Proportion setup, include = F}
Proportion<-read.csv("Proportions.csv", stringsAsFactors = T)
Proportion$ID<-as.factor(Proportion$ID)
Proportion$AProp<-asin(sqrt(Proportion$Proportion))
```

```{r Proportion Double Check, echo = F}
str(Proportion)
```

> All variables are correctly categorized

## Stacked barplot for proportion

> I also decided to make a stacked barplot to show the proportion of total time spent performing each behavior.

```{r Proportion Summary, echo = F}
MeanProp<-summarySE(Proportion, "Proportion", groupvars = c("G.Env", "Behavior", "Sentinel"))
MeanProp
```

> Summary values are good, but graphs are better. Let's plot a stacked bar graph.

```{r Proportion Barplot, echo = F}
Prop<-ggplot(MeanProp, aes(x = Sentinel, y = Proportion, fill = Behavior))+
  geom_bar(stat = 'identity', position = 'stack')+
  geom_text(aes(label = paste0(signif(Proportion*100), "%")), position = position_stack(vjust = 0.5)
            , size = 3)+
  scale_y_continuous(labels = scales::percent) +
  theme_bw()+
  theme(text=element_text(size = 10)
        , axis.title.x = element_blank())+
  ylab("Proportion of time")+
  scale_x_discrete(labels=c("Sentinel Absent", "Sentinel Present"))+
  scale_fill_manual(values = cbPalette, labels = c("Foraging", "Alert", "Moving"))+
  theme(legend.position = "bottom")+
  facet_grid(~G.Env)
Prop
```

> The proportion figure shows something interesting: crows spent longer times with their heads down in green areas than in commercial ones. Additionally, in the absence of a sentinel, individuals in green areas seemingly increased the proportion of time spent 'alert'.\
> \
> This could be caused by food being easier to find in commercial areas. Crows in green areas need to spend more time looking for food. This reinforces the need to include "Baited" in the model. In the absence of a sentinel, individuals are expected to rely more on individual vigilance, explaining the apparent increase in the proportion of time spent alert in green areas with no sentinels.

## To transform, or not to transform?

> In order to determine whether our fixed effects have significant effects on the allocation of time to each behavior, I decided to first arcsin square root transform the proportions.

```{r Proportion Histograms Individual, include = F}
Untransformed.Prop<-hist(Proportion$Proportion)
Transformed.Prop<-hist(Proportion$AProp, main="Transformed")
```

```{r Proportion Histograms Combined, echo=F}
op <- par(mfrow = c(1,2))
plot(Untransformed.Prop, main="Untransformed", ylab="Frequency", xlab="Proportion")
plot(Transformed.Prop, main="Transformed", ylab="", xlab="Asin(Sqrt(Proportion)")
par(op)
```

> Transformed data seems more normal than untransformed data. Will continue to use transformed data.

## Simple Model

> Let's plot the simple model and test the assumptions.

```{r Proportion Simple Assumptions, echo = F}
Simple.Prop <-lmer(AProp~Behavior+Sentinel*G.Env+Baited+(1|ID), data = Proportion)
plot(Simple.Prop)
```

```{r Is Singular?}
isSingular(Simple.Prop)
```

> Not great. This means my model could have collinearity in its fixed effects (two variables explaining the same thing). Let's try dropping fixed effects one at a time.

```{r Drop one}
drop1(Simple.Prop, test="Chisq")

```

> Ah ha! Behavior completely explains the proportion of time allocated to each behavior, causing the singularity error to appear. Not sure where to proceed from here. One thing you can do is make different models and then do AIC on them to find the best fit model.

## AIC

```{r AIC}
Simple.Prop2 <- lmer(AProp~Behavior+Sentinel*G.Env+(1|ID), data = Proportion)
#Also, I would make a separate column for your transformed data in your dataframe. Doing the calculations inline with the functions can get confusing with all the () you have to keep track of.
Simple.Prop3 <- lmer(asin(sqrt(Proportion))~Behavior+Sentinel +(1|ID), data = Proportion)
Simple.Prop4 <- lmer(asin(sqrt(Proportion))~Behavior+G.Env +(1|ID), data = Proportion)
Simple.Prop5 <- lmer(asin(sqrt(Proportion))~Behavior+(1|ID), data = Proportion)

anova(Simple.Prop, Simple.Prop2, Simple.Prop3, Simple.Prop4, Simple.Prop5)
#Don't worry about the warning if you get it. I just have to change a parameter. Basically, the model with the lowest AIC is the one with only behaviour. So you can confidently say that the type of behaviour has an effect on the proportion of time in a behaviour, and nothing else matters. Think about what this means together with duration. Why things have an effect on duration but not proportion?
```

<!-- --- -->

<!-- ```{r Proportion Simple Model Results, include = F} -->

<!-- Anova(Simple.Prop) -->

<!-- ranova(Simple.Prop) -->

<!-- ``` -->

<!-- > Not sure how to interpret the last output. However, behavior has a significant effect. Let's use the robust LMM to fit the model. -->

<!-- ```{r Robust Proportion, include = F} -->

<!-- RLMM.Prop<-rlmer(asin(sqrt(Proportion))~Behavior*Sentinel*G.Env + (1|ID), data = Proportion) -->

<!-- plot(RLMM.Prop) -->

<!-- ``` -->

<!-- > Looks... ok... Let's see the output -->

<!-- ```{r Summary RLMM.Prop, include = F} -->

<!-- ##### Table of results ##### -->

<!-- sjPlot::tab_model(RLMM.Prop -->

<!--                   , show.re.var = T -->

<!--                   , title = "Effects of generalized environment, number of crows and the presence of a sentinel on the proportion of time" -->

<!--                   , dv.labels = " Effects on proportion") -->

<!-- ##### Effect Size ##### -->

<!-- sjPlot::plot_model(RLMM.Prop -->

<!--                    , show.values=T -->

<!--                    , show.p=T -->

<!--                    , value.offset = 0.5 -->

<!--                    , wrap.title = 70 -->

<!--                    , value.size = 2 -->

<!--                    , title = "Effects of generalized environment, number of crows and the presence of a sentinel on the proportion of time") -->

<!-- ``` -->

<!-- > Curious... Contrary to the simple model, the robust LMM does not identify behavior as a significant effect. The interaction between Moving and G.Env is significant. -->

<!-- --- -->

# Peck rate

> Using the peck data can allow us to infer if crows foraging effectiveness. More pecks per minute would imply that crows forage more actively. To analyse this, I used a linear mixed model with G.Env and presence of a sentinel and its interaction as fixed effects, and observation ID as the random effect.
>
> I calculated the pecks per minute of each individual recorded by dividing the number of pecks by the total recorded duration (excluded time out of frame).

```{r Peck Setup, include=F}
Peck<-read.csv("Peck.csv", stringsAsFactors = T)
Peck$ID<-as.factor(Peck$ID)
Peck$Sentinel<-as.factor(Peck$Sentinel)
Peck$Num.Rec<-as.factor(Peck$Num.Rec)
Peck<-rename(Peck, Number.of.Crows = Num.Rec)
Peck<-rename(Peck, PPM = Pecks.Per.Min)
Peck$LPek<-log(Peck$PPM)
```

```{r Peck Double Check, echo = F}
str(Peck)
```

> Just making sure! Seems like everything is ok.

## Histogram & Transformation

> First, I made a histogram and decided if transformation was appropriate.

```{r Peck Per Minute Histograms Individual, include = F}
Untransformed.PPM<-hist(Peck$PPM, breaks = 12)
Transformed.PPM<-hist(Peck$LPek, breaks = 12)
```

```{r Peck Per Minute Histograms Combined, echo=F}
op <- par(mfrow = c(1,2))
plot(Untransformed.PPM, main="Untransformed", ylab="Frequency", xlab="Pecks per minute")
plot(Transformed.PPM, main="Transformed", ylab="", xlab="Log(Pecks per minute)")
par(op)
```

> Some gaps are present in the histogram. Log transformation made the distribution more normal.\
> Let's summarize the results:

## Peck Summary

```{r Peck Summary, echo = F}
### Summarize
MeanPeck<-summarySE(Peck, "PPM", c("G.Env", "Sentinel"))
MeanPeck
```

> Again, summary stats are good, but plots are better. Let's do that. Seems like peck rate is lower in the presence of a sentinel and in green areas. Standard error is relatively equal.

## Peck Plot

```{r Peck Plots, echo = F}
PeckPlot<-ggplot(MeanPeck, aes(x = G.Env, y = PPM, colour = Sentinel))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3)+
  geom_text(aes(label = signif(PPM))
            , size = 3
            , position=position_dodge2(width=1, preserve='single')
            , hjust = -0.5
            , vjust = 0.1)+
  geom_errorbar(aes(ymin=(PPM-se), ymax=(PPM+se))
                , width = 0.2
                , position = position_dodge(.9))+
  theme_bw()+
  labs(colour = "Presence of a sentinel")+
  ylab("Mean pecks per minute")+
  scale_color_manual(values = cbPalette
                       , labels=c("Sentinel Absent", "Sentinel Present")) +
  scale_x_discrete(labels=c("Commercial", "Green Area"))+
  theme(legend.position = 'bottom'
        , legend.title = element_blank()
        , axis.title.x = element_blank()
        , text=element_text(size = 12))
PeckPlot
```

> It appears that there are fewer pecks per minute in Green areas, regardless of the presence of a sentinel.

## Simple Model

> At this point, let's use the model. As before, we are using the presence of a sentinel, the type of environment and their interaction as fixed effects, with the trial ID as a random effect.

```{r Simple peck, echo = F}
LMM.Peck<-lmer(LPek~Sentinel*G.Env+Baited+(1|ID), data = Peck)
plot(LMM.Peck)
```

> Uhh... not good. Need to figure out why.Let's try the robust version of the linear mixed model.

## Robust LMM

```{r Robust Peck, echo = F}
RLMM.Peck<-rlmer(LPek~Sentinel*G.Env+Baited+(1|ID), data = Peck)
plot(RLMM.Peck)
```

> First graph is identical, others seem ok.

```{r Simple Peck Summary}
Anova(LMM.Peck)
ranova(LMM.Peck)
```

> All not significant, including random effect. Presence of a sentinel is marginally insignificant (p.val = 0.05419). Let's try the robust model.

```{r Robust Peck Summary, echo = F}
sjPlot::tab_model(RLMM.Peck
                  , show.re.var = T
                  , title = "Effects of generalized environment, bait, and the presence of a sentinel on the peck rate of foraging crows"
                  , dv.labels = " Effects on log of pecks per minute")
```

> Interesting! Looks like there are no effects on peck rate. Let's plot it. The difference between the robust LMM and the simpler model is likely due to how the p-values are estimated, and how the robust lmm decreases the chance of committing Type I error.

```{r Robust Peck Effect size}
sjPlot::plot_model(RLMM.Peck
                   , show.values=T
                   , show.p=T
                   , value.offset = 0.25
                   , value.size = 4
                   , wrap.title = 70
                   , title = "Effects of generalized environment, bait, and the presence of a sentinel on the peck rate of foraging crows")
```

KMG: Great job! I agree with your methods and conclusions. Looks like you are pretty much done with your analysis! The hard part now will be interpreting the results! For results, my suggestion is to stick with plots of effect size and means w/ SE. Make sure your stats methods parallels your results section, and in the resuts section present everything in the same way/order. It might seem redundant, but it's actually a lot harder to understand if things are in all directions. I think I commented on this above. I have tried to put all my comments with a KMG: in front. I might have forgotten a few. You can check the two most recent histories to see what changes I made.

# Disturbances
To determine whether there were more or fewer disturbances in different areas, I decided to extract the number of disturbances from across my videos, then compare across the same factors.

```{r Disturbance Files, include = F}
Distub<-read.csv("Disturbance.csv", stringsAsFactors = T)
```

```{r Disturbance Summary, echo=F}
summary(Disturb)

### 2023-05-02 Data & Scripts/Calculate Summary Table with SE, SD, CI by grouping variables.R

@@ -1,0 +1,29 @@
+#calculate means and standard errors
+    summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE, conf.interval=.95) {
+    require(doBy)
+
+    # New version of length which can handle NA's: if na.rm==T, don't count them
+    length2 <- function (x, na.rm=FALSE) {
+        if (na.rm) sum(!is.na(x))
+        else       length(x)
+    }
+
+    # Collapse the data
+    formula <- as.formula(paste(measurevar, paste(groupvars, collapse=" + "), sep=" ~ "))
+    datac <- summaryBy(formula, data=data, FUN=c(length2,mean,sd), na.rm=na.rm)
+
+    # Rename columns
+    names(datac)[ names(datac) == paste(measurevar, ".mean",    sep="") ] <- measurevar
+    names(datac)[ names(datac) == paste(measurevar, ".sd",      sep="") ] <- "sd"
+    names(datac)[ names(datac) == paste(measurevar, ".length2", sep="") ] <- "N"
+    
+    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
+    
+    # Confidence interval multiplier for standard error
+    # Calculate t-statistic for confidence interval: 
+    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
+    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
+    datac$ci <- datac$se * ciMult
+    
+    return(datac)
+}

### 2023-05-02 Systematic Review/Alex Popescu - MSc Thesis - SR Results - 2022-06-27.R

@@ -1,0 +1,75 @@
+##### Working Directory #####
+#setwd("~/Brock U - MSc/Thesis/Systematic Review")
+setwd("D:/Systematic Review")
+
+
+##### Installation & dependencies #####
+#install.packages("BiocManager")
+#install.packages("metagear")
+#BiocManager::install("EBImage")
+
+##### Library #####
+library(metagear)
+
+##### Files #####
+#SandB<-read.csv("Alex Popescu - MSc Thesis - S + B_V2.csv", header = TRUE)
+#effort_distribute(SandB,
+                  #reviewers = "Alex 1",
+                  #initialize = TRUE,
+                  #save_split = TRUE)
+#effort_distribute(SandB,
+                  #reviewers = "Alex 2",
+                  #initialize = TRUE,
+                  #save_split = TRUE)
+#effort_distribute(SandB,
+                  #reviewers = "Alex 3",
+                  #initialize = TRUE,
+                  #save_split = TRUE)
+
+#SorVandB<-read.csv("Alex Popescu - MSc Thesis - SorV+b.csv", header = TRUE)
+#effort_distribute(SorVandB,
+                  #reviewers = c("Alex Popescu_1", "Alex Popescu_2", "Alex Popescu_3"),
+                  #initialize = TRUE,
+                  #save_split = TRUE)
+
+##### Metagear #####
+#abstract_screener("effort_Alex 1.csv", "Alex 1",
+                  #highlightKeywords=c("Sentinel", "sentinel"))
+#abstract_screener("effort_Alex 2.csv", "Alex 2",
+                  #highlightKeywords=c("Sentinel", "sentinel"))
+#abstract_screener("effort_Alex 3.csv", "Alex 3",
+                  #highlightKeywords=c("Sentinel", "sentinel"))
+
+#effort_summary(read.csv("effort_Alex 1.csv"))
+
+Effort<-read.csv("effort_Alex Combined_V2.csv")
+Effort_Yes<-read.csv("effort_Alex Yes.csv")
+effort_summary(Effort_Yes)
+effort_summary(Effort)
+
+##### Collect references #####
+#dir.create("metagear_downloads")
+PDFs_collect(aDataFrame = subset(Effort_Yes, Include = "maybe"), DOIcolumn = "DOI"
+             , FileNamecolumn = "STUDY_ID", directory = "metagear_downloads"
+             , WindowsProxy = T, randomize = T, buffer = T,)
+# SUUUUUUUCKS
+# PDFs collected manually
+
[…]

### 2023-05-12 Data & Scripts/MSC R Script.Rmd
March 12th
New version of the R script + PCAs. Includes pathway analyses.

@@
 ---
-title: "R Script V4"
+title: "R Script V5"
 author: "Alex Popescu"
-date: "2023-01-23"
+date: "2023-05-12"
 output: github_document
@@
 
-V4 - Made some revisions and added plots. All colors are colorblind friendly.
+V5 - Major overhaul and data/analysis improvement.
 
@@
 Therefore, the baited crows are expected to move less than non-baited crows.
-> #KMG: I can imagine a reviewer would want to know hwo the baiting was standardized - something to keep in mind when you are writing.
-I removed the observations in agricultural settings because they had no sentinels present.
+>I removed the observations in agricultural settings because they had no sentinels present.
 This could be talked about in the discussion: agricultural fields may lack perches for individuals to act as sentinels.
 Vineyards may have some structures, but the visibility is limited when the vines are laden with fruit.\
-> #KMG: This is something that can go in the methods.
-You had three sites, but one was removed due to lack of sentinals, and then you can have a bit in the discussion.
-I removed observations past Sept 29 because I was told behaviors can change wildly in the winter.
+>I removed observations past Sept 29 because I was told behaviors can change wildly in the winter.
 The number of crows spotted decreased past that date, suggesting the migrants have left or were leaving.
@@
 library("car")
+library("ggplot2")
 library("ggpattern")
@@
 library("doBy")
+library("FactoMineR")
+library("factoextra")
 ```
@@
 Crow$Number.of.Crows<-as.numeric(Crow$Number.of.Crows)
-Crow$LDur<-log(Crow$Duration)
+Crow$CODED_BAIT<-as.numeric(Crow$CODED_BAIT)
+Crow$CODED_G.ENV<-as.numeric(Crow$CODED_G.ENV)
+Crow$CODED_NB_CROWS<-as.numeric(Crow$CODED_NB_CROWS)
+Crow$CODED_SENTINEL<-as.numeric(Crow$CODED_SENTINEL)
+Crow$CODED_BEHAVIOR<-as.numeric(Crow$CODED_BEHAVIOR)
+Crow$LDur<-log(Crow$Duration+1)
 ```
@@
 I set up a histogram with the distibution of durations across all behaviors.
-I had to log-transform the duration to get a better distribution.\
+I had to log-transform the duration to get a better distribution (log(Duration+1)).\
 
@@
 Box
-#KMG suggestion: I'd make the hatch marks the one where sentinal is absent.
-They remind me of X's which makes me think there is no sentinal present.
-Just me, though, so not a big deal.
-Also a reminder to check if these are colour blind firendly-I don't think the blue and green are.
-
-#AP: I swapped the crosshatch in the figure and used a colorblind-friendly palette.
 ```
@@
 
-> #KMG: Is there a way to group the violin plot like the boxplot?
-So the violins for foraging sentinal present/absent are next to each other?
-Will be easier to see your conclusions.
-Also, have you looked at plotting means and SE?
-Here's a plot #Kiyoko plot of means and SE.
-KMG: I've put the function to generate means, SE, SD, and CI in the github.
-With this plot, it's easier to see the differences in the commercial area where the sentinal presence increases the behaviours.
-It's much messier in the green areas.
-I wonder if it's a refuge or visibility thing?
+>These plots are nice, but it is hard to really see the underlying trends. To best visualise them, I'll next plot a dot plot.
+
+### The Dot Plot
 
@@
                      , groupvars = c("G.Env", "Behavior", "Sentinel"))
-#switching the order of groupvars will change how they are grouped.
+
 plot(x = c(1:12)
@@
 > Very cool!
-I modified the violin and boxplot, but I think the last plot is the most informative.
+I think the last plot is the most informative.
 We can see that the presence of a sentinel increases the duration of bouts in commercial areas.
-In green areas, however, this effect is reduced apart from in alert behavior.
+In green areas, however, this effect is not observed.
+In fact, the bouts of Alert behavior seem significantly reduced in green areas and in the presence of a sentinel.
 
 ## The Tests
+
+### PCAs
+```{r Bout Duration PCA packages, include=F}
+library("FactoMineR")
+library("factoextra")
+```
+```{r Bout Duration PCA Book keeping, include=F}
+PCA.Crow.Data<-Crow[,c(4,6,8,10,12:13)]
+head(PCA.Crow.Data)
+```
+```{r Bout Duration PCA, echo=F}
+PCA.Crow<-PCA(PCA.Crow.Data[,-5], ncp=5)
+```
+
+>I'm new to PCAs, so bear with me please.
+From what I see, presence of bait and sentinel, and the number of crows all contibute to PC1, while Generalized environment seems to be contributing to PC2.
+The behavior appears to contribute to PC2, but not a lot.
+
+#### Eigenvalues
+
+>Let's have a look at the eigenvalues.
+
+``` {r Bout Duration PCA eigenvalues, echo = F}
+Crow.eig.val<-get_eigenvalue(PCA.Crow)
+Crow.eig.val
+```
+
+>Pretty small values. The second column shows how much contribution each dimension has to explain the total variance in the data. Let's quickly plot that.
+
+#### Scree Plot
+
+```{r Bout Duration PCA Scree Plot, echo = F}
+fviz_eig(PCA.Crow, addlabels=T, ylim = c(0,40))
+```
+
+>Pretty cool. The first 3 PCs explain ~77% of the variance. Adding the 4th PC explains ~93% of the variation.
+
+#### Contribution of variables to Principal Components
+
+```{r Bout Duration PCA var results, include = F}
+Crow.var<-get_pca_var(PCA.Crow)
+```
+```{r Bout Duration PCA Contribution Dim 1, echo=F}
+fviz_contrib(PCA.Crow, choice="var", axes = 1, top = 20)
+```
+
+>The red dotted line represents the expected average contribution, assuming the contribution of the variables were uniform.
+In other words, the line is equal to 1/the number of variables; in this case 1/5 or 20%.
+
+> From this plot, I can see that, as I saw previously in the correlation circle, the number of crows, the presence of a sentinel and the presence of bait contribute the most to PC1 (Dimension 1)
+
+```{r Bout Duration PCA Contribution Dim 2, echo=F}
+fviz_contrib(PCA.Crow, choice="var", axes = 2, top = 20)
+```
+
+> The generalized environment contributes the most to PC2.
+
+```{r Bout Duration PCA Contribution Dim 3, echo=F}
+fviz_contrib(PCA.Crow, choice="var", axes = 3, top = 20)
+```
+
+>The coded behavior contributes most to PC3.
+
+```{r Bout Duration PCA Contribution Dim 4, echo=F}
+fviz_contrib(PCA.Crow, choice="var", axes = 4, top = 20)
+```
+
+>OK!
+Now we're getting to the interesting bits.
+The presence of bait and the presence of a Sentinel contribute most to PC4.
+This makes me curious to see if there is an interaction between the two factors.
+
+```{r Bout Duration PCA Contribution Dim 5, echo=F}
+fviz_contrib(PCA.Crow, choice="var", axes = 5, top = 20)
+```
+
+>>The presence of a sentinel and the size of the group contribute most to PC 5.
+I believe this would imply that the interaction between the two could be interesting to look at.
+Considering these two variables are also major contributors to PC1, this makes the inclusion of the interaction in the model a must.
+
+>From this, I believe I should look at the interaction between NB_Crows, Presence of a sentinel, and the presence of bait.
+
+#### Plotting with Behaviors
+
+>I'd like to plot the dimensions using the behaviors of the bouts.
+
+```{r Bout Duration PCA Plot with Behavior PC1-2, echo=F}
+fviz_pca_ind(PCA.Crow
+             , geom.ind="point"
+             , col.ind=PCA.Crow.Data$Behavior
+             , palette = cbPalette
+             , addEllipses = T
+             , legend.title = "Behaviors"
+             )
+```
+
+>Messy. The mean points of the behaviors are relatively centered.
+
+```{r Bout Duration PCA Plot with Behavior PC2-3, echo=F}
+fviz_pca_ind(PCA.Crow
+             , axes = c(2,3)
+             , geom.ind="point"
+             , col.ind=PCA.Crow.Data$Behavior
+             , palette = cbPalette
+             , addEllipses = T
+             , legend.title = "Behaviors"
+             )
+```
+
+>Very clear separation by behavior type, with 'Head Up' shifted upward, 'Head Down' shifted downward, and 'Moving' in the middle.
+
+```{r Bout Duration PCA Plot with Behavior PC1-3, echo=F}
+fviz_pca_ind(PCA.Crow
+             , axes = c(1,3)
+             , geom.ind="point"
+             , col.ind=PCA.Crow.Data$Behavior
+             , palette = cbPalette
+             , addEllipses = T
+             , legend.title = "Behaviors"
+             )
+```
+
+>Good separation, but the concentration ellipses have increased in size. Curious.
+
+#### Biplot
+
+>Finally, let's look at the biplot. From what I understand, it is essentially the correlation circle overlaid onto the data.
+
+```{r Bout Duration PCA Biplot PC1-2, echo=F}
+fviz_pca_biplot(PCA.Crow
+                ,geom.ind = "point"
+                , repel = T
+                , col.ind=PCA.Crow.Data$Behavior
+                , addEllipses = T
+                , label = "var"
+                , palette=cbPalette
+                , col.var="black"
+                , legend.title = "Behavior")
+```
+```{r Bout Duration PCA Biplot PC2-3, echo=F}
+fviz_pca_biplot(PCA.Crow
+                , axes = c(2,3)
+                , geom.ind = "point"
+                , repel = T
+                , col.ind=PCA.Crow.Data$Behavior
+                , addEllipses = T
+                , label = "var"
+                , palette=cbPalette
+                , col.var="black"
+                , legend.title = "Behavior")
+```
+```{r Bout Duration PCA Biplot PC1-3, echo=F}
+fviz_pca_biplot(PCA.Crow
+                , axes = c(1,3)
+                , geom.ind = "point"
+                , repel = T
+                , col.ind=PCA.Crow.Data$Behavior
+                , addEllipses = T
+                , label = "var"
+                , palette=cbPalette
+                , col.var="black"
+                , legend.title = "Behavior")
+```
+
+>Nice! I think I've done the PCA analysis to the best of my abilities, but I am very much open to comments and feedback!
+
+>From what I can gather, I should model the data as follows:/
+>LDur~Behavior+NB_Crows*Bait*Sentinel+G.Env+(1|ID)
 
@@
 I then plotted the models to test the assumptions of the model.
-The model I used was "log(Duration) \~ Behavior + (Sentinel\* G.Env) + Number of crows + Baited + (1\|ID)".
-KMG: what is the backslash?
->
+
 > Behavior is expected to be affected by both the presence of a sentinel and the environment the crows are in.
@@
 
-```{r Duration RLMM, echo=F}
-RLMM<-rlmer(LDur~Behavior+Sentinel*G.Env+Number.of.Crows + Baited + (1|ID), data = Crow)
-plot(RLMM)
+```{r Bout Duration Model Book keeping, include=F}
+Crow$CODED_NB_CROWS<-as.factor(Crow$CODED_NB_CROWS)
 ```
 
-> That took AGES to run. There's a big "cliff" on the normal Q-Q vs residuals plot. This means that there are more extreme values than would be expected if they came from a truly normal population.
+```{r Duration RLMM, echo=F}
+RLMM.Dur.I<-rlmer(LDur~Behavior+Sentinel*CODED_NB_CROWS*Baited+G.Env+ (1|ID), data = Crow)
+plot(RLMM.Dur.I)
+```
 
-### Bout Results
+> That took pretty long to run.
+Let's next see the results of the model.
+I don't see huge deviations from the assumptions.
+According to the error message, my fixed-effect model matrix is rank deficient.
+I don't believe I need to worry too much about this warning.
 
 ```{r Duration RLMM Summary, echo = F}
-summary(RLMM)
+summary(RLMM.Dur.I)
 ```
@@
 ```{r Duration Better Summary, echo = F}
-sjPlot::tab_model(RLMM
+sjPlot::tab_model(RLMM.Dur.I
                   , show.re.var = T
-                  , title = "Effects of generalized environment, bait, number of crows and the presence of a sentinel on durations of behaviors"
+                  , title = "Model Results with Interactions"
                   , dv.labels = " Effects of on duration of behaviors")
-#KMG: is baiting not in here as well?
+
 ```
 
-> Like this!
-This is much better!
-Lets try a simpler output.
-Curious...
-Removing the three-way interaction caused many factors to become significant.
-Number of crows remains non-significant.
-#KMG: Can we visualize the the baited effects on duration as well?
+>Oh fun! Adding the three-way interaction caused all but the behavior to be non-significant. Let's try removing the interactions.
+
+```{r Duration RLMM No Interactions, echo=F}
+RLMM.Dur.NoI<-rlmer(LDur~Behavior+Sentinel+CODED_NB_CROWS+Baited+G.Env+ (1|ID), data = Crow)
+sjPlot::tab_model(RLMM.Dur.NoI
+                  , show.re.var = T
+                  , title = "Model Results without Interactions"
+                  , dv.labels = " Effects of on duration of behaviors")
+```
+
+>Same result!
+
+>Let's try the good old regular lmer and see if that changes anything.
 
 ```{r Simpler output, echo = F}
-Simple.Dur<-lmer(LDur~Behavior+Sentinel*G.Env+Number.of.Crows + Baited + (1|ID), data = Crow)
-Anova(Simple.Dur)
-ranova(Simple.Dur)
+Simple.Dur.I<-lmer(LDur~Behavior+Sentinel*CODED_NB_CROWS*Baited + G.Env + (1|ID), data = Crow)
+Simple.Dur.NoI<-lmer(LDur~Behavior+Sentinel+CODED_NB_CROWS+Baited + G.Env + (1|ID), data = Crow)
+anova(Simple.Dur.NoI, Simple.Dur.I)
 ```
 
-> Simple model shows significant effects of Behavior, and the interaction between generalized environment and Behavior, and generalized environment and the presence of a sentinel.
+>Using the simple model, I looked at the AIC values.
+The model without interactions had the lowest AIC value, therefore is preferred.Let's actually see what removing some factors in the model does to the AIC.
+
+#### AIC Testing
+
+```{r Bout Duration Drop1,echo=F}
+drop1(Simple.Dur.NoI)
+```
+
+>Seems like dropping any of the variables will have no effect on the model's results. In other words, I can drop all factors except for 'behavior'.
+
+```{r Bout Duration AIC testing}
+Simplest.Dur<-lmer(LDur~Behavior+(1|ID), data=Crow)
+anova(Simple.Dur.NoI, Simplest.Dur)
+```
+
+>Yep! The simplest model (LDur~Behavio+(1|ID)) is preferred. Let's try subsetting the data.
+
+#### Subset Data
+```{r Bout Duration Subset, include=F}
+HU.Dur<-Crow[which(Crow$Behavior=="HU"),]
+HD.Dur<-Crow[which(Crow$Behavior=="HD"),]
+M.Dur<-Crow[which(Crow$Behavior=="M"),]
+```
+```{r Simple model on Subset HU data, echo=F}
+Simple.HU.Dur<-lmer(LDur~Sentinel+CODED_NB_CROWS+Baited + G.Env + (1|ID), data =HU.Dur)
+Anova(Simple.HU.Dur)
+```
+
+>Simple model on Head Up subset with no interactions shows no significance.
+
+```{r Simple model on Subset HD data, echo=F}
+Simple.HD.Dur<-lmer(LDur~Sentinel+CODED_NB_CROWS+Baited + G.Env + (1|ID), data =HD.Dur)
+Anova(Simple.HD.Dur)
+```
+
+>Simple model on Head Down subset shows a significant effect of foraging group size, presence of bait and generalized environment.
+
+```{r Simple model on Subset M data, echo=F}
+Simple.M.Dur<-lmer(LDur~Sentinel+CODED_NB_CROWS+Baited + G.Env + (1|ID), data =M.Dur)
+Anova(Simple.M.Dur)
+```
+
+> Simple model shows no significant effects.
 
@@
 
-> I generated a plot to better see what the interaction between Generalized Environment and the presence of a sentinel. First, let's make a summary of the results.
+>Ok! Considering only the 'Head Down' subset had significant effects, let's plot those results.
 
-```{r Duration Summary, echo=F}
-MeanDur
-```
+#### Bait
 
-> There are too many combinations to effectively see any trends, let's look at the significant interactions from the models.
-
-### Bait
-
-```{r Bait effect on bout, echo = F}
-BaitPlot<- Crow %>%
+```{r Bait effect on HD bout, echo = F}
+BaitPlot<- HD.Dur %>%
   summarySE(measurevar = "Duration"
@@
   theme(text=element_text(size = 14))+
-  xlab("")+
-  ylab("Mean Bout Duration (s)")+
+  xlab("Presence of Bait")+
+  ylab("Mean Head Down Bout Duration (s)")+
   scale_x_discrete(labels=c("Not Baited", "Baited"))+
   theme(legend.position = 'bottom'
-        , axis.title.x = element_blank())
+        )
 BaitPlot
@@
 
-### Plot the Interactions
+>Fun! The plot shows overlap between the standard error bars. In the presence of bait, crows will decrease the duration of their bouts of 'Head Down'.
 
-<!-- --- -->
+#### G.Env
 
-<!-- ```{r Behavior x G.Env Plot, echo = F} -->
-
-<!-- BGInteraction<-Crow %>% -->
-
-<!--   group_by(Behavior, G.Env) %>% -->
-
-<!--   summarize(mean = mean(log(Duration)) -->
-
-<!--             , se = sd(log(Duration))/sqrt(length(log(Duration))) -->
-
-<!--             , upper = mean(log(Duration))+sd(log(Duration))/sqrt(length(log(Duration))) -->
-
-<!--             , lower = mean(log(Duration))-sd(log(Duration))/sqrt(length(log(Duration))) -->
-
-<!--             , .groups = 'keep') %>% -->
-
-<!--   ggplot(aes(x = G.Env, y = mean, colour = Behavior))+ -->
-
-<!--   geom_point(position = position_dodge(width = 0.9)) + -->
-
-<!--   geom_text(aes(label = signif(mean)) -->
-
-<!--             , size = 3 -->
-
-<!--             , position=position_dodge2(width=0.9, preserve='single') -->
-
-<!--             , hjust = 1.1 -->
-
-<!--             , vjust = -.5)+ -->
-
-<!--   geom_errorbar(aes(ymin=lower, ymax=upper) -->
-
-<!--                 , width = 0.2 -->
-
-<!--                 , position = position_dodge(width=.9))+ -->
-
-<!--   theme_bw()+ -->
-
-<!--   theme(text=element_text(size = 14))+ -->
-
-<!--   labs(colour = "Behavior")+ -->
-
-<!--   xlab("")+ -->
-
-<!--   ylab("Mean log Bout Duration")+ -->
-
-<!--   scale_colour_manual(values = c("orange", "blue", "green") -->
-
-<!--                     , labels = c("Foraging", "Alert", "Moving"))+ -->
-
-<!--   scale_x_discrete(labels=c("Commercial", "Green Area"))+ -->
-
-<!--   theme(legend.position = 'bottom') -->
-
-<!-- BGInteraction -->
-
-<!-- ``` -->
-
-<!-- > The error bars here represent the standard error.
-It looks like in both environments, the duration of bouts of alertness are smaller than the other two behaviors.
-Interestingly, in Green areas, individuals decrease the durations of bouts of movement and increase the duration of bouts of foraging.
-Let's next check out the interaction between the presence of a sentinel and generalized environment.
--->
-
-```{r Sentinel x G.Env Plot, echo = F}
-SGInteraction<- Crow %>%
+```{r G.Env effect on HD bout, echo = F}
+EnvPlot<- HD.Dur %>%
   summarySE(measurevar = "Duration"
-            , groupvars = c("G.Env", "Sentinel")) %>%
-  ggplot(aes(x = G.Env, y = Duration, colour = Sentinel))+
+            , groupvars = "G.Env") %>%
+  ggplot(aes(x = G.Env, y = Duration))+
   geom_point(position = position_dodge(width = 0.9)
@@
   theme(text=element_text(size = 14))+
-  labs(colour = "Sentinel")+
-  xlab("")+
-  ylab("Mean Bout Duration (s)")+
-  scale_colour_manual(values = cbPalette
-                    , labels = c("Sentinel Absent", "Sentinel Present"))+
-  scale_x_discrete(labels=c("Commercial", "Green Area"))+
+  xlab("Generalized Environment")+
+  ylab("Mean Head Down Bout Duration (s)")+
   theme(legend.position = 'bottom'
-        , axis.title.x = element_blank())
-SGInteraction
+        )
+EnvPlot
 ```
 
-> Very cool!
-I believe I'm getting better at this (or I make good annotations on other scripts...).
-There is a difference in bout duration between foragers in the presence and absence of a sentinel.
-The difference is reduced in green areas.
-Again, the error bars are the standard error.
-KMG: Cool.
-You'll want to think about what exactly do you want to present.
-You can generate all sorts of figures.
-I personally think these means and SE plots are better.
-You can add the raw data in as a cloud of points, or you can add histograms sideways next to the plots as well.
+>Very nice! We can see that in Green Areas, the bout duration of head down behavior is increased. The error bars, as before, are standard error.
 
-### Effect Size
+#### Foraging Group Size
 
-> I next plotted the effect size for each effect.
-
-```{r Duration Effect Size, echo = F}
-sjPlot::plot_model(RLMM
-                   , show.values=T
-                   , show.p=T
-                   , value.offset = 0.25
-                   , value.size = 4
-                   , title = "Effects of generalized environment, number of crows and the presence of a sentinel on the duration of behaviors")
+```{r NB_Crows effect on HD bout, echo = F}
+NBPlot<- HD.Dur %>%
+  summarySE(measurevar = "Duration"
+            , groupvars = "CODED_NB_CROWS") %>%
+  ggplot(aes(x = CODED_NB_CROWS, y = Duration))+
+  geom_point(position = position_dodge(width = 0.9)
+             , size = 3) +
+  geom_text(aes(label = signif(Duration))
+            , size = 3
+            , position=position_dodge2(width=0.9, preserve='single')
+            , hjust = -0.5
+            , vjust = 0.1)+
+  geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
+                , width = 0.2
+                , position = position_dodge(width=.9))+
+  theme_bw()+
+  theme(text=element_text(size = 14))+
+  xlab("Group Size")+
+  ylab("Mean Head Down Bout Duration (s)")+
+  scale_x_discrete(labels=c("Small", "Large"))+
+  theme(legend.position = 'bottom'
+        )
+NBPlot
 ```
 
-> Interesting...
-According to my model, Alert behavior had a significant negative effect on duration, while the presence of a sentinel and Green Areas have a significant positive effect on duration.
-The interaction between Moving and Green areas also had a significant negative effect on bout duration.
-I'm still relatively green when it comes to LMMs, so this may be evident, but I'll make my assumptions clear.
-From the previous figure, I see that "Head Down", G.Env[Commercial], Baited[N], etc are missing.
-Therefore I assume that the comparisons are made with those.
-#KMG: THat is correct.
-Therefore, I can conclude that: - Crows spend significantly less time with their head's up than down.
-- Crows spend significantly less time moving than with their heads down.
-- The duration of behaviors decreases significantly in the absence of a sentinel.
-- The duration of behaviors decreases significantly in Commercial areas.
-- The duration of behaviors decreases significantly in baited sites.
-KMG: When you present these kinds of things, its helpful to write in the same direction.
-So instead of significantly less and increases, perhaps significantly more and increases, or significantly less and decreases.
-- There was an interaction between the presence of a sentinel and the generalized environment.
-That being said, many papers start removing variables to fine tune their model.
-This is something I am considering in order to better identify the effects of the presence of a sentinel and the type of environment in which the individual forages in.
+>Very nice! In larger foraging groups, individuals increase their bouts of head down behavior.
 
-### What about the group size?
-
-> From the previous results, the number of crows had little effect on the individual vigilance of foragers.
-This is slightly unexpected, since we would assume there is safety in numbers.
-By that logic, and hypotheses like the 'many eyes' hypothesis, we would expect to observe a decrease in the duration of some behaviors, namely "Alert".
-To test this, I decided to plot the model's estimates on the duration of behavior "Alert" to see if there is any trend.
-
-```{r Effects of Group Size, echo = F}
-##### Plot estimates with data - Number of crows #####
-effects_Num<-effects::effect(term="Number.of.Crows", mod = RLMM)
-summary(effects_Num)
-Num<-as.data.frame(effects_Num)
-Num.crows<-ggplot() +
-  geom_point(data = subset(Crow, Behavior == "HU")
-             , aes(x = Number.of.Crows, y = log(Duration))
-                   , color = "black") + 
-  geom_point(data = Num
-             , aes(x = Number.of.Crows, y = fit)
-             , color = "blue") +
-  geom_line(data = Num
-            , aes( x = Number.of.Crows, y = fit, group = 1)
-            , color = "blue") +
-  geom_ribbon(data = Num
-              , aes( x = Number.of.Crows, ymin = lower, ymax = upper, group = 1)
-              , alpha = 0.3
-              , fill = "blue") + 
-  labs(x="Number of crows (centered and scaled)"
-       , y = "log Duration (s)")
-Num.crows
-```
-
-> Looks like the data does not follow the predictions from the model. I therefore conclude that the group size had no effect on the duration of alert behavior, at any group size (that I tested).
+>I can therefore make the following conclusions:
+>-Crows do not modify the duration of bouts of head up or moving
+>-Crows will modify the duration of bouts of head down behavior.
+>-In larger groups, crows will increase the duration of bouts of 'Head Down", possibly due to the many eyes hypothesis.
+>-In the presence of Bait, crows will decrease the duration of bouts of 'Head Down'.
+This could be due to the larger concentration of food in the patch, and therefore less time required to look for food.
+>-In Green Areas, crows will increase the duration of bouts of 'Head Down'.
 
@@
 ```{r Proportion setup, include = F}
-Proportion<-read.csv("Proportions.csv", stringsAsFactors = T)
-Proportion$ID<-as.factor(Proportion$ID)
-Proportion$AProp<-asin(sqrt(Proportion$Proportion))
+DATA<-read.csv("DATA.csv", stringsAsFactors = T)
+DATA$ID<-as.factor(DATA$ID)
+DATA$ASIN_PROPORTION<-asin(sqrt(DATA$BEHAVIOR_PROPORTION))
 ```
@@
 ```{r Proportion Double Check, echo = F}
-str(Proportion)
+str(DATA)
 ```
@@
 ```{r Proportion Summary, echo = F}
-MeanProp<-summarySE(Proportion, "Proportion", groupvars = c("G.Env", "Behavior", "Sentinel"))
+MeanProp<-summarySE(DATA, "BEHAVIOR_PROPORTION", groupvars = c("GENERALIZED_ENVIRONMENT", "BEHAVIOR", "SENTINEL_PRESENCE"))
 MeanProp
@@
 ```{r Proportion Barplot, echo = F}
-Prop<-ggplot(MeanProp, aes(x = Sentinel, y = Proportion, fill = Behavior))+
+Prop<-ggplot(MeanProp, aes(x = SENTINEL_PRESENCE, y = BEHAVIOR_PROPORTION, fill = BEHAVIOR))+
   geom_bar(stat = 'identity', position = 'stack')+
-  geom_text(aes(label = paste0(signif(Proportion*100), "%")), position = position_stack(vjust = 0.5)
+  geom_text(aes(label = paste0(signif(BEHAVIOR_PROPORTION*100), "%")), position = position_stack(vjust = 0.5)
             , size = 3)+
@@
   scale_x_discrete(labels=c("Sentinel Absent", "Sentinel Present"))+
-  scale_fill_manual(values = cbPalette, labels = c("Foraging", "Alert", "Moving"))+
+  scale_fill_manual(values = cbPalette, labels = c("Foraging", "Alert", "Moving"), name="Behavior")+
   theme(legend.position = "bottom")+
-  facet_grid(~G.Env)
+  facet_grid(~GENERALIZED_ENVIRONMENT)
 Prop
@@
 
-> The proportion figure shows something interesting: crows spent longer times with their heads down in green areas than in commercial ones.
+> The proportion figure shows something interesting: crows seem to spend longer times with their heads down in green areas than in commercial ones.
 Additionally, in the absence of a sentinel, individuals in green areas seemingly increased the proportion of time spent 'alert'.\
@@
 ```{r Proportion Histograms Individual, include = F}
-Untransformed.Prop<-hist(Proportion$Proportion)
-Transformed.Prop<-hist(Proportion$AProp, main="Transformed")
+Untransformed.Prop<-hist(DATA$BEHAVIOR_PROPORTION, breaks=25)
+Transformed.Prop<-hist(DATA$ASIN_PROPORTION, breaks=25)
 ```
@@
 
+## PCAs
+
+>Let's run another PCA
+
+```{r DATA PCA SHEET, include=F}
+PCA.DATA.SHEET<-DATA[,c(5,7,10,14,16,18,22,23,24,39)]
+head(PCA.DATA.SHEET)
+str(PCA.DATA.SHEET)
+```
+```{R DATA PCA, echo=F}
+PCA.DATA<-PCA(PCA.DATA.SHEET[,c(-8,-9)])
+```
+
+>There are many variables here. I won't bother describing them here.
+
+>Since the presence of a sentinel is not "God-Given", we'll run another PCA without that factor.
+
+```{r DATA PCA NO SENTINEL, echo=F}
+PCA.DATA.NOSENT<-PCA(PCA.DATA.SHEET[,c(-5,-8,-9)])
+```
+
+>Very similar to the one above.
+
+### Scree Plot
+
+```{r DATA PCA Scree Plot, echo = F}
+get_eigenvalue(PCA.DATA)
+fviz_eig(PCA.DATA, addlabels=T, ylim = c(0,40))
+```
+
+>The first PC covers ~40% of the variation. The first 4 PCs cover ~82% of the variation.
+
+```{r DATA PCA Scree Plot NO SENTINEL, echo = F}
+get_eigenvalue(PCA.DATA.NOSENT)
+fviz_eig(PCA.DATA.NOSENT, addlabels=T, ylim = c(0,40))
+```
+
+>Excluding sentinel, the first PC covers ~43% of the variation, and the first 4 PCs cover ~86% of the total variation.
+
+### Contribution of variables to Principal Components
+
+```{r PCA DATA var results, include = F}
+DATA.var<-get_pca_var(PCA.DATA)
+DATA.NOSENT.var<-get_pca_var(PCA.DATA.NOSENT)
+```
+```{r PCA DATA Contribution Dim 1, echo=F}
+fviz_contrib(PCA.DATA, choice="var", axes = 1, top = 12.5)
+```
+
+>Very weird, Julian Date, Decimal time, group size and generalized environment contribute greatly to PC1.
+
+```{r PCA DATA Contribution Dim 2, echo=F}
+fviz_contrib(PCA.DATA, choice="var", axes = 2, top = 12.5)
+```
+
+>Frequency of disturbances, group size (again), temperature and the presence of a sentinel contribute most to PC2.
+
+```{r PCA DATA Contribution Dim 3, echo=F}
+fviz_contrib(PCA.DATA, choice="var", axes = 3, top = 12.5)
+```
+
+>Temperature, Bait presence and decimal time contribute most to PC3.
+
+```{r PCA DATA Contribution Dim 4, echo=F}
+fviz_contrib(PCA.DATA, choice="var", axes = 4, top = 12.5)
+```
+
+>Sentinel presence, environment and frequency of disturbances contribute most to PC4.
+
+>I'm still a little unclear as to how I should interpret these results.
+
+#### No Sentinel
+
+```{r PCA DATA.NOSENT Contribution Dim 1, echo=F}
+fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 1, top = 14.29)
+```
+
+>Same 4 factors contribute the most to PC1 in the absence of the sentinel factor.
+
+```{r PCA DATA.NOSENT Contribution Dim 2, echo=F}
+fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 2, top = 14.29)
+```
+
+>Same as above, frequency of disturbances, temperature and group size contribute most to PC2 in the absence of the sentinel factor.
+
+```{r PCA DATA.NOSENT Contribution Dim 3, echo=F}
+fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 3, top = 14.29)
+```
+
+>Temperature, Bait presence and frequency of disturbances contribute the most to PC3 in the absence of a sentinel.
+
+```{r PCA DATA.NOSENT Contribution Dim 4, echo=F}
+fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 4, top = 14.29)
+```
+
+>Environment contributes most to PC4 in the absence of a sentinel.
+
+### BiPlots
+```{r DATA PCA Biplot PC1-2, echo=F}
+fviz_pca_biplot(PCA.DATA
+                ,geom.ind = "point"
+                , repel = T
+                , col.ind=PCA.DATA.SHEET$BEHAVIOR
+                , addEllipses = T
+                , label = "var"
+                , palette=cbPalette
+                , col.var="black"
+                , legend.title = "Behavior")
+```
+
+>Messy, can't see where the behaviors lie.
+
+```{r DATA PCA Biplot PC1-4, echo=F}
+fviz_pca_biplot(PCA.DATA
+                , axes = c(1,4)
+                , geom.ind = "point"
+                , repel = T
+                , col.ind=PCA.DATA.SHEET$BEHAVIOR
+                , addEllipses = T
+                , label = "var"
+                , palette=cbPalette
+                , col.var="black"
+                , legend.title = "Behavior")
+```
+
+>Still messy.
+
+```{r DATA PCA Biplot PC2-3, echo=F}
+fviz_pca_biplot(PCA.DATA
+                , axes = c(2,3)
+                , geom.ind = "point"
+                , repel = T
+                , col.ind=PCA.DATA.SHEET$BEHAVIOR
+                , addEllipses = T
+                , label = "var"
+                , palette=cbPalette
+                , col.var="black"
+                , legend.title = "Behavior")
+```
+
+>Biplots are messy if we don't include behavior.
+
+>Same is observed if we run the biplot in the absence of the sentinel factor.
+
+>I'm not sure how to interpret these results.
+
 ## Simple Model
 
-> Let's plot the simple model and test the assumptions.
+> Let's plot the simple model and test the assumptions. I'm using the same fixed effects as in the bout duration analyses. I am also adding some of the factors identified in the PCAs.
 
 ```{r Proportion Simple Assumptions, echo = F}
-Simple.Prop <-lmer(AProp~Behavior+Sentinel*G.Env+Baited+(1|ID), data = Proportion)
+Simple.Prop <-lmer(ASIN_PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+TOTAL_FREQUENCY_OF_DISTURBANCES+(1|ID)+(1|JULIAN_DATE)+(1|DECIMAL_TIME)+(1|TEMPERATURE), data = DATA)
 plot(Simple.Prop)
@@
 
-```{r Is Singular?}
-isSingular(Simple.Prop)
+> Not great. The warning I got means my model could have collinearity in its fixed effects (two variables explaining the same thing). Let's try dropping fixed effects one at a time.
+
+>Let's see the results before we do anything.
+
+```{r Proportion Model Results, echo = F}
+Anova(Simple.Prop)
+ranova(Simple.Prop)
 ```
 
-> Not great. This means my model could have collinearity in its fixed effects (two variables explaining the same thing). Let's try dropping fixed effects one at a time.
+>Interesting. For one, the results are insignificant. The other, the ranova results suggest that the model without random effects is preferred.
 
@@
 
-> Ah ha!
-Behavior completely explains the proportion of time allocated to each behavior, causing the singularity error to appear.
-Not sure where to proceed from here.
-One thing you can do is make different models and then do AIC on them to find the best fit model.
+>Removing factors has no significant effects on the model's output.
 
-## AIC
+>Let's go to an even simpler model
 
-```{r AIC}
-Simple.Prop2 <- lmer(AProp~Behavior+Sentinel*G.Env+(1|ID), data = Proportion)
-#Also, I would make a separate column for your transformed data in your dataframe. Doing the calculations inline with the functions can get confusing with all the () you have to keep track of.
-Simple.Prop3 <- lmer(asin(sqrt(Proportion))~Behavior+Sentinel +(1|ID), data = Proportion)
-Simple.Prop4 <- lmer(asin(sqrt(Proportion))~Behavior+G.Env +(1|ID), data = Proportion)
-Simple.Prop5 <- lmer(asin(sqrt(Proportion))~Behavior+(1|ID), data = Proportion)
-
-anova(Simple.Prop, Simple.Prop2, Simple.Prop3, Simple.Prop4, Simple.Prop5)
-#Don't worry about the warning if you get it.
-I just have to change a parameter.
-Basically, the model with the lowest AIC is the one with only behaviour.
-So you can confidently say that the type of behaviour has an effect on the proportion of time in a behaviour, and nothing else matters.
-Think about what this means together with duration.
-Why things have an effect on duration but not proportion?
+### The even simpler models
+```{r Proportion Simpler Models, echo = F}
+Simpler.Prop1 <-lmer(ASIN_PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+(1|ID), data = DATA)
+Simpler.Prop2 <-lmer(ASIN_PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+(1|ID), data = DATA)
+Simpler.Prop3 <-lmer(ASIN_PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+(1|ID), data = DATA)
+Simpler.Prop4 <-lmer(ASIN_PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+(1|ID), data = DATA)
+Simpler.Prop5 <-lmer(ASIN_PROPORTION~BEHAVIOR+(1|ID), data = DATA)
+anova(Simpler.Prop1,Simpler.Prop2,Simpler.Prop3,Simpler.Prop4,Simpler.Prop5)
 ```
 
-<!-- --- -->
+>The simplest model, the one with only behavior as a FE and ID as a RE has the lowest AIC value. Weird...
 
-<!-- ```{r Proportion Simple Model Results, include = F} -->
+```{r Proportion Simper Model 5 Results, echo = F}
+Anova(Simpler.Prop5)
+```
 
-<!-- Anova(Simple.Prop) -->
-
-<!-- ranova(Simple.Prop) -->
-
-<!-- ``` -->
-
-<!-- > Not sure how to interpret the last output. However, behavior has a significant effect. Let's use the robust LMM to fit the model. -->
-
-<!-- ```{r Robust Proportion, include = F} -->
-
-<!-- RLMM.Prop<-rlmer(asin(sqrt(Proportion))~Behavior*Sentinel*G.Env + (1|ID), data = Proportion) -->
-
-<!-- plot(RLMM.Prop) -->
-
-<!-- ``` -->
-
-<!-- > Looks... ok... Let's see the output -->
-
-<!-- ```{r Summary RLMM.Prop, include = F} -->
-
-<!-- ##### Table of results ##### -->
-
-<!-- sjPlot::tab_model(RLMM.Prop -->
-
-<!--                   , show.re.var = T -->
-
-<!--                   , title = "Effects of generalized environment, number of crows and the presence of a sentinel on the proportion of time" -->
-
-<!--                   , dv.labels = " Effects on proportion") -->
-
-<!-- ##### Effect Size ##### -->
-
-<!-- sjPlot::plot_model(RLMM.Prop -->
-
-<!--                    , show.values=T -->
-
-<!--                    , show.p=T -->
-
-<!--                    , value.offset = 0.5 -->
-
-<!--                    , wrap.title = 70 -->
-
-<!--                    , value.size = 2 -->
-
-<!--                    , title = "Effects of generalized environment, number of crows and the presence of a sentinel on the proportion of time") -->
-
-<!-- ``` -->
-
-<!-- > Curious... Contrary to the simple model, the robust LMM does not identify behavior as a significant effect. The interaction between Moving and G.Env is significant. -->
-
-<!-- --- -->
+>Behavior is still insignificant. Could this be caused by the data transformation?
 
@@
 >
-> I calculated the pecks per minute of each individual recorded by dividing the number of pecks by the total recorded duration (excluded time out of frame).
+> I calculated the pecks per minute of each individual recorded by dividing the number of pecks by the total head down duration (excluded time out of frame).
 
 ```{r Peck Setup, include=F}
-Peck<-read.csv("Peck.csv", stringsAsFactors = T)
-Peck$ID<-as.factor(Peck$ID)
-Peck$Sentinel<-as.factor(Peck$Sentinel)
-Peck$Num.Rec<-as.factor(Peck$Num.Rec)
-Peck<-rename(Peck, Number.of.Crows = Num.Rec)
-Peck<-rename(Peck, PPM = Pecks.Per.Min)
-Peck$LPek<-log(Peck$PPM)
+DATA$LPek<-log(DATA$PECK_RATE)
 ```
-
-```{r Peck Double Check, echo = F}
-str(Peck)
-```
-
-> Just making sure! Seems like everything is ok.
 
@@
 ```{r Peck Per Minute Histograms Individual, include = F}
-Untransformed.PPM<-hist(Peck$PPM, breaks = 12)
-Transformed.PPM<-hist(Peck$LPek, breaks = 12)
+Untransformed.PPM<-hist(DATA$PECK_RATE, na.rm=T, breaks = 12)
+Transformed.PPM<-hist(DATA$LPek, breaks = 12)
 ```
@@
 ### Summarize
-MeanPeck<-summarySE(Peck, "PPM", c("G.Env", "Sentinel"))
+MeanPeck<-summarySE(DATA, "PECK_RATE", c("GENERALIZED_ENVIRONMENT", "SENTINEL_PRESENCE"), na.rm=T)
 MeanPeck
@@
 ```{r Peck Plots, echo = F}
-PeckPlot<-ggplot(MeanPeck, aes(x = G.Env, y = PPM, colour = Sentinel))+
+PeckPlot<-ggplot(MeanPeck, aes(x = GENERALIZED_ENVIRONMENT, y = PECK_RATE, colour = SENTINEL_PRESENCE))+
   geom_point(position = position_dodge(width = 0.9)
              , size = 3)+
-  geom_text(aes(label = signif(PPM))
+  geom_text(aes(label = signif(PECK_RATE))
             , size = 3
@@
             , vjust = 0.1)+
-  geom_errorbar(aes(ymin=(PPM-se), ymax=(PPM+se))
+  geom_errorbar(aes(ymin=(PECK_RATE-se), ymax=(PECK_RATE+se))
                 , width = 0.2
@@
 
-> It appears that there are fewer pecks per minute in Green areas, regardless of the presence of a sentinel.
+> It appears that there more pecks per minute in commercial areas in the presence of a sentinel.
 
@@
 
-```{r Simple peck, echo = F}
-LMM.Peck<-lmer(LPek~Sentinel*G.Env+Baited+(1|ID), data = Peck)
+```{r Simple peck, include= F}
+DATA.PECK<-DATA[complete.cases(DATA),]
+DATA.PECK<-DATA.PECK[DATA.PECK$BEHAVIOR_PROPORTION!=0,]
+LMM.Peck<-lmer(LPek~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+(1|ID), data = DATA.PECK)
 plot(LMM.Peck)
 ```
-
-> Uhh... not good. Need to figure out why.Let's try the robust version of the linear mixed model.
-
-## Robust LMM
-
-```{r Robust Peck, echo = F}
-RLMM.Peck<-rlmer(LPek~Sentinel*G.Env+Baited+(1|ID), data = Peck)
-plot(RLMM.Peck)
-```
-
-> First graph is identical, others seem ok.
 
@@
 
-> All not significant, including random effect. Presence of a sentinel is marginally insignificant (p.val = 0.05419). Let's try the robust model.
+> Generalized environment and bait presence are significant. Weirdly, it seems to also prefer the model without the random effect.
 
-```{r Robust Peck Summary, echo = F}
-sjPlot::tab_model(RLMM.Peck
-                  , show.re.var = T
-                  , title = "Effects of generalized environment, bait, and the presence of a sentinel on the peck rate of foraging crows"
-                  , dv.labels = " Effects on log of pecks per minute")
+## Plot time
+
+### Generalized Environment
+
+```{r G.Env effect on Peck Rate, echo = F}
+EnvPlot.PECK<- DATA.PECK %>%
+  summarySE(measurevar = "PECK_RATE"
+            , groupvars = "GENERALIZED_ENVIRONMENT") %>%
+  ggplot(aes(x = GENERALIZED_ENVIRONMENT, y = PECK_RATE))+
+  geom_point(position = position_dodge(width = 0.9)
+             , size = 3) +
+  geom_text(aes(label = signif(PECK_RATE))
+            , size = 3
+            , position=position_dodge2(width=0.9, preserve='single')
+            , hjust = -0.5
+            , vjust = 0.1)+
+  geom_errorbar(aes(ymin=(PECK_RATE-se), ymax=(PECK_RATE+se))
+                , width = 0.2
+                , position = position_dodge(width=.9))+
+  theme_bw()+
+  theme(text=element_text(size = 14))+
+  xlab("Generalized Environment")+
+  ylab("Peck Rate in Pecks per minute")+
+  theme(legend.position = 'bottom'
+        )
+EnvPlot.PECK
 ```
 
-> Interesting!
-Looks like there are no effects on peck rate.
-Let's plot it.
-The difference between the robust LMM and the simpler model is likely due to how the p-values are estimated, and how the robust lmm decreases the chance of committing Type I error.
+>Very cool! Peck rate is increased in commercial areas.
 
-```{r Robust Peck Effect size}
-sjPlot::plot_model(RLMM.Peck
-                   , show.values=T
-                   , show.p=T
-                   , value.offset = 0.25
-                   , value.size = 4
-                   , wrap.title = 70
-                   , title = "Effects of generalized environment, bait, and the presence of a sentinel on the peck rate of foraging crows")
+### Bait
+
+```{r Bait effect on Peck Rate, echo = F}
+BAITPlot.PECK<- DATA.PECK %>%
+  summarySE(measurevar = "PECK_RATE"
+            , groupvars = "BAIT_PRESENCE") %>%
+  ggplot(aes(x = BAIT_PRESENCE, y = PECK_RATE))+
+  geom_point(position = position_dodge(width = 0.9)
+             , size = 3) +
+  geom_text(aes(label = signif(PECK_RATE))
+            , size = 3
+            , position=position_dodge2(width=0.9, preserve='single')
+            , hjust = -0.5
+            , vjust = 0.1)+
+  geom_errorbar(aes(ymin=(PECK_RATE-se), ymax=(PECK_RATE+se))
+                , width = 0.2
+                , position = position_dodge(width=.9))+
+  theme_bw()+
+  theme(text=element_text(size = 14))+
+  xlab("Bait Presence")+
+  ylab("Peck Rate in Pecks per minute")+
+  theme(legend.position = 'bottom'
+        )
+BAITPlot.PECK
 ```
 
-KMG: Great job!
-I agree with your methods and conclusions.
-Looks like you are pretty much done with your analysis!
-The hard part now will be interpreting the results!
-For results, my suggestion is to stick with plots of effect size and means w/ SE.
-Make sure your stats methods parallels your results section, and in the resuts section present everything in the same way/order.
-It might seem redundant, but it's actually a lot harder to understand if things are in all directions.
-I think I commented on this above.
-I have tried to put all my comments with a KMG: in front.
-I might have forgotten a few.
-You can check the two most recent histories to see what changes I made.
-
-# Disturbances
-To determine whether there were more or fewer disturbances in different areas, I decided to extract the number of disturbances from across my videos, then compare across the same factors.
-
-```{r Disturbance Files, include = F}
-Distub<-read.csv("Disturbance.csv", stringsAsFactors = T)
-```
-
-```{r Disturbance Summary, echo=F}
-summary(Disturb)
+>Seems like peck rate increases significantly in the presence of bait.

### 2023-05-12 Data & Scripts/MSC R Script.Rmd
March 12th
New version of the R script + PCAs. Includes pathway analyses.

@@
 Therefore, the baited crows are expected to move less than non-baited crows.
->I removed the observations in agricultural settings because they had no sentinels present.
+> #KMG: I can imagine a reviewer would want to know hwo the baiting was standardized - something to keep in mind when you are writing.
+I removed the observations in agricultural settings because they had no sentinels present.
 This could be talked about in the discussion: agricultural fields may lack perches for individuals to act as sentinels.
 Vineyards may have some structures, but the visibility is limited when the vines are laden with fruit.\
->I removed observations past Sept 29 because I was told behaviors can change wildly in the winter.
+> #KMG: This is something that can go in the methods.
+You had three sites, but one was removed due to lack of sentinals, and then you can have a bit in the discussion.
+I removed observations past Sept 29 because I was told behaviors can change wildly in the winter.
 The number of crows spotted decreased past that date, suggesting the migrants have left or were leaving.

### 2023-05-27 Data & Scripts/MSC R Script.Rmd
OE3C Presentation
Amended files for the OE3C Presentation

--- 2023-05-12 Data & Scripts/MSC R Script.Rmd
@@
 Crow<-read.csv("Crow.csv", stringsAsFactors = T)
+#Removing outliers
+Crow<-Crow[which(Crow$Duration>0.01),]
+
 #Some management #
@@
 DurDot<-ggplot(data = MeanDur
-               , aes(x = G.Env
+               , aes(x = Behavior
                      , y = Duration
@@
                 , position = position_dodge(width=0.9))+
-  theme_bw() +
+  theme_classic() +
   ylab("Mean Bout Duration (s)") +
   scale_colour_manual(values = cbPalette
-                      , labels = c("Foraging", "Alert", "Moving")) +
+                      , guide="none") +
   scale_shape_manual(values = c(16,17)
@@
                                   , "Sentinel Present")) +
+  scale_x_discrete(labels = c("Foraging", "Alert", "Moving"))+
   theme(axis.title.x = element_blank()
@@
         , legend.box="vertical"
-        , legend.margin=margin())
+        , legend.margin=margin()) +
+  facet_grid(~G.Env)
 
@@
 
-### PCAs
-```{r Bout Duration PCA packages, include=F}
-library("FactoMineR")
-library("factoextra")
-```
-```{r Bout Duration PCA Book keeping, include=F}
-PCA.Crow.Data<-Crow[,c(4,6,8,10,12:13)]
-head(PCA.Crow.Data)
-```
-```{r Bout Duration PCA, echo=F}
-PCA.Crow<-PCA(PCA.Crow.Data[,-5], ncp=5)
-```
-
->I'm new to PCAs, so bear with me please.
+<!-- ### PCAs -->
+<!-- ```{r Bout Duration PCA packages, include=F} -->
+<!-- library("FactoMineR") -->
+<!-- library("factoextra") -->
+<!-- ``` -->
+<!-- ```{r Bout Duration PCA Book keeping, include=F} -->
+<!-- PCA.Crow.Data<-Crow[,c(4,6,8,10,12:13)] -->
+<!-- head(PCA.Crow.Data) -->
+<!-- ``` -->
+<!-- ```{r Bout Duration PCA, echo=F} -->
+<!-- PCA.Crow<-PCA(PCA.Crow.Data[,-5], ncp=5) -->
+<!-- ``` -->
+
+<!-- >I'm new to PCAs, so bear with me please.
 From what I see, presence of bait and sentinel, and the number of crows all contibute to PC1, while Generalized environment seems to be contributing to PC2.
 The behavior appears to contribute to PC2, but not a lot.
-
-#### Eigenvalues
-
->Let's have a look at the eigenvalues.
-
-``` {r Bout Duration PCA eigenvalues, echo = F}
-Crow.eig.val<-get_eigenvalue(PCA.Crow)
-Crow.eig.val
-```
-
->Pretty small values. The second column shows how much contribution each dimension has to explain the total variance in the data. Let's quickly plot that.
-
-#### Scree Plot
-
-```{r Bout Duration PCA Scree Plot, echo = F}
-fviz_eig(PCA.Crow, addlabels=T, ylim = c(0,40))
-```
-
->Pretty cool. The first 3 PCs explain ~77% of the variance. Adding the 4th PC explains ~93% of the variation.
-
-#### Contribution of variables to Principal Components
-
-```{r Bout Duration PCA var results, include = F}
-Crow.var<-get_pca_var(PCA.Crow)
-```
-```{r Bout Duration PCA Contribution Dim 1, echo=F}
-fviz_contrib(PCA.Crow, choice="var", axes = 1, top = 20)
-```
-
->The red dotted line represents the expected average contribution, assuming the contribution of the variables were uniform.
+-->
+
+<!-- #### Eigenvalues -->
+
+<!-- >Let's have a look at the eigenvalues. -->
+
+<!-- ``` {r Bout Duration PCA eigenvalues, echo = F} -->
+<!-- Crow.eig.val<-get_eigenvalue(PCA.Crow) -->
+<!-- Crow.eig.val -->
+<!-- ``` -->
+
+<!-- >Pretty small values. The second column shows how much contribution each dimension has to explain the total variance in the data. Let's quickly plot that. -->
+
+<!-- #### Scree Plot -->
+
+<!-- ```{r Bout Duration PCA Scree Plot, echo = F} -->
+<!-- fviz_eig(PCA.Crow, addlabels=T, ylim = c(0,40)) -->
+<!-- ``` -->
+
+<!-- >Pretty cool. The first 3 PCs explain ~77% of the variance. Adding the 4th PC explains ~93% of the variation. -->
+
+<!-- #### Contribution of variables to Principal Components -->
+
+<!-- ```{r Bout Duration PCA var results, include = F} -->
+<!-- Crow.var<-get_pca_var(PCA.Crow) -->
+<!-- ``` -->
+<!-- ```{r Bout Duration PCA Contribution Dim 1, echo=F} -->
+<!-- fviz_contrib(PCA.Crow, choice="var", axes = 1, top = 20) -->
+<!-- ``` -->
+
+<!-- >The red dotted line represents the expected average contribution, assuming the contribution of the variables were uniform.
 In other words, the line is equal to 1/the number of variables; in this case 1/5 or 20%.
-
-> From this plot, I can see that, as I saw previously in the correlation circle, the number of crows, the presence of a sentinel and the presence of bait contribute the most to PC1 (Dimension 1)
-
-```{r Bout Duration PCA Contribution Dim 2, echo=F}
-fviz_contrib(PCA.Crow, choice="var", axes = 2, top = 20)
-```
-
-> The generalized environment contributes the most to PC2.
-
-```{r Bout Duration PCA Contribution Dim 3, echo=F}
-fviz_contrib(PCA.Crow, choice="var", axes = 3, top = 20)
-```
-
->The coded behavior contributes most to PC3.
-
-```{r Bout Duration PCA Contribution Dim 4, echo=F}
-fviz_contrib(PCA.Crow, choice="var", axes = 4, top = 20)
-```
-
->OK!
+-->
+
+<!-- > From this plot, I can see that, as I saw previously in the correlation circle, the number of crows, the presence of a sentinel and the presence of bait contribute the most to PC1 (Dimension 1) -->
+
+<!-- ```{r Bout Duration PCA Contribution Dim 2, echo=F} -->
+<!-- fviz_contrib(PCA.Crow, choice="var", axes = 2, top = 20) -->
+<!-- ``` -->
+
+<!-- > The generalized environment contributes the most to PC2. -->
+
+<!-- ```{r Bout Duration PCA Contribution Dim 3, echo=F} -->
+<!-- fviz_contrib(PCA.Crow, choice="var", axes = 3, top = 20) -->
+<!-- ``` -->
+
+<!-- >The coded behavior contributes most to PC3. -->
+
+<!-- ```{r Bout Duration PCA Contribution Dim 4, echo=F} -->
+<!-- fviz_contrib(PCA.Crow, choice="var", axes = 4, top = 20) -->
+<!-- ``` -->
+
+<!-- >OK!
 Now we're getting to the interesting bits.
@@
 This makes me curious to see if there is an interaction between the two factors.
-
-```{r Bout Duration PCA Contribution Dim 5, echo=F}
-fviz_contrib(PCA.Crow, choice="var", axes = 5, top = 20)
-```
-
->>The presence of a sentinel and the size of the group contribute most to PC 5.
+-->
+
+<!-- ```{r Bout Duration PCA Contribution Dim 5, echo=F} -->
+<!-- fviz_contrib(PCA.Crow, choice="var", axes = 5, top = 20) -->
+<!-- ``` -->
+
+<!-- >>The presence of a sentinel and the size of the group contribute most to PC 5.
 I believe this would imply that the interaction between the two could be interesting to look at.
 Considering these two variables are also major contributors to PC1, this makes the inclusion of the interaction in the model a must.
-
->From this, I believe I should look at the interaction between NB_Crows, Presence of a sentinel, and the presence of bait.
-
-#### Plotting with Behaviors
-
->I'd like to plot the dimensions using the behaviors of the bouts.
-
-```{r Bout Duration PCA Plot with Behavior PC1-2, echo=F}
-fviz_pca_ind(PCA.Crow
-             , geom.ind="point"
-             , col.ind=PCA.Crow.Data$Behavior
-             , palette = cbPalette
-             , addEllipses = T
-             , legend.title = "Behaviors"
-             )
-```
-
->Messy. The mean points of the behaviors are relatively centered.
-
-```{r Bout Duration PCA Plot with Behavior PC2-3, echo=F}
-fviz_pca_ind(PCA.Crow
-             , axes = c(2,3)
-             , geom.ind="point"
-             , col.ind=PCA.Crow.Data$Behavior
-             , palette = cbPalette
-             , addEllipses = T
-             , legend.title = "Behaviors"
-             )
-```
-
->Very clear separation by behavior type, with 'Head Up' shifted upward, 'Head Down' shifted downward, and 'Moving' in the middle.
-
-```{r Bout Duration PCA Plot with Behavior PC1-3, echo=F}
-fviz_pca_ind(PCA.Crow
-             , axes = c(1,3)
-             , geom.ind="point"
-             , col.ind=PCA.Crow.Data$Behavior
-             , palette = cbPalette
-             , addEllipses = T
-             , legend.title = "Behaviors"
-             )
-```
-
->Good separation, but the concentration ellipses have increased in size. Curious.
-
-#### Biplot
-
->Finally, let's look at the biplot. From what I understand, it is essentially the correlation circle overlaid onto the data.
-
-```{r Bout Duration PCA Biplot PC1-2, echo=F}
-fviz_pca_biplot(PCA.Crow
-                ,geom.ind = "point"
-                , repel = T
-                , col.ind=PCA.Crow.Data$Behavior
-                , addEllipses = T
-                , label = "var"
-                , palette=cbPalette
-                , col.var="black"
-                , legend.title = "Behavior")
-```
-```{r Bout Duration PCA Biplot PC2-3, echo=F}
-fviz_pca_biplot(PCA.Crow
-                , axes = c(2,3)
-                , geom.ind = "point"
-                , repel = T
-                , col.ind=PCA.Crow.Data$Behavior
-                , addEllipses = T
-                , label = "var"
-                , palette=cbPalette
-                , col.var="black"
-                , legend.title = "Behavior")
-```
-```{r Bout Duration PCA Biplot PC1-3, echo=F}
-fviz_pca_biplot(PCA.Crow
-                , axes = c(1,3)
-                , geom.ind = "point"
-                , repel = T
-                , col.ind=PCA.Crow.Data$Behavior
-                , addEllipses = T
-                , label = "var"
-                , palette=cbPalette
-                , col.var="black"
-                , legend.title = "Behavior")
-```
-
->Nice! I think I've done the PCA analysis to the best of my abilities, but I am very much open to comments and feedback!
-
->From what I can gather, I should model the data as follows:/
->LDur~Behavior+NB_Crows*Bait*Sentinel+G.Env+(1|ID)
+-->
+
+<!-- >From this, I believe I should look at the interaction between NB_Crows, Presence of a sentinel, and the presence of bait. -->
+
+<!-- #### Plotting with Behaviors -->
+
+<!-- >I'd like to plot the dimensions using the behaviors of the bouts. -->
+
+<!-- ```{r Bout Duration PCA Plot with Behavior PC1-2, echo=F} -->
+<!-- fviz_pca_ind(PCA.Crow -->
+<!--              , geom.ind="point" -->
+<!--              , col.ind=PCA.Crow.Data$Behavior -->
+<!--              , palette = cbPalette -->
+<!--              , addEllipses = T -->
+<!--              , legend.title = "Behaviors" -->
+<!--              ) -->
+<!-- ``` -->
+
+<!-- >Messy. The mean points of the behaviors are relatively centered. -->
+
+<!-- ```{r Bout Duration PCA Plot with Behavior PC2-3, echo=F} -->
+<!-- fviz_pca_ind(PCA.Crow -->
+<!--              , axes = c(2,3) -->
+<!--              , geom.ind="point" -->
+<!--              , col.ind=PCA.Crow.Data$Behavior -->
+<!--              , palette = cbPalette -->
+<!--              , addEllipses = T -->
+<!--              , legend.title = "Behaviors" -->
+<!--              ) -->
+<!-- ``` -->
+
+<!-- >Very clear separation by behavior type, with 'Head Up' shifted upward, 'Head Down' shifted downward, and 'Moving' in the middle. -->
+
+<!-- ```{r Bout Duration PCA Plot with Behavior PC1-3, echo=F} -->
+<!-- fviz_pca_ind(PCA.Crow -->
+<!--              , axes = c(1,3) -->
+<!--              , geom.ind="point" -->
+<!--              , col.ind=PCA.Crow.Data$Behavior -->
+<!--              , palette = cbPalette -->
+<!--              , addEllipses = T -->
+<!--              , legend.title = "Behaviors" -->
+<!--              ) -->
+<!-- ``` -->
+
+<!-- >Good separation, but the concentration ellipses have increased in size. Curious. -->
+
+<!-- #### Biplot -->
+
+<!-- >Finally, let's look at the biplot. From what I understand, it is essentially the correlation circle overlaid onto the data. -->
+
+<!-- ```{r Bout Duration PCA Biplot PC1-2, echo=F} -->
+<!-- fviz_pca_biplot(PCA.Crow -->
+<!--                 ,geom.ind = "point" -->
+<!--                 , repel = T -->
+<!--                 , col.ind=PCA.Crow.Data$Behavior -->
+<!--                 , addEllipses = T -->
+<!--                 , label = "var" -->
+<!--                 , palette=cbPalette -->
+<!--                 , col.var="black" -->
+<!--                 , legend.title = "Behavior") -->
+<!-- ``` -->
+<!-- ```{r Bout Duration PCA Biplot PC2-3, echo=F} -->
+<!-- fviz_pca_biplot(PCA.Crow -->
+<!--                 , axes = c(2,3) -->
+<!--                 , geom.ind = "point" -->
+<!--                 , repel = T -->
+<!--                 , col.ind=PCA.Crow.Data$Behavior -->
+<!--                 , addEllipses = T -->
+<!--                 , label = "var" -->
+<!--                 , palette=cbPalette -->
+<!--                 , col.var="black" -->
+<!--                 , legend.title = "Behavior") -->
+<!-- ``` -->
+<!-- ```{r Bout Duration PCA Biplot PC1-3, echo=F} -->
+<!-- fviz_pca_biplot(PCA.Crow -->
+<!--                 , axes = c(1,3) -->
+<!--                 , geom.ind = "point" -->
+<!--                 , repel = T -->
+<!--                 , col.ind=PCA.Crow.Data$Behavior -->
+<!--                 , addEllipses = T -->
+<!--                 , label = "var" -->
+<!--                 , palette=cbPalette -->
+<!--                 , col.var="black" -->
+<!--                 , legend.title = "Behavior") -->
+<!-- ``` -->
+
+<!-- >Nice! I think I've done the PCA analysis to the best of my abilities, but I am very much open to comments and feedback! -->
+
+<!-- >From what I can gather, I should model the data as follows:/ -->
+<!-- >LDur~Behavior+NB_Crows*Bait*Sentinel+G.Env+(1|ID) -->
 
@@
 ```{r Duration RLMM, echo=F}
-RLMM.Dur.I<-rlmer(LDur~Behavior+Sentinel*CODED_NB_CROWS*Baited+G.Env+ (1|ID), data = Crow)
+RLMM.Dur.I<-rlmer(LDur~Behavior+Sentinel*G.Env+Sentinel*CODED_NB_CROWS+Baited+Sentinel+G.Env+ (1|ID), data = Crow)
 plot(RLMM.Dur.I)
@@
                   , title = "Model Results with Interactions"
-                  , dv.labels = " Effects of on duration of behaviors")
+                  , dv.labels = " Effects of on duration of behaviors"
+                  , file="Tab_Dur_Model.html")
 
@@
 
+```{r Effect Sizes}
+sjPlot::plot_model(RLMM.Dur.I
+                   , show.values=T
+                   , show.p=T
+                   , value.offset = 0.4
+                   , value.size = 3.5
+                   , wrap.title = 48
+                   , title = "Effects of environment, presence of a sentinel, number of crows and bait on durations of behaviors")# Looks much better
+```
 >Let's try the good old regular lmer and see if that changes anything.
@@
 ```{r Bait effect on HD bout, echo = F}
-BaitPlot<- HD.Dur %>%
+BaitPlot<- Crow[which(Crow$Behavior=="HD"),] %>%
   summarySE(measurevar = "Duration"
@@
              , size = 3) +
-  geom_text(aes(label = signif(Duration))
-            , size = 3
-            , position=position_dodge2(width=0.9, preserve='single')
-            , hjust = -0.5
-            , vjust = 0.1)+
   geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
-                , width = 0.2
+                , width = 0.1
                 , position = position_dodge(width=.9))+
-  theme_bw()+
+  theme_classic()+
   theme(text=element_text(size = 14))+
-  xlab("Presence of Bait")+
-  ylab("Mean Head Down Bout Duration (s)")+
+  xlab("")+
+  ylab("Mean Bout Duration (s)")+
   scale_x_discrete(labels=c("Not Baited", "Baited"))+
-  theme(legend.position = 'bottom'
+  theme(legend.position = 'none'
         )
@@
   geom_bar(stat = 'identity', position = 'stack')+
-  geom_text(aes(label = paste0(signif(BEHAVIOR_PROPORTION*100), "%")), position = position_stack(vjust = 0.5)
-            , size = 3)+
+  geom_text(aes(label = paste0(formattable::digits(BEHAVIOR_PROPORTION*100, dig=2), "%"))
+            , position = position_stack(vjust = 0.5)
+            , size = 3) +
   scale_y_continuous(labels = scales::percent) +
-  theme_bw()+
+  theme_classic()+
   theme(text=element_text(size = 10)
@@
 ```{r Peck Per Minute Histograms Individual, include = F}
-Untransformed.PPM<-hist(DATA$PECK_RATE, na.rm=T, breaks = 12)
-Transformed.PPM<-hist(DATA$LPek, breaks = 12)
+Untransformed.PPM<-hist(DATA$PECK_RATE, na.rm=T, breaks = 25)
+Transformed.PPM<-hist(DATA$LPek, breaks = 25)
 ```
@@
 ```{r Peck Plots, echo = F}
-PeckPlot<-ggplot(MeanPeck, aes(x = GENERALIZED_ENVIRONMENT, y = PECK_RATE, colour = SENTINEL_PRESENCE))+
+PeckPlot<-ggplot(MeanPeck, aes(x = SENTINEL_PRESENCE, y = PECK_RATE, color=SENTINEL_PRESENCE))+
   geom_point(position = position_dodge(width = 0.9)
              , size = 3)+
-  geom_text(aes(label = signif(PECK_RATE))
-            , size = 3
-            , position=position_dodge2(width=1, preserve='single')
-            , hjust = -0.5
-            , vjust = 0.1)+
+  # geom_text(aes(label = signif(PECK_RATE))
+  #           , size = 3
+  #           , position=position_dodge2(width=1, preserve='single')
+  #           , hjust = -0.5
+  #           , vjust = 0.1)+
   geom_errorbar(aes(ymin=(PECK_RATE-se), ymax=(PECK_RATE+se))
-                , width = 0.2
+                , width = 0.1
                 , position = position_dodge(.9))+
-  theme_bw()+
+  theme_classic()+
   labs(colour = "Presence of a sentinel")+
@@
                        , labels=c("Sentinel Absent", "Sentinel Present")) +
-  scale_x_discrete(labels=c("Commercial", "Green Area"))+
-  theme(legend.position = 'bottom'
+  scale_x_discrete(labels=c("Sentinel Absent", "Sentinel Present"))+
+  theme(legend.position = ''
         , legend.title = element_blank()
         , axis.title.x = element_blank()
-        , text=element_text(size = 12))
+        , text=element_text(size = 12)) +
+  facet_grid(~GENERALIZED_ENVIRONMENT)
 PeckPlot
@@
 LMM.Peck<-lmer(LPek~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+(1|ID), data = DATA.PECK)
-plot(LMM.Peck)
+plot(residuals(LMM.Peck))
 ```
@@
              , size = 3) +
-  geom_text(aes(label = signif(PECK_RATE))
-            , size = 3
-            , position=position_dodge2(width=0.9, preserve='single')
-            , hjust = -0.5
-            , vjust = 0.1)+
+  # geom_text(aes(label = signif(PECK_RATE))
+  #           , size = 3
+  #           , position=position_dodge2(width=0.9, preserve='single')
+  #           , hjust = -0.5
+  #           , vjust = 0.1)+
   geom_errorbar(aes(ymin=(PECK_RATE-se), ymax=(PECK_RATE+se))
-                , width = 0.2
+                , width = 0.1
                 , position = position_dodge(width=.9))+
-  theme_bw()+
+  theme_classic()+
   theme(text=element_text(size = 14))+

### 2023-06-09 Data & Scripts/BOUT.Rmd
June 9th
Meeting with KMG

---
title: "BOUT"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

# BOUT

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

## Preamble

This script will deal with analysing the relationship of the duration of bouts of behavior - head down, head up, and moving - with the presence of a sentinel, the presence of bait, the foraging environment and the group size.

As random effects, I've chosen the individual ID, the (julian) date, and the (decimal) time.

This analysis will answer the following question:\
Do individual vary the duration of behavioral bouts in response to their perception of risk and sentinel coverage in their foraging environment.

In other words, if the proportion of time allocated to each behavior does not vary, then does the individual alter the frequency, and therefore the duration of bouts of each behavior in response to, for example, an environment with shorter lines of sight for the sentinel.

To do so, I extracted the bouts of each individual using the following string in the 'advanced event filtering' tool in BORIS v.8.20.3:

"No focal subject\|[Sentinel State]" & "Individual X\|[Behavior]"

I then compiled a dataset and named it "BOUT"

```{r Open Data}
BOUT<-read.csv("BOUT.csv", stringsAsFactors = T)
names(BOUT)[14]<-"DURATION"
str(BOUT)
```

I will also remove duration values smaller than 0.01, since it is likely an artifact from coding in BORIS. Sometimes things get cut weirdly, resulting in impossibly small values.

```{r Remove Small Values}
BOUT<-BOUT[which(BOUT$DURATION>0.01),]
```

This removed 21 points, leaving me with 5070 bouts.

```{r Bout Summary}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
BOUT.MEAN <- summarySE(data = BOUT
                      , measurevar = "DURATION"
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "BEHAVIOR"
                                      , "SENTINEL_PRESENCE"
                                      )
                      )
BOUT.MEAN
```

This has given me the mean duration of bouts, grouped into each combination of "main" factors of interest.

Onto the analysis

## To transform or not

That is the question.

Due to the nature of the data, the best type of transformation is a log transformation.

```{r Duration Histograms Individual, include = F}
Untransformed.Dur<-hist(BOUT$DURATION, breaks = 25)
Transformed.Dur<-hist(log(BOUT$DURATION), breaks = 25)
```

```{r Duration Histograms Combined, echo=F}
op <- par(mfrow = c(1,2))
plot(Untransformed.Dur, main="Untransformed", ylab="Frequency", xlab="Duration (s)")
plot(Transformed.Dur, main="Transformed", ylab="", xlab="Log Duration (s)")
par(op)
```

We can clearly see that the untransformed data is right-skewed, and always positive. This is perfect for log-transformations. We see in the histogram of the transformed data, the data is now roughly normally distributed.

Therefore, for the statistical analyses, we will use **LOG-Transformed** durations.

```{r LOG BOUT}
BOUT$LDURATION<-log(BOUT$DURATION)
```

Here is the histogram of the log-transformed data separated by behavior.

```{r Crow Histogram, echo=F}
HIST<-ggplot(data = BOUT) +
  geom_histogram(aes(fill = BEHAVIOR, x = LDURATION)
                 , alpha = 1
                 , bins = 25)+
  scale_fill_manual(values=cbPalette
                    , labels = c("Foraging", "Alert", "Moving"))+
  xlab("Log Duration (s)")+
  ylab("Count")
HIST
```

There is a gap in the distribution. That is unfortunate.

## The Dot Plot

Here's where things get beautiful.

```{r BOUT Dot Plot}
BOUT.DOT<-ggplot(data = BOUT.MEAN
               , aes(x = BEHAVIOR
                     , y = DURATION
                     , color = BEHAVIOR
                     , shape = SENTINEL_PRESENCE))+
  geom_point(position = position_dodge2(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.2
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none") +
  scale_shape_manual(values = c(16,17)
                     , labels = c("Sentinel Absent"
                                  , "Sentinel Present")
                     , name = "") +
  scale_x_discrete(labels = c("Foraging", "Alert", "Moving"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  facet_grid(~GENERALIZED_ENVIRONMENT)

BOUT.DOT
```

> The error bars represent the standard error, while the symbols are the mean duration of bouts.

Beautiful! It looks like the duration of bouts of alertness and foraging increase increases when sentinels are present in commercial areas. In green areas, this effect is reversed, with bouts of alertness and foraging having decreased duration in the presence of a sentinel. Meanwhile, bouts of movement seem affected by neither the environment nor the presence of a sentinel.

Let's fit the models.

## The Models

Let's fit a model. We are testing principally for the effects of the presence of a sentinel and the generalized environment. However, group size and the presence of bait could also affect the duration of bouts.

I originally had temperature, time and data as random effects, but the model was singular. Running rePCA showed that the random effects explained the same variance.

The first model fitted will be a simple model with no interactions.

```{r BOUT MODEL 1}
BOUT.MOD1<-lmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+(1|ID), data = BOUT)

sjPlot::tab_model(BOUT.MOD1
                  , show.re.var = T
                  , title = "Bout Model 1 Output"
                  , dv.labels = " Effects of on duration of bouts of behaviors")
```

The simple model without interactions shows behavior and the generalized environment having significant effects on the duration of bouts.

Now, we would expect to see bouts of different behaviors being different. It takes longer to manipulate food than it does to look out for sources of threat.\
However, the presence of a sentinel and the environment could modify this effect. If a sentinel is present, then I would expect the duration of bouts of alertness to decrease, while a riskier environment should cause the duration of these bouts to increase.\
As such, in the next model, I will include an interaction between behavior and the presence of a sentinel, as well as behavior and the generalized environment.

The generalized environment has a significant effect, and this supports, at least in part, my predictions.

However, the presence of a sentinel did not affect the duration of bouts. It may be that the effects of the presence of a sentinel is interactive with the type of environment. I will therefore include an interaction between the type of environment and the presence of a sentinel in the next model.

Interestingly, the presence of bait and the group size did not have an effect. I would have expected the bait, a concentrated patch of food, to have an effect on the duration of behavior. It may be the case that the presence of bait only affects the duration of bouts of foraging. As a result, I will also include an interaction between the presence of bait and the behavior in the subsequent model.

Group size provides antipredator benefits to the entire group. If there are many individuals, then either the individual risk of predation decreases (dilution effect) or there are more foragers capable of detecting threats (many-eyes).\
As such, it is possible that, once again, the group size's effects occur on certain behaviors rather than others. I will include this interaction in the next model.

```{r BOUT MODEL 2}
BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+(1|ID), data = BOUT)

sjPlot::tab_model(BOUT.MOD2
                  , show.re.var = T
                  , title = "Bout Model 2 Output"
                  , dv.labels = " Effects of on duration of bouts of behaviors")
```

In the second model with the added interactions, the behavior, generalized environment, group size and a number of interactions are significant.

The group size becoming significant is puzzling. I don't understand why now it would become significant.

The interaction between the behavior type and the environment are significant. This would imply that individuals change the duration of bouts of behaviors in response to different environments, with the effect differing between behaviors.\

The interaction between the presence of a sentinel and the generalized environment is also significant. As before, it could be worthwhile to also look at the three-way interaction between behavior, the presence of a sentinel and the generalized environment. This could reveal different effects caused by the presence of a sentinel and the environment on different behaviors.

The interaction between behavior and group size is significant. Again, this is unexpected. I will include a three way interaction between behavior, group size and the presence of a sentinel in the next model, if the interactive effects of group size and behavior are in some way affected by sentinel behavior.

Lastly, the interaction between behavior "Moving" and the presence of bait is significant. This is curious, especially since the effect size appears to be positive when bait is present. My interpretation of this is moving bout durations increase when bait is present.

Let's run the final and most complex model.

```{r BOUT MOD 3}
BOUT.MOD3<-lmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+BEHAVIOR*GENERALIZED_ENVIRONMENT*SENTINEL_PRESENCE+BEHAVIOR*GROUP_SIZE*SENTINEL_PRESENCE+BEHAVIOR*BAIT_PRESENCE+(1|ID), data = BOUT)

sjPlot::tab_model(BOUT.MOD3
                  , show.re.var = T
                  , title = "Bout Model 3 Output"
                  , dv.labels = " Effects of on duration of bouts of behaviors")
```

Ok, so there are a number of differences. The only significant effects detected are the generalized environment, the interaction between "moving" and green areas, and the interaction between "moving" and bait presence.

This is weird. Let's see what the AICs are.

```{r BOUT AIC}
anova(BOUT.MOD1,BOUT.MOD2,BOUT.MOD3)
```

Nice! The second model has the lowest AIC value and is therefore prefered.

In case there are any violations in the assumptions of the linear mixed model, I will run the robust linear mixed model, using model 2's formula.

```{r RLMM MOD 2}
BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+(1|ID), data = BOUT)

sjPlot::tab_model(BOUT.RMOD2
                  , show.re.var = T
                  , title = "Bout Robust Model 2 Output"
                  , dv.labels = " Effects of on duration of bouts of behaviors")
```

Same results as MOD2, but different p-values.

```{r LMM Assumptions}
plot(BOUT.RMOD2)
```

The assumptions of the LMM do not appear to be violated.

### Model Conclusions

Using model 2, here are the conclusions I draw:

-   Behavior, generalized environment and group size have significant effects. The former is to be expected, while the latter two are interesting, but not very informative (alone).
-   The interaction between behavior and generalized environment are significant. The effect size appears negative for both "head up" and "moving" behavior in green areas.
-   There is a significant interaction of sentinel presence and generalized environment.
-   There is a significant interaction of behavior and group size for head up and moving.
-   There is a significant interaction of "moving" and bait presence.

To better visualize these results, I will plot them.

#### Behavior X Generalized Environment

```{r BxG.ENV BOUT Dot Plot}
BxG.ENV.DOT<-BOUT %>%
  summarySE(measurevar = "DURATION"
            , groupvars = c("BEHAVIOR", "GENERALIZED_ENVIRONMENT")) %>%
    ggplot(aes(x = BEHAVIOR
                     , y = DURATION
                     , color = BEHAVIOR))+
  geom_point(position = position_dodge2(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  scale_x_discrete(labels = c("Foraging", "Alert", "Moving"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  facet_grid(~GENERALIZED_ENVIRONMENT)

BxG.ENV.DOT
```

Very interesting! Duration of bouts of alertness and moving remain largely the same, while the duration of foraging bout in green areas increases dramatically. \

This could be explained by needing more time to locate and handle food in green areas, especially if the food is not on concrete or in a dense patch.

#### Sentinel Presence X Generalized Environment

```{r SPxG.ENV BOUT Dot Plot}
SPxG.ENV.DOT<-BOUT %>%
  summarySE(measurevar = "DURATION"
            , groupvars = c("SENTINEL_PRESENCE", "GENERALIZED_ENVIRONMENT")) %>%
    ggplot(aes(x = SENTINEL_PRESENCE
                     , y = DURATION
                     , color = SENTINEL_PRESENCE))+
  geom_point(position = position_dodge2(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  scale_x_discrete(labels = c("Sentinel Absent", "Sentinel Present"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  facet_grid(~GENERALIZED_ENVIRONMENT)

SPxG.ENV.DOT
```

In the presence of a sentinel, bout durations of all behaviors remain relatively similar across environments. However, when no sentinels are present the bout duration of all behaviors increases dramatically in green areas.

I will need to find a way to explain this result, especially since the duration of behaviors when no sentinel is present is much shorter in commercial areas and much longer in green areas than in the presence of a sentinel, regardless of the environment.

```{r BxGS BOUT Dot Plot}
BxGS.DOT<-BOUT %>%
  summarySE(measurevar = "DURATION"
            , groupvars = c("BEHAVIOR", "GROUP_SIZE")) %>%
    ggplot(aes(x = BEHAVIOR
               , y = DURATION
               , color = BEHAVIOR
               , shape = GROUP_SIZE
               , group = BEHAVIOR))+
  geom_point(position = position_dodge2(width = 0.5)
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge2())+
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  scale_x_discrete(labels = c("Foraging", "Alert", "Moving"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())

BxGS.DOT
```

GGPLOT! The error bars are not over the dots.

#### Behavior x Bait Presence

```{r BxBP BOUT}
BxBP.DOT<-BOUT %>%
  summarySE(measurevar = "DURATION"
            , groupvars = c("BEHAVIOR", "BAIT_PRESENCE")) %>%
    ggplot(aes(x = BEHAVIOR
               , y = DURATION
               , color = BEHAVIOR
               , shape = BAIT_PRESENCE
               , group = BEHAVIOR))+
  geom_point(position = position_dodge2(width = 0.5)
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge2())+
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  scale_x_discrete(labels = c("Foraging", "Alert", "Moving"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())

BxBP.DOT
```

### 2023-06-16 Data & Scripts/MSC R Script.Rmd
Another push
What I had on the desktop

@@
 
+```{r DATA setup, include = F}
+DATA<-read.csv("DATA.csv", stringsAsFactors = T)
+DATA<-DATA[,c(-3,-4,-6)]
+DATA$ID<-as.factor(DATA$ID)
+DATA$LOG_DURATION<-log(DATA$MEAN_BOUT_DURATION)
+DATA$LOG_DURATION[is.infinite(DATA$LOG_DURATION)]<-NA
+DATA$ASIN_PROPORTION<-asin(sqrt(DATA$BEHAVIOR_PROPORTION))
+
+```
+
 ```{r Crow Double checking, echo = F}
@@
 ```{r Duration Histograms Individual, include = F}
-Untransformed.Dur<-hist(Crow$Duration, breaks = 25)
-Transformed.Dur<-hist(Crow$LDur, breaks = 25)
+Untransformed.Dur<-hist(DATA$MEAN_BOUT_DURATION, breaks = 25)
+Transformed.Dur<-hist(DATA$LOG_DURATION, breaks = 25)
 ```
@@
 ```{r Crow Histogram, echo=F}
-Hist<-ggplot(data = Crow) +
-  geom_histogram(aes(fill = Behavior, x = LDur)
+Hist<-ggplot(data = DATA) +
+  geom_histogram(aes(fill = BEHAVIOR, x = LOG_DURATION)
                  , alpha = 1
@@
 
-### Boxplot
-
-```{r Crow boxplot, echo = F}
-Box<-ggplot(data = Crow)+
-  geom_boxplot_pattern(aes(x = G.Env, y = LDur, fill = Behavior, pattern = Sentinel)
-                       , pattern_angle = 45
-                       , pattern_size = 0.05
-                       , pattern_spacing = 0.02
-                       , color = "black")+
-  theme(legend.position = "bottom")+
-  scale_fill_manual(values = cbPalette
-                    , labels = c("Foraging", "Alert", "Moving"))+
-  scale_pattern_manual(values = c("crosshatch", "none")
-                       , labels = c("Sentinel Absent", "Sentinel Present"))+
-  guides(pattern = guide_legend(title = "Presence of a sentinel"
-                                , override.aes = list(fill = c("white")))
-         , fill = guide_legend(override.aes = list(pattern = "none")))+
-  xlab("Generalized Environment")+
-  ylab("Log Duration(s)") +
-  theme(axis.title.x = element_blank()
-        , legend.position = "bottom"
-        , legend.box="vertical"
-        , legend.margin=margin())
-Box
-```
-
-### Violin plot
-
-```{r Crow Violin plot, echo = F}
-Violin<-ggplot(data = Crow
-               , aes(x = Behavior, y = LDur, fill = Sentinel))+
-  geom_violin()+
-  theme_bw()+
-  theme(legend.position = "bottom")+
-  scale_fill_manual(values = cbPalette
-                    , labels = c("Sentinel Absent", "Sentinel Present"))+
-  scale_x_discrete(labels=c("Foraging", "Alert", "Moving"))+
-  xlab("Generalized Environment")+
-  ylab("Log Duration(s)")+
-  facet_grid(~G.Env)+
-  stat_summary(fun = "mean"
-               , geom = "crossbar"
-               , position = "dodge"
-               , color = "red")
-Violin
-```
-
->These plots are nice, but it is hard to really see the underlying trends. To best visualise them, I'll next plot a dot plot.
+<!-- ### Boxplot -->
+
+<!-- ```{r Crow boxplot, echo = F} -->
+<!-- Box<-ggplot(data = Crow)+ -->
+<!--   geom_boxplot_pattern(aes(x = G.Env, y = LDur, fill = Behavior, pattern = Sentinel) -->
+<!--                        , pattern_angle = 45 -->
+<!--                        , pattern_size = 0.05 -->
+<!--                        , pattern_spacing = 0.02 -->
+<!--                        , color = "black")+ -->
+<!--   theme(legend.position = "bottom")+ -->
+<!--   scale_fill_manual(values = cbPalette -->
+<!--                     , labels = c("Foraging", "Alert", "Moving"))+ -->
+<!--   scale_pattern_manual(values = c("crosshatch", "none") -->
+<!--                        , labels = c("Sentinel Absent", "Sentinel Present"))+ -->
+<!--   guides(pattern = guide_legend(title = "Presence of a sentinel" -->
+<!--                                 , override.aes = list(fill = c("white"))) -->
+<!--          , fill = guide_legend(override.aes = list(pattern = "none")))+ -->
+<!--   xlab("Generalized Environment")+ -->
+<!--   ylab("Log Duration(s)") + -->
+<!--   theme(axis.title.x = element_blank() -->
+<!--         , legend.position = "bottom" -->
+<!--         , legend.box="vertical" -->
+<!--         , legend.margin=margin()) -->
+<!-- Box -->
+<!-- ``` -->
+
+<!-- ### Violin plot -->
+
+<!-- ```{r Crow Violin plot, echo = F} -->
+<!-- Violin<-ggplot(data = Crow -->
+<!--                , aes(x = Behavior, y = LDur, fill = Sentinel))+ -->
+<!--   geom_violin()+ -->
+<!--   theme_bw()+ -->
+<!--   theme(legend.position = "bottom")+ -->
+<!--   scale_fill_manual(values = cbPalette -->
+<!--                     , labels = c("Sentinel Absent", "Sentinel Present"))+ -->
+<!--   scale_x_discrete(labels=c("Foraging", "Alert", "Moving"))+ -->
+<!--   xlab("Generalized Environment")+ -->
+<!--   ylab("Log Duration(s)")+ -->
+<!--   facet_grid(~G.Env)+ -->
+<!--   stat_summary(fun = "mean" -->
+<!--                , geom = "crossbar" -->
+<!--                , position = "dodge" -->
+<!--                , color = "red") -->
+<!-- Violin -->
+<!-- ``` -->
+
+<!-- >These plots are nice, but it is hard to really see the underlying trends. To best visualise them, I'll next plot a dot plot. -->
 
@@
 
-```{r Base R Plot, include = F}
-source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
-MeanDur <- summarySE(data = Crow, measurevar = "Duration"
-                     , groupvars = c("G.Env", "Behavior", "Sentinel"))
-
-plot(x = c(1:12)
-     , y = MeanDur$Duration
-     , xlim = c(0,12)
-     , ylim = c(1,3)
-     , type = "n"
-     , ylab = "Duration (seconds)", xlab = "Generalized Environment")
-arrows(c(1:12), MeanDur$Duration - MeanDur$se
-       , c(1:12), MeanDur$Duration + MeanDur$se
-       , length = 0.1, angle = 90, code = 3, lwd = 2)
-points(x = c(1:12), y = MeanDur$Duration
-       , pch = c(rep(c(16, 17), 3)), cex = 2
-       , col = c("orange", "orange", "blue", "blue", "green", "green"))
-legend(x = "bottomright", c("Sentinel Absent", "Sentinal Present", "Foraging", "Alert", "Moving")
-       , bty = "o", cex = 1.05, bg = "white"
-       , pch = c(16, 17, 15, 15, 15), col = c("black", "black", "orange", "blue", "green"))
-#KMG: I'm not going to fix the X-axis labels and those kinds of things right now.
+<!-- ```{r Base R Plot, include = F} -->
+<!-- source("./Calculate Summary Table with SE, SD, CI by grouping variables.R") -->
+<!-- MeanDur <- summarySE(data = Crow, measurevar = "Duration" -->
+<!--                      , groupvars = c("G.Env", "Behavior", "Sentinel")) -->
+
+<!-- plot(x = c(1:12) -->
+<!--      , y = MeanDur$Duration -->
+<!--      , xlim = c(0,12) -->
+<!--      , ylim = c(1,3) -->
+<!--      , type = "n" -->
+<!--      , ylab = "Duration (seconds)", xlab = "Generalized Environment") -->
+<!-- arrows(c(1:12), MeanDur$Duration - MeanDur$se -->
+<!--        , c(1:12), MeanDur$Duration + MeanDur$se -->
+<!--        , length = 0.1, angle = 90, code = 3, lwd = 2) -->
+<!-- points(x = c(1:12), y = MeanDur$Duration -->
+<!--        , pch = c(rep(c(16, 17), 3)), cex = 2 -->
+<!--        , col = c("orange", "orange", "blue", "blue", "green", "green")) -->
+<!-- legend(x = "bottomright", c("Sentinel Absent", "Sentinal Present", "Foraging", "Alert", "Moving") -->
+<!--        , bty = "o", cex = 1.05, bg = "white" -->
+<!--        , pch = c(16, 17, 15, 15, 15), col = c("black", "black", "orange", "blue", "green")) -->
+<!-- #KMG: I'm not going to fix the X-axis labels and those kinds of things right now.
 I can do that if you like this plot or things it's informative.
@@
 Legend should be correct.
-```
+-->
+<!-- ``` -->
 
@@
 source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
-MeanDur <- summarySE(data = Crow, measurevar = "Duration"
-                     , groupvars = c("G.Env", "Behavior", "Sentinel"))
+MeanDur <- summarySE(data = DATA, measurevar = "MEAN_BOUT_DURATION"
+                     , groupvars = c("GENERALIZED_ENVIRONMENT", "BEHAVIOR", "SENTINEL_PRESENCE"))
 DurDot<-ggplot(data = MeanDur
-               , aes(x = Behavior
-                     , y = Duration
-                     , color = Behavior
-                     , shape = Sentinel))+
+               , aes(x = BEHAVIOR
+                     , y = MEAN_BOUT_DURATION
+                     , color = BEHAVIOR
+                     , shape = SENTINEL_PRESENCE))+
   geom_point(position = position_dodge2(width = 0.9)
              , size = 3) +
-  geom_errorbar(aes(ymin=(Duration-se)
-                    , ymax=(Duration+se))
+  geom_errorbar(aes(ymin=(MEAN_BOUT_DURATION-se)
+                    , ymax=(MEAN_BOUT_DURATION+se))
                 , width = 0.2
@@
         , legend.margin=margin()) +
-  facet_grid(~G.Env)
+  facet_grid(~GENERALIZED_ENVIRONMENT)
 
@@
 
+```{r DATA Duration RLMM, echo=F}
+RLMM.Dur.DATA<-rlmer(LOG_DURATION~BEHAVIOR+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GROUP_SIZE+BAIT_PRESENCE+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+(1|ID), data = DATA)
+plot(RLMM.Dur.DATA)
+```
+
 > That took pretty long to run.
@@
 ```{r Duration Better Summary, echo = F}
-sjPlot::tab_model(RLMM.Dur.I
+sjPlot::tab_model(RLMM.Dur.DATA
                   , show.re.var = T
                   , title = "Model Results with Interactions"
-                  , dv.labels = " Effects of on duration of behaviors"
-                  , file="Tab_Dur_Model.html")
+                  , dv.labels = " Effects of on duration of behaviors")
 
@@
 
-```{r Proportion setup, include = F}
-DATA<-read.csv("DATA.csv", stringsAsFactors = T)
-DATA$ID<-as.factor(DATA$ID)
-DATA$ASIN_PROPORTION<-asin(sqrt(DATA$BEHAVIOR_PROPORTION))
-```
-
 ```{r Proportion Double Check, echo = F}
@@
 
->Let's run another PCA
-
-```{r DATA PCA SHEET, include=F}
-PCA.DATA.SHEET<-DATA[,c(5,7,10,14,16,18,22,23,24,39)]
-head(PCA.DATA.SHEET)
-str(PCA.DATA.SHEET)
-```
-```{R DATA PCA, echo=F}
-PCA.DATA<-PCA(PCA.DATA.SHEET[,c(-8,-9)])
-```
-
->There are many variables here. I won't bother describing them here.
-
->Since the presence of a sentinel is not "God-Given", we'll run another PCA without that factor.
-
-```{r DATA PCA NO SENTINEL, echo=F}
-PCA.DATA.NOSENT<-PCA(PCA.DATA.SHEET[,c(-5,-8,-9)])
-```
-
->Very similar to the one above.
-
-### Scree Plot
-
-```{r DATA PCA Scree Plot, echo = F}
-get_eigenvalue(PCA.DATA)
-fviz_eig(PCA.DATA, addlabels=T, ylim = c(0,40))
-```
-
->The first PC covers ~40% of the variation. The first 4 PCs cover ~82% of the variation.
-
-```{r DATA PCA Scree Plot NO SENTINEL, echo = F}
-get_eigenvalue(PCA.DATA.NOSENT)
-fviz_eig(PCA.DATA.NOSENT, addlabels=T, ylim = c(0,40))
-```
-
->Excluding sentinel, the first PC covers ~43% of the variation, and the first 4 PCs cover ~86% of the total variation.
-
-### Contribution of variables to Principal Components
-
-```{r PCA DATA var results, include = F}
-DATA.var<-get_pca_var(PCA.DATA)
-DATA.NOSENT.var<-get_pca_var(PCA.DATA.NOSENT)
-```
-```{r PCA DATA Contribution Dim 1, echo=F}
-fviz_contrib(PCA.DATA, choice="var", axes = 1, top = 12.5)
-```
-
->Very weird, Julian Date, Decimal time, group size and generalized environment contribute greatly to PC1.
-
-```{r PCA DATA Contribution Dim 2, echo=F}
-fviz_contrib(PCA.DATA, choice="var", axes = 2, top = 12.5)
-```
-
->Frequency of disturbances, group size (again), temperature and the presence of a sentinel contribute most to PC2.
-
-```{r PCA DATA Contribution Dim 3, echo=F}
-fviz_contrib(PCA.DATA, choice="var", axes = 3, top = 12.5)
-```
-
->Temperature, Bait presence and decimal time contribute most to PC3.
-
-```{r PCA DATA Contribution Dim 4, echo=F}
-fviz_contrib(PCA.DATA, choice="var", axes = 4, top = 12.5)
-```
-
->Sentinel presence, environment and frequency of disturbances contribute most to PC4.
-
->I'm still a little unclear as to how I should interpret these results.
-
-#### No Sentinel
-
-```{r PCA DATA.NOSENT Contribution Dim 1, echo=F}
-fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 1, top = 14.29)
-```
-
->Same 4 factors contribute the most to PC1 in the absence of the sentinel factor.
-
-```{r PCA DATA.NOSENT Contribution Dim 2, echo=F}
-fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 2, top = 14.29)
-```
-
->Same as above, frequency of disturbances, temperature and group size contribute most to PC2 in the absence of the sentinel factor.
-
-```{r PCA DATA.NOSENT Contribution Dim 3, echo=F}
-fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 3, top = 14.29)
-```
-
->Temperature, Bait presence and frequency of disturbances contribute the most to PC3 in the absence of a sentinel.
-
-```{r PCA DATA.NOSENT Contribution Dim 4, echo=F}
-fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 4, top = 14.29)
-```
-
->Environment contributes most to PC4 in the absence of a sentinel.
-
-### BiPlots
-```{r DATA PCA Biplot PC1-2, echo=F}
-fviz_pca_biplot(PCA.DATA
-                ,geom.ind = "point"
-                , repel = T
-                , col.ind=PCA.DATA.SHEET$BEHAVIOR
-                , addEllipses = T
-                , label = "var"
-                , palette=cbPalette
-                , col.var="black"
-                , legend.title = "Behavior")
-```
-
->Messy, can't see where the behaviors lie.
-
-```{r DATA PCA Biplot PC1-4, echo=F}
-fviz_pca_biplot(PCA.DATA
-                , axes = c(1,4)
-                , geom.ind = "point"
-                , repel = T
-                , col.ind=PCA.DATA.SHEET$BEHAVIOR
-                , addEllipses = T
-                , label = "var"
-                , palette=cbPalette
-                , col.var="black"
-                , legend.title = "Behavior")
-```
-
->Still messy.
-
-```{r DATA PCA Biplot PC2-3, echo=F}
-fviz_pca_biplot(PCA.DATA
-                , axes = c(2,3)
-                , geom.ind = "point"
-                , repel = T
-                , col.ind=PCA.DATA.SHEET$BEHAVIOR
-                , addEllipses = T
-                , label = "var"
-                , palette=cbPalette
-                , col.var="black"
-                , legend.title = "Behavior")
-```
-
->Biplots are messy if we don't include behavior.
-
->Same is observed if we run the biplot in the absence of the sentinel factor.
-
->I'm not sure how to interpret these results.
+<!-- >Let's run another PCA -->
+
+<!-- ```{r DATA PCA SHEET, include=F} -->
+<!-- PCA.DATA.SHEET<-DATA[,c(5,7,10,14,16,18,22,23,24,39)] -->
+<!-- head(PCA.DATA.SHEET) -->
+<!-- str(PCA.DATA.SHEET) -->
+<!-- ``` -->
+<!-- ```{R DATA PCA, echo=F} -->
+<!-- PCA.DATA<-PCA(PCA.DATA.SHEET[,c(-8,-9)]) -->
+<!-- ``` -->
+
+<!-- >There are many variables here. I won't bother describing them here. -->
+
+<!-- >Since the presence of a sentinel is not "God-Given", we'll run another PCA without that factor. -->
+
+<!-- ```{r DATA PCA NO SENTINEL, echo=F} -->
+<!-- PCA.DATA.NOSENT<-PCA(PCA.DATA.SHEET[,c(-5,-8,-9)]) -->
+<!-- ``` -->
+
+<!-- >Very similar to the one above. -->
+
+<!-- ### Scree Plot -->
+
+<!-- ```{r DATA PCA Scree Plot, echo = F} -->
+<!-- get_eigenvalue(PCA.DATA) -->
+<!-- fviz_eig(PCA.DATA, addlabels=T, ylim = c(0,40)) -->
+<!-- ``` -->
+
+<!-- >The first PC covers ~40% of the variation. The first 4 PCs cover ~82% of the variation. -->
+
+<!-- ```{r DATA PCA Scree Plot NO SENTINEL, echo = F} -->
+<!-- get_eigenvalue(PCA.DATA.NOSENT) -->
+<!-- fviz_eig(PCA.DATA.NOSENT, addlabels=T, ylim = c(0,40)) -->
+<!-- ``` -->
+
+<!-- >Excluding sentinel, the first PC covers ~43% of the variation, and the first 4 PCs cover ~86% of the total variation. -->
+
+<!-- ### Contribution of variables to Principal Components -->
+
+<!-- ```{r PCA DATA var results, include = F} -->
+<!-- DATA.var<-get_pca_var(PCA.DATA) -->
+<!-- DATA.NOSENT.var<-get_pca_var(PCA.DATA.NOSENT) -->
+<!-- ``` -->
+<!-- ```{r PCA DATA Contribution Dim 1, echo=F} -->
+<!-- fviz_contrib(PCA.DATA, choice="var", axes = 1, top = 12.5) -->
+<!-- ``` -->
+
+<!-- >Very weird, Julian Date, Decimal time, group size and generalized environment contribute greatly to PC1. -->
+
+<!-- ```{r PCA DATA Contribution Dim 2, echo=F} -->
+<!-- fviz_contrib(PCA.DATA, choice="var", axes = 2, top = 12.5) -->
+<!-- ``` -->
+
+<!-- >Frequency of disturbances, group size (again), temperature and the presence of a sentinel contribute most to PC2. -->
+
+<!-- ```{r PCA DATA Contribution Dim 3, echo=F} -->
+<!-- fviz_contrib(PCA.DATA, choice="var", axes = 3, top = 12.5) -->
+<!-- ``` -->
+
+<!-- >Temperature, Bait presence and decimal time contribute most to PC3. -->
+
+<!-- ```{r PCA DATA Contribution Dim 4, echo=F} -->
+<!-- fviz_contrib(PCA.DATA, choice="var", axes = 4, top = 12.5) -->
+<!-- ``` -->
+
+<!-- >Sentinel presence, environment and frequency of disturbances contribute most to PC4. -->
+
+<!-- >I'm still a little unclear as to how I should interpret these results. -->
+
+<!-- #### No Sentinel -->
+
+<!-- ```{r PCA DATA.NOSENT Contribution Dim 1, echo=F} -->
+<!-- fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 1, top = 14.29) -->
+<!-- ``` -->
+
+<!-- >Same 4 factors contribute the most to PC1 in the absence of the sentinel factor. -->
+
+<!-- ```{r PCA DATA.NOSENT Contribution Dim 2, echo=F} -->
+<!-- fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 2, top = 14.29) -->
+<!-- ``` -->
+
+<!-- >Same as above, frequency of disturbances, temperature and group size contribute most to PC2 in the absence of the sentinel factor. -->
+
+<!-- ```{r PCA DATA.NOSENT Contribution Dim 3, echo=F} -->
+<!-- fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 3, top = 14.29) -->
+<!-- ``` -->
+
+<!-- >Temperature, Bait presence and frequency of disturbances contribute the most to PC3 in the absence of a sentinel. -->
+
+<!-- ```{r PCA DATA.NOSENT Contribution Dim 4, echo=F} -->
+<!-- fviz_contrib(PCA.DATA.NOSENT, choice="var", axes = 4, top = 14.29) -->
+<!-- ``` -->
+
+<!-- >Environment contributes most to PC4 in the absence of a sentinel. -->
+
+<!-- ### BiPlots -->
+<!-- ```{r DATA PCA Biplot PC1-2, echo=F} -->
+<!-- fviz_pca_biplot(PCA.DATA -->
+<!--                 ,geom.ind = "point" -->
+<!--                 , repel = T -->
+<!--                 , col.ind=PCA.DATA.SHEET$BEHAVIOR -->
+<!--                 , addEllipses = T -->
+<!--                 , label = "var" -->
+<!--                 , palette=cbPalette -->
+<!--                 , col.var="black" -->
+<!--                 , legend.title = "Behavior") -->
+<!-- ``` -->
+
+<!-- >Messy, can't see where the behaviors lie. -->
+
+<!-- ```{r DATA PCA Biplot PC1-4, echo=F} -->
+<!-- fviz_pca_biplot(PCA.DATA -->
+<!--                 , axes = c(1,4) -->
+<!--                 , geom.ind = "point" -->
+<!--                 , repel = T -->
+<!--                 , col.ind=PCA.DATA.SHEET$BEHAVIOR -->
+<!--                 , addEllipses = T -->
+<!--                 , label = "var" -->
+<!--                 , palette=cbPalette -->
+<!--                 , col.var="black" -->
+<!--                 , legend.title = "Behavior") -->
+<!-- ``` -->
+
+<!-- >Still messy. -->
+
+<!-- ```{r DATA PCA Biplot PC2-3, echo=F} -->
+<!-- fviz_pca_biplot(PCA.DATA -->
+<!--                 , axes = c(2,3) -->
+<!--                 , geom.ind = "point" -->
+<!--                 , repel = T -->
+<!--                 , col.ind=PCA.DATA.SHEET$BEHAVIOR -->
+<!--                 , addEllipses = T -->
+<!--                 , label = "var" -->
+<!--                 , palette=cbPalette -->
+<!--                 , col.var="black" -->
+<!--                 , legend.title = "Behavior") -->
+<!-- ``` -->
+
+<!-- >Biplots are messy if we don't include behavior. -->
+
+<!-- >Same is observed if we run the biplot in the absence of the sentinel factor. -->
+
+<!-- >I'm not sure how to interpret these results. -->
 
@@
 >Seems like peck rate increases significantly in the presence of bait.
+
+# Sentinel Presence
+
+I'll try to use a correlation test to determine if the presence of a sentinel (Yes/No, binary data) is correlated with the type of environment, the frequency of disturbances and the group size.
+
+```{r Correlation test}
+cor(DATA[, c("CODED_SENTINEL_PRESENCE", "CODED_ENV", "CODED_GROUP_SIZE", "TOTAL_FREQUENCY_OF_DISTURBANCES")])
+```

### 2023-06-23 Data & Scripts/BOUT.Rmd
Desktop Push
Push from desktop

@@
 BOUT<-read.csv("BOUT.csv", stringsAsFactors = T)
-names(BOUT)[14]<-"DURATION"
 str(BOUT)
@@
 ```{r BOUT MODEL 1}
-BOUT.MOD1<-lmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+(1|ID), data = BOUT)
+BOUT.MOD1<-lmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+DISTURBANCE_FREQUENCY+(1|ID), data = BOUT)
 
@@
 
+Finally, the frequency of disturbances had a significant effect. I will include an interaction between this factor and generalized environment, as well as with sentinel presence, and behavior.
+
 ```{r BOUT MODEL 2}
-BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+(1|ID), data = BOUT)
+BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|ID)+(1|VIDEO_ID), data = BOUT)
 
@@
 This is curious, especially since the effect size appears to be positive when bait is present.
-My interpretation of this is moving bout durations increase when bait is present.
+My interpretation of this is moving bout duration increase when bait is present.
+
+By adding the interactions, the frequency of disturbances, whether alone or interacting with another factor, becomes insignificant, so I've removed the interactions.
 
@@
 ```{r RLMM MOD 2}
-BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+(1|ID), data = BOUT)
+BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|ID), data = BOUT)
 
@@
 The assumptions of the LMM do not appear to be violated.
+
+#### Post Hoc testing
+
+```{r Post Hoc}
+DIFF<-emmeans::emmeans(BOUT.MOD2, ~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BAIT_PRESENCE,by = "BEHAVIOR", adjust = "fdr",pbkrtest.limit = 5070)
+PAIRS<-emmeans::contrast(DIFF, "pairwise", adjust="fdr")
+PAIRS
+```
 
@@
                      , color = BEHAVIOR))+
-  geom_point(position = position_dodge2(width = 0.9)
+  geom_point(position = "dodge"
              , size = 3) +
@@
                      , color = SENTINEL_PRESENCE))+
-  geom_point(position = position_dodge2(width = 0.9)
+  geom_point(position = position_dodge(width = 0.9)
              , size = 3) +
@@
                , color = BEHAVIOR
-               , shape = GROUP_SIZE
-               , group = BEHAVIOR))+
-  geom_point(position = position_dodge2(width = 0.5)
+               , shape = GROUP_SIZE))+
+  geom_point(stat="identity"
+             , position = position_dodge(width = 0.9)
              , size = 3) +
@@
                 , width = 0.1
-                , position = position_dodge2())+
+                , position = position_dodge(width=0.9))+
   theme_classic() +
@@
                , color = BEHAVIOR
-               , shape = BAIT_PRESENCE
-               , group = BEHAVIOR))+
-  geom_point(position = position_dodge2(width = 0.5)
+               , shape = BAIT_PRESENCE))+
+  geom_point(position = position_dodge(width = 0.9)
              , size = 3) +
@@
 ```
+
+#### Disturbance Frequency
+
+```{r Disturb}
+DF.DOT<-BOUT %>%
+    ggplot(aes(x = DISTURBANCE_FREQUENCY
+               , y = DURATION
+               , color = BEHAVIOR))+
+  geom_point(position = "jitter"
+             , size = 3) +
+  theme_classic() +
+  ylab("Mean Bout Duration (s)") +
+  scale_colour_manual(values = cbPalette
+                      , guide="none")+
+  theme(axis.title.x = element_text("Disturbance per Minute")
+        , legend.position = "bottom")
+
+DF.DOT
+```

### 2023-06-23 Data & Scripts/BOUT.Rmd
Auto stash before merge of "main" and "origin/main"

@@
 ```{r BOUT MODEL 1}
-BOUT.MOD1<-lmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+DISTURBANCE_FREQUENCY+(1|ID), data = BOUT)
+BOUT.MOD1<-lmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
 
@@
 ```{r BOUT MODEL 2}
-BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|ID)+(1|VIDEO_ID), data = BOUT)
+BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
 
@@
 
-In the second model with the added interactions, the behavior, generalized environment, group size and a number of interactions are significant.
-
-The group size becoming significant is puzzling. I don't understand why now it would become significant.
+In the second model with the added interactions, the behavior, generalized environment and a number of interactions are significant.
+The interaction between behavior and generalized environment is significant, as well as the interaction between the presence of a sentinel and generalized environment.
+The interaction between behavior and group size is significant.
+Lastly, moving and bait presence were significant.
 
@@
 ```{r BOUT MOD 3}
-BOUT.MOD3<-lmer(LDURATION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+BEHAVIOR*GENERALIZED_ENVIRONMENT*SENTINEL_PRESENCE+BEHAVIOR*GROUP_SIZE*SENTINEL_PRESENCE+BEHAVIOR*BAIT_PRESENCE+(1|ID), data = BOUT)
+BOUT.MOD3<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+(1|VIDEO_ID/ID), data = BOUT)
 
@@
 
-Nice! The second model has the lowest AIC value and is therefore prefered.
+Nice! The second model has the lowest AIC value and is therefore preferred.
 
@@
 ```{r RLMM MOD 2}
-BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|ID), data = BOUT)
+BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
 
@@
                   , title = "Bout Robust Model 2 Output"
-                  , dv.labels = " Effects of on duration of bouts of behaviors")
+                  , dv.labels = " Effects on duration of bouts of behaviors")
 ```
@@
 -   There is a significant interaction of sentinel presence and generalized environment.
--   There is a significant interaction of behavior and group size for head up and moving.
+-   There is a significant interaction of behavior and group size
 -   There is a significant interaction of "moving" and bait presence.
@@
 
-#### Behavior X Generalized Environment
+### Behavior X Generalized Environment
 
@@
 
-Very interesting! Duration of bouts of alertness and moving remain largely the same, while the duration of foraging bout in green areas increases dramatically. \
+Very interesting! Duration of bouts of alertness and moving remain largely the same, while the duration of foraging bout in green areas increases dramatically.\
 
@@
 
-#### Sentinel Presence X Generalized Environment
+### Sentinel Presence X Generalized Environment
 
@@
 
+It could be that sentinels have a stabilizing effect on the duration of bouts
+
 ```{r BxGS BOUT Dot Plot}
@@
 
-#### Behavior x Bait Presence
+### Behavior x Bait Presence
 

### 2023-06-23 Data & Scripts/BOUT.Rmd
June 23rd
Late night, couldn't sleep :(

@@
 ---
-
-# BOUT
 
@@
 library("doBy")
+library("emmeans")
 ```
@@
 
-This script will deal with analysing the relationship of the duration of bouts of behavior - head down, head up, and moving - with the presence of a sentinel, the presence of bait, the foraging environment and the group size.
+This script will deal with analyzing the relationship of the duration of bouts of behavior - head down, head up, and moving - with the presence of a sentinel, the presence of bait, the foraging environment, the group size and the frequency of disturbances.
 
-As random effects, I've chosen the individual ID, the (julian) date, and the (decimal) time.
+As random effects, I've chosen the video and individual ID
 
 This analysis will answer the following question:\
-Do individual vary the duration of behavioral bouts in response to their perception of risk and sentinel coverage in their foraging environment.
+Do individual vary the duration of behavioral bouts in response to their perception of risk and sentinel coverage in their foraging environment?
 
@@
 
-```{r Crow Histogram, echo=F}
-HIST<-ggplot(data = BOUT) +
+```{r BOUT Histogram, echo=F}
+BOUT.HIST<-ggplot(data = BOUT) +
   geom_histogram(aes(fill = BEHAVIOR, x = LDURATION)
@@
   ylab("Count")
-HIST
+BOUT.HIST
 ```
@@
 ```{r BOUT MODEL 2}
-BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
+BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
 
@@
 ```{r RLMM MOD 2}
-BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
+BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
 
@@
 
-The assumptions of the LMM do not appear to be violated.
+There's a horizontal line on the Q-Q vs. random effects. In the outputs of the models, the videoID alone does not account for any variation.
 
@@
 
-```{r Post Hoc}
-DIFF<-emmeans::emmeans(BOUT.MOD2, ~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BAIT_PRESENCE,by = "BEHAVIOR", adjust = "fdr",pbkrtest.limit = 5070)
-PAIRS<-emmeans::contrast(DIFF, "pairwise", adjust="fdr")
-PAIRS
+```{r BOUT Post Hoc}
+BOUT.DIFF<-emmeans(BOUT.MOD2, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
+test(pairs(BOUT.DIFF, by= "BEHAVIOR"), adjust="fdr")
 ```
+Very cool! These are the results of pairwise comparisons averaged over the levels of group size and bait presence (**Not disturbance frequency**).
+
+These pairwise comparisons show there are no significant differences in the bout duration of 'moving' or 'head up' between each combination of 'sentinel_presence' and 'generalized_environment'.
+
+However, there are differences in the bout duration of 'head down' behavior.
+There are differences caused by the presence of a sentinel in commercial areas.
+There are significant differences between bout durations in commercial areas without a sentinel and bout durations in green areas, in both the presence and absence of a sentinel.
+
+Let's quickly have a look at that:
+
+```{r Pairwise comparisons}
+BOUT.DOT
+```
+
+Huh... I would have expected that there would be significant differences in bouts of 'head up' behavior.
 
@@
                      , color = BEHAVIOR))+
-  geom_point(position = "dodge"
+  geom_point(position = position_dodge(width=0.9)
              , size = 3) +
@@
 
-GGPLOT! The error bars are not over the dots.
+Ok, we see that smaller group sizes lead to shorter bout durations.
 
@@
 
+The presence of bait has the effect of decreasing the duration of 'head down' and 'head up' behavior, yet increased the duration of bouts of movement.
+
 #### Disturbance Frequency
@@
                , color = BEHAVIOR))+
-  geom_point(position = "jitter"
-             , size = 3) +
+  geom_point(size = 3) +
   theme_classic() +
-  ylab("Mean Bout Duration (s)") +
-  scale_colour_manual(values = cbPalette
-                      , guide="none")+
-  theme(axis.title.x = element_text("Disturbance per Minute")
-        , legend.position = "bottom")
+  labs(x="Mean Bout Duration (s)", y="Disturbances per Minute", color= "Behavior") +
+  scale_colour_manual(values = cbPalette, labels = c("Foraging", "Alert", "Moving"))+
+  scale_x_continuous(n.breaks=14)
+  theme(legend.position = "bottom")
 
@@
 ```
+
+Finally, it seems like as the frequency of disturbances increases, the duration of bouts decreases.

### 2023-06-23 Data & Scripts/PROP.Rmd
June 23rd
Late night, couldn't sleep :(

---
title: "PROP"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

## Preamble

This script will deal with analyzing the relationship of the proportion of time allocated to each behavior - head down, head up, and moving - with the presence of a sentinel, the presence of bait, the foraging environment, group size and the frequency of disturbances.

As random effects, I've chosen the video and individual ID

This analysis will answer the following question:\
Do individual vary the proportion of time allocated to each behavioral in response to their perception of risk and sentinel coverage in their foraging environment?

To accomplish this, I extracted the summaries from the videos is BORIS and compiled a data table named "DATA".

```{r Open Data}
PROP<-read.csv("DATA.csv", stringsAsFactors = T)

str(PROP)
```
```{r Bout Summary}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
PROP.MEAN <- summarySE(data = PROP
                      , measurevar = "BEHAVIOR_PROPORTION"
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "BEHAVIOR"
                                      , "SENTINEL_PRESENCE"
                                      )
                      )
PROP.MEAN
```

We now have a summary of the proportion data.

Onto the analysis.

## To transform or not

In order to determine whether our fixed effects have significant effects on the allocation of time to each behavior, I decided to first arcsin square root transform the proportions.

```{r Proportion Histograms Individual, include = F}
Untransformed.Prop<-hist(PROP$BEHAVIOR_PROPORTION, breaks=25)
Transformed.Prop<-hist(asin(sqrt(PROP$BEHAVIOR_PROPORTION)), breaks=25)
```

```{r Proportion Histograms Combined, echo=F}
op <- par(mfrow = c(1,2))
plot(Untransformed.Prop, main="Untransformed", ylab="Frequency", xlab="Proportion")
plot(Transformed.Prop, main="Transformed", ylab="", xlab="Asin(Sqrt(Proportion)")
par(op)
```

Transformed data seems more normal than untransformed data. Will continue to use transformed data.

```{r Data Transformation}
PROP$ASIN_PROPORTION<-asin(sqrt(PROP$BEHAVIOR_PROPORTION))
```

Here is the histogram of the log-transformed data separated by behavior.

```{r PROP Histogram, echo=F}
PROP.HIST<-ggplot(data = PROP) +
  geom_histogram(aes(fill = BEHAVIOR, x = ASIN_PROPORTION)
                 , alpha = 1
                 , bins = 25)+
  scale_fill_manual(values=cbPalette
                    , labels = c("Foraging", "Alert", "Moving"))+
  xlab("Transformed Proportion")+
  ylab("Count")
PROP.HIST
```

There are some gaps in the distribution.

## Stacked barplot

GGPlot time

```{r Proportion Barplot}
PROP.BAR<-ggplot(PROP.MEAN
                 , aes(x = SENTINEL_PRESENCE
                       , y = BEHAVIOR_PROPORTION
                       , fill = BEHAVIOR))+
  geom_bar(stat = 'identity'
           , position = 'stack')+
  geom_text(aes(label = paste0(formattable::digits(BEHAVIOR_PROPORTION*100, dig=2)
                               , "%"))
            , position = position_stack(vjust = 0.5)
            , size = 3) +
  scale_y_continuous(labels = scales::percent) +
  theme_classic()+
  theme(text=element_text(size = 10)
        , axis.title.x = element_blank())+
  ylab("Proportion of time")+
  scale_x_discrete(labels=c("Sentinel Absent"
                            , "Sentinel Present"))+
  scale_fill_manual(values = cbPalette, labels = c("Foraging"
                                                   , "Alert"
                                                   , "Moving")
                    , name="Behavior")+
  theme(legend.position = "bottom")+
  facet_grid(~GENERALIZED_ENVIRONMENT)
PROP.BAR
```

Ok, so we can see that there are a couple of differences between the proportions.

It seems that the presence of a sentinel increases the proportion of 'moving' behavior. \
It seems like 'Alert' behavior remains relatively consistent across environments, with the presence of a sentinel decreasing the proportion of time allocated to that behavior.
Lastly, it seems like 'foraging' behavior remains consistent throughout environments and the presence of a sentinel.

Let's fit some models.

## The Models

Let's fit the models. We are principally testing for the effects of the presence of a sentinel and the generalized environment. In addition to those, we will include the presence of bait, group size and the frequency of disturbances as fixed effects.

As for the random effects, I will use VIDEO_ID/ID. The ID is nested within the VIDEO ID, explaining the structure.

Let's fit the simplest model.

```{r PROP MODEL 1}
PROP.MOD1<-lmer(ASIN_PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = PROP)

sjPlot::tab_model(PROP.MOD1
                  , show.re.var = T
                  , title = "Proportion Model 1 Output"
                  , dv.labels = " Effects of on the proportion of behaviors")
```

huh, seems like the model is singular. This could mean that the model is overfitted.

Let's look at the results.

```{r Proportion Model Results, echo = F}
Anova(PROP.MOD1)
ranova(PROP.MOD1)
```

Ok, so none of the factors are significant, and the results of the ranova suggest that the model without random effects is preferred. Let's improve the model with some interactions. We will use the same formula for MOD2 as the one used in the BOUT data.

```{r PROP MODEL 2}
PROP.MOD2<-lmer(ASIN_PROPORTION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = PROP)

sjPlot::tab_model(PROP.MOD2
                  , show.re.var = T
                  , title = "Proportion Model 2 Output"
                  , dv.labels = " Effects on proportion of behaviors")
```

Like MOD1, none of the factors have an effect.

Maybe I'm not using the correct approach. Instead of performing a LMM on transformed data, I'll instead use a quasibinomial GLM for proportion data. Since the random effects in the previous models explained none of the variance, we can, in this case, omit them.

```{r PROP MODEL 3}
PROP.MOD3<-glm(BEHAVIOR_PROPORTION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY, family=quasibinomial, data = PROP)

sjPlot::tab_model(PROP.MOD3
                  , show.re.var = T
                  , title = "Proportion Model 3 Output"
                  , dv.labels = " Effects on proportion of behaviors")
```

All factors are non-significant apart from the interaction between the proportion of 'moving' and the generalized environment.

I believe this may be the best approach since I am unable to use the quasibinomial family in a glmm.

Let's look at the comparisons between groups.

```{r PROP Post Hoc}
PROP.DIFF<-emmeans(PROP.MOD3, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
test(pairs(PROP.DIFF, by= "BEHAVIOR"), adjust="fdr")
```

None of the groups are significantly different from one another when comparing within each behavior. Next, let's look at the differences throughout behaviors.

```{r PROP Post Hoc 2}
PROP.DIFF2<-emmeans(PROP.MOD3, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
PROP.PWISE<-test(pairs(PROP.DIFF2), adjust="fdr")
PROP.PWISE
```

Ok! Across behaviors we have few significant and marginally significant results.

```{r PROP PWISE}
PROP.PWISE[which(PROP.PWISE$p.value<0.1),]
```
Very cool! These are the results of pairwise comparisons averaged over the levels of group size and bait presence (**not disturbance frequency**).

There do not appear to be any differences across environments or the presence of a sentinel, but between proportions of behaviors.

Let's have another look at it.

```{r Proportion Pairwise}
PROP.BAR
```

Indeed, there is a difference between Moving and Alert proportions in commercial areas in the absence of a sentinel.
Likewise, there are differences in the proportion of time allocated to head up/down and moving behaviors in green spaces in the absence of a sentinel.
Possibly the most interesting comparison is the difference between the proportion of time allocated to 'alert' behavior and 'moving' in the absence of a sentinel and across the two environments.

Yet, these differences are marginally significant, apart from 'HU NO Green Area - M NO Green Area'

## Conclusions

It seems that the proportion of time allocated to each behavior remains relatively consistent, and not affected by the presence of a sentinel, nor the environment in which the individual forages in. \
However, the differences appear between the proportion of behaviors, suggesting that the allocation of time to each behavior is relatively dependent on the type of behavior. \

This would suggest that if the presence of a sentinel and the environment were to affect the behavior of foragers, the effects would be seen in the frequency of behaviors. In other words, instead of changing the overall proportion of time allocated to each behavior, they instead change the duration of bouts of their behaviors to provide better coverage over time (more, but shorter gaps between instances of vigilance).

### 2023-08-17 Data & Scripts/BOUT.Rmd
Aug.17 pt.2
Desktop commit

@@
                      , shape = SENTINEL_PRESENCE))+
-  geom_point(position = position_dodge2(width = 0.9)
+  geom_point(position = position_dodge(width = 0.9)
              , size = 3) +
@@
   theme_classic() +
-  labs(x="Mean Bout Duration (s)", y="Disturbances per Minute", color= "Behavior") +
+  labs(y="Mean Bout Duration (s)", x="Disturbances per Minute", color= "Behavior")+
   scale_colour_manual(values = cbPalette, labels = c("Foraging", "Alert", "Moving"))+
-  scale_x_continuous(n.breaks=14)
+  scale_x_continuous(n.breaks=14)+
   theme(legend.position = "bottom")

### 2023-08-17 Data & Scripts/PCA.Rmd
Aug.17 pt.2
Desktop commit

---
title: "PCA"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: html_document
---

---

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
library("factoextra")
library("FactoMineR")
library("corrr")
library("ggcorrplot")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

## Preamble

This script will cover the pathway analysis. I've already performed a random permutation test in BORIS, but I would like to look at the individual level.

To do this, I've extracted the frequencies of one behavior following another (Head Down => Head Up) and will perform a pca to reduce the number of models I need to run.

That being said, I would, at most, need to run 9 models. However, I believe that certain transitions can be grouped together (e.g. transitions to 'alert' state).

To do this, I exported the behavioral sequences in BORIS, then calculated the frequencies of each transition, per individual, and compiled the 'PTWY' dataset.

```{r Open PTWY}
PCA<-read.csv("DATA.SR.csv", stringsAsFactors = T)
str(PCA)
```

Visualizing the data will be difficult in R, but the summaries and their visualizations can be found in Alex-Popescu-Thesis-2023\BORIS\Pathway Analysis\Flow Diagrams.

## Data distribution

```{r Data prep}
PCA.DATA<-PCA[1:81,c(2, 13, 31:34, 36:39, 41:44, 47, 50, 53, 56, 59, 62, 65, 68, 71, 74)]
PCA.DATA$HU_RATE<-PCA.DATA$HU_NUMBER_OF_BOUTS/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HD_RATE<-PCA.DATA$HD_NUMBER_OF_BOUTS/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$M_RATE<-PCA.DATA$M_NUMBER_OF_BOUTS/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HD.HDP_RATE<-PCA.DATA$NB_HD.HDP/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HD.HU_RATE<-PCA.DATA$NB_HD.HU/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HD.M_RATE<-PCA.DATA$NB_HD.M/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HDP.HU_RATE<-PCA.DATA$NB_HDP.HU/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HDP.M_RATE<-PCA.DATA$NB_HDP.M/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HU.HD_RATE<-PCA.DATA$NB_HU.HD/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$HU.M_RATE<-PCA.DATA$NB_HU.M/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$M.HD_RATE<-PCA.DATA$NB_M.HD/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA$M.HU_RATE<-PCA.DATA$NB_M.HU/(PCA.DATA$RECORDED_DURATION/60)
PCA.DATA<-PCA.DATA[1:81,c(1,4:6,8:10,12:15,25:36)]
```
```{r Distribution RAW}
PCA.DATA.LONG<-na.omit(pivot_longer(PCA.DATA, -ID))

ggplot(PCA.DATA.LONG
       , aes(x=value)) +
  geom_histogram() +
  geom_vline(data=PCA.DATA.LONG
       , aes(xintercept=mean(value), color="red")) +
  geom_vline(data=PCA.DATA.LONG
       , aes(xintercept=median(value), color="blue"))+
  facet_wrap(~name)
```

```{r Distribution LOG, include = F}
LOG<-as.data.frame(sapply(PCA.DATA[,-1], function(x) {log(x+0.01)}))
LOG.LONGER<-na.omit(pivot_longer(LOG, 1:22))

ggplot(LOG.LONGER
       , aes(x=value)) +
  geom_histogram() +
  geom_vline(data=LOG.LONGER
       , aes(xintercept=mean(value), color="red")) +
  geom_vline(data=LOG.LONGER
       , aes(xintercept=median(value), color="blue")) +
  facet_wrap(~name)
```

```{r Distribution ASIN SCALED}
SC.LOG<-as.data.frame(scale(LOG))
SC.LOG<-na.omit(SC.LOG)
SC.LOG.LONG<-pivot_longer(SC.LOG, 1:22)

ID = "ID."
ggplot(SC.LOG.LONG
       , aes(x=value)) +
  geom_histogram() +
  geom_vline(data=SC.LOG.LONG
       , aes(xintercept=mean(value), color="red")) +
  geom_vline(data=SC.LOG.LONG
       , aes(xintercept=median(value), color="blue")) +
  facet_wrap(~name)
```

## PCA

```{r PTWY pca}
#Computing correlation matrix
corr_matrix <- cor(na.omit(SC.LOG))

#Showing the correlation matrix
ggcorrplot(corr_matrix)

```

Pretty interesting. Redder = most positively correlated. Bluer = most negatively correlated. Let's apply the PCA.

```{r PCA scores}
#Computing the PCA
pca.data<-prcomp(SC.LOG)

#Show the results
summary(pca.data)
```

Pretty interesting. 
The first principal component explains 25.32% of the total variance. 
The second principal component explains 16.80% of the total variance. 
The third principal component explains 12.14% of the total variance.
The fourth principal component explains 9.74% of the total variance.
The fifth principal component explains 8.76% of the total variance.
The sixth principal component explains 7.54% of the total variance.
The seventh principal component explains 5.94% of the total variance.
The eighth principal component explains 4.86% of the total variance.

Let's visualize this with a scree plot.

```{r pca scree}
#The Scree plot to show how much of the variance each PC explains
fviz_eig(pca.data, addlabels = T, ylim = c(0,50))

```
The first eight PCs explain 91.11% of the variation!

## Graph of individuals

```{r PCA ind}
fviz_pca_ind(pca.data,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = cbPalette[1:3],
             repel = TRUE     # Avoid text overlapping
             )
```

## Graph of Variables

```{r PCA var}
fviz_pca_var(pca.data,
             col.var = "contrib", # Color by the quality of representation
             gradient.cols = cbPalette[1:3],
             repel = TRUE     # Avoid text overlapping
             )
```

## Eigenvalues

```{r PCA eig}
eig.val <- get_eigenvalue(pca.data)
eig.val

mean(eig.val$eigenvalue)
```

OK! Using the Kaiser-Guttman criterion, we end up retaining 8 PCs (mean eigenvalue = 0.4430, anything with an eigenvalue < 0.4430 is rejected). Let's look at how each factor fits into each principal component using their loadings.

```{r pca loadings}
-1*pca.data$rotation[,1:8]

#Criterion for "large" values
sqrt(1/ncol(SC.LOG)) #0.2132
```
Using the criterion for "large" results (=0.2132), we find that "peck rate", "HU Rate", "HU>HD Rate", and "HU.M Rate" do not fall into any PCs we've kept.

Should I remove them? Yes. Let's do that.

```{r remove columns}
colnames(SC.LOG)
SC.LOG2<-SC.LOG[,c(-10,-11,-19,-20)]
```

And now we re-run the PCA

## PCA2

```{r PTWY pca2}
#Computing correlation matrix
corr_matrix2 <- cor(na.omit(SC.LOG2))

#Showing the correlation matrix
ggcorrplot(corr_matrix2)

```

```{r PCA2 scores}
#Computing the PCA
pca.data2<-prcomp(SC.LOG2)

#Show the results
summary(pca.data2)
```

Nice!

The first 8 PCs explain a total of 92.11% of the variation, slightly more than the first PCA.

As before, let's plot a scree plot.

```{r PCA2 scree}
#The Scree plot to show how much of the variance each PC explains
fviz_eig(pca.data2, addlabels = T, ylim = c(0,50))

```

Generally, the PCs explain more variation that in the first PCA.

## Graph of individuals 2

```{r PCA ind}
fviz_pca_ind(pca.data2,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = cbPalette[1:3],
             repel = TRUE     # Avoid text overlapping
             )
```

## Graph of Variables 2

```{r PCA var}
fviz_pca_var(pca.data2,
             col.var = "contrib", # Color by the quality of representation
             gradient.cols = cbPalette[1:3],
             repel = TRUE     # Avoid text overlapping
             )
```

## Eigenvalues 2

```{r PCA eig}
eig.val <- get_eigenvalue(pca.data2)
eig.val

mean(eig.val$eigenvalue)
```

The mean eigenvalue is 0.5208. As a result, we keep 7 PCs. These PCs explain 87.17% of the total variation.

Let's look at the loadings again.

```{r pca loadings}
-1*pca.data2$rotation[,1:7]

#Criterion for "large" values
sqrt(1/ncol(SC.LOG2)) #0.2357
```

The magic number this time is 0.2357. All factors are included in the PCs. The next question is can I make this better? Is there any way for me to explain how these factors fit into each PC? Can I reduce the number of PCs?

RUN A MANOVA!!!

### 2023-08-17 Data & Scripts/PROP.Rmd
Aug.17 pt.2
Desktop commit

@@
 ```{r Open Data}
-PROP<-read.csv("DATA.csv", stringsAsFactors = T)
+DATA<-read.csv("DATA.csv", stringsAsFactors = T)
 
-str(PROP)
+str(DATA)
 ```
@@
 source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
-PROP.MEAN <- summarySE(data = PROP
+PROP.MEAN <- summarySE(data = DATA
                       , measurevar = "BEHAVIOR_PROPORTION"
@@
 ```{r Proportion Histograms Individual, include = F}
-Untransformed.Prop<-hist(PROP$BEHAVIOR_PROPORTION, breaks=25)
-Transformed.Prop<-hist(asin(sqrt(PROP$BEHAVIOR_PROPORTION)), breaks=25)
+Untransformed.Prop<-hist(DATA$BEHAVIOR_PROPORTION, breaks=25)
+Transformed.Prop<-hist(asin(sqrt(DATA$BEHAVIOR_PROPORTION)), breaks=25)
 ```
@@
 ```{r Data Transformation}
-PROP$ASIN_PROPORTION<-asin(sqrt(PROP$BEHAVIOR_PROPORTION))
+DATA$ASIN_PROPORTION<-asin(sqrt(DATA$BEHAVIOR_PROPORTION))
 ```
@@
 ```{r PROP Histogram, echo=F}
-PROP.HIST<-ggplot(data = PROP) +
+PROP.HIST<-ggplot(data = DATA) +
   geom_histogram(aes(fill = BEHAVIOR, x = ASIN_PROPORTION)
@@
 ```{r PROP MODEL 1}
-PROP.MOD1<-lmer(ASIN_PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = PROP)
+PROP.MOD1<-lmer(ASIN_PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = DATA)
 
@@
 
-huh, seems like the model is singular. This could mean that the model is overfitted.
+huh, seems like the model is singular. This could mean that the model is over fitted.
 
@@
 ```{r PROP MODEL 2}
-PROP.MOD2<-lmer(ASIN_PROPORTION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = PROP)
+PROP.MOD2<-lmer(ASIN_PROPORTION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = DATA)
 
@@
 
-Like MOD1, none of the factors have an effect.
+Like MOD1, none of the factors have an effect. Still singular.
 
@@
 ```{r PROP MODEL 3}
-PROP.MOD3<-glm(BEHAVIOR_PROPORTION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY, family=quasibinomial, data = PROP)
+PROP.MOD3<-glm(BEHAVIOR_PROPORTION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY, family=quasibinomial, data = DATA)
 
@@
 ```{r PROP Post Hoc}
-PROP.DIFF<-emmeans(PROP.MOD3, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
+PROP.DIFF<-emmeans(PROP.MOD3, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT)
 test(pairs(PROP.DIFF, by= "BEHAVIOR"), adjust="fdr")
@@
 ```{r PROP Post Hoc 2}
-PROP.DIFF2<-emmeans(PROP.MOD3, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
+PROP.DIFF2<-emmeans(PROP.MOD3, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT)
 PROP.PWISE<-test(pairs(PROP.DIFF2), adjust="fdr")

### 2023-08-17 Data & Scripts/PTWY.Rmd
Aug.17 pt.2
Desktop commit

---
title: "PTWY"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---
```{r Libraries}
library("lavaan")
library("lavaanPlot")
```

```{r Open DATA.SR}
PTWY<-read.csv("DATA.SR.csv", stringsAsFactors = T)
str(PTWY)
```
```{r PTWY Analysis}
PTWY.ENV <- '

# measurement model

ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TOTAL_FREQUENCY_OF_DISTURBANCES

# intercepts

CODED_ENV ~ 1
CODED_BAIT ~ 1
DECIMAL_TIME ~ 1
TOTAL_FREQUENCY_OF_DISTURBANCES ~ 1
'

PTWY.Fit.ENV <- sem(PTWY.ENV, data = PTWY)

summary(PTWY.Fit.ENV, fit.measures = TRUE, standardized = TRUE)
```

```{r PTWY.SOC Analysis}
PTWY.Mod.SOC <- '

# measurement model 
SOCIAL =~ CODED_GROUP_SIZE

# intercepts
CODED_GROUP_SIZE ~ 1
'

PTWY.Fit.SOC <- sem(PTWY.Mod.SOC, data = PTWY)

summary(PTWY.Fit.SOC, fit.measures = TRUE, standardized = TRUE)
```

```{r PTWY Analysis}
PTWY.Mod <- '

# measurement model

ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TEMPERATURE + TOTAL_FREQUENCY_OF_DISTURBANCES

VIGILANCE =~ PTWY_HDP.HU + PTWY_HD.HU + PTWY_M.HU

FORAGING =~ PTWY_HD.HDP + PTWY_HU.HD + PTWY_M.HD

# regressions

VIGILANCE ~ ENVIRONMENT + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE

FORAGING ~ ENVIRONMENT + VIGILANCE + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE

'

PTWY.Fit <- sem(PTWY.Mod, data = PTWY)

summary(PTWY.Fit, fit.measures = TRUE, standardized = TRUE)
```

### 2023-08-17 Data & Scripts/SENT.Rmd
Aug.17 pt.2
Desktop commit

---
title: "Sentinel Likelihood"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---
```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

# Preamble
This script will deal with determining if different environmental factors affect the likelihood of a sentinel being present in the area. These factors will include group size, temperature, time of day, and disturbance frequency to name a few.

The main test we will be implementing is the binary logistic regression to determine the relationship between a binary target variable (sentinel presence) and a set of independent variables.

We will be reusing the 'Single Row' dataset for this script.

```{r Open SENT}
SENT<-read.csv("DATA.SR.csv", stringsAsFactors = T)
SENT<-SENT[1:81,]
str(SENT)
```
Next, we will subset the dataframe to include only the variables we want to keep.

```{r SENT Subset}
SENT<-SENT[1:81,c(1,2,7,10,11,15,17,22,30,31)]
str(SENT)
dim(SENT)
```
81 observations across 9 variables. BUT that's with all the individuals. Let's remove those replicates.

```{r Selecting unique observations}
SENT<- SENT[,-2] %>%
  distinct()
str(SENT)
```
33 observations across 8 variables. This includes the videos where a transition in sentinel presence occurred.

## Fit the model

```{r SENT model - Simple}
SENT.mod1<-glm(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+GROUP_SIZE+TOTAL_FREQUENCY_OF_DISTURBANCES+DECIMAL_TIME+TEMPERATURE
              , family = binomial()
              , data = SENT)

summary(SENT.mod1)
```

OOF. By removing the duplicates, we now have no significant effects of any variable.
Next, let's try and see if we can run a glmm and include the video ID as a random effect.

```{r SENT GLMM}
SENT.mod2<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+TOTAL_FREQUENCY_OF_DISTURBANCES+DECIMAL_TIME+TEMPERATURE+(1|VIDEO_ID.)
                , family = binomial()       
                , data = SENT
                       )

summary(SENT.mod2)
```
Ok, the model is singular. Let's see if we can improve the model by reducing the number of factors.

```{r AIC testing}
drop1(SENT.mod2, tests = "F")
```
From the results of the drop1 function, we see that the removal of temperature and decimal time would improve the model fit. Let's do that next.

```{r SENT GLMM2}
SENT.mod3<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+TOTAL_FREQUENCY_OF_DISTURBANCES+(1|VIDEO_ID.)
                , family = binomial()       
                , data = SENT
                       )

summary(SENT.mod3)
```
Still singular. It seems like my sample size is too small to determine any significant effects. Let's see if we can remove additional variables.

```{r AIC testing2}
drop1(SENT.mod3, tests = "F")
```
Removing Generalized Environment would improve the model. Let's fit the final model.

```{r SENT GLMM3}
SENT.mod4<-glmer(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES+(1|TEMPERATURE)
                , family = binomial()       
                , data = SENT
                       )

summary(SENT.mod4)
```
Bummer... Still singular. I will need to look into bootstrapping or something else to improve the sample size. As it stands, the total frequency of disturbances appears to have the greatest effect on the likelihood of a sentinel being present. Next, I will run another 'simple' glm.

```{r SENT model - Simple 2}
SENT.mod5<-glm(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES
              , family = binomial()
              , data = SENT)

summary(SENT.mod5)
```

Same results as the glmm. Let's quickly try code I found on R-bloggers (https://www.r-bloggers.com/2015/08/bootstrap-with-logistic-regression/)

```{r Bootstrapped model}
library(boot)
logit_test <- function(d,indices) {  
d <- d[indices,]  
fit <- glmer(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES+(1|VIDEO_ID.)
                , family = binomial()       
                , data = d
                       ) 
return(coef(fit))  
}
boot_fit <- boot(  
   data = SENT, 
   statistic = logit_test, 
   R = 1e5
) 
```

Ok, still singular and the bootstrap failed. It is possible that there are too many levels in the frequency of disturbances. Let's try to simplify by binning the disturbance frequencies.

```{r SENT GLMM 4}
SENT.mod6<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+DECIMAL_TIME+TEMPERATURE+(1|VIDEO_ID.)
                , family = binomial()       
                , data = SENT
                       )

summary(SENT.mod6)
```

Still singular. Let's remove some factors using drop 1.

```{r AIC testing3}
drop1(SENT.mod6, tests = "F")
```

We can remove TEMPERATURE.

```{r SENT GLMM 5}
SENT.mod7<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+DECIMAL_TIME+(1|VIDEO_ID.)
                , family = binomial()       
                , data = SENT
                       )

summary(SENT.mod7)
drop1(SENT.mod7) #remove decimal time
```

```{r SENT GLMM 6}
SENT.mod8<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+(1|VIDEO_ID.)
                , family = binomial()       
                , data = SENT
                       )

summary(SENT.mod8)
drop1(SENT.mod8) #remove environment
```

```{r SENT GLMM 7}
SENT.mod9<-glmer(SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY+(1|VIDEO_ID.)
                , family = binomial()       
                , data = SENT
                       )

summary(SENT.mod9)
drop1(SENT.mod9)
```

Still singular. Let's try removing the random effect and running yet another simple model.

```{r SENT model - Simple 3}
SENT.mod10<-glm(SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY
              , family = binomial()
              , data = SENT)

summary(SENT.mod10)
```

Disturbance frequency is not significant. What about another bootstrap?

```{r Bootstrapped model 2}
library(boot)
logit_test <- function(d,indices, formula) {  
d <- d[indices,]  
fit <- glm(formula
           , family = binomial()
           , data = d
           ) 
return(coef(fit))  
}

boot_fit <- boot(
   data = SENT
   , statistic = logit_test
   , R = 1000
   , formula=SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY
   )

```

RUN A CHI SQUARED!!!

### 2023-08-18 Data & Scripts/PECK.Rmd
Aug 18th
Knitted Scripts, SENT and PTWY done. PCA was archived.

---
title: "PECK"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

## Preamble

This script will deal with analyzing the relationship between the peck rate of foragers and the presence of a sentinel, the presence of bait, the foraging environment, the group size and the frequency of disturbances.

As random effects, I've chosen the video and individual ID

This analysis will answer the following question:\
Is the foraging efficiency of foraging individuals affected by the presence of a sentinel and the environment in which they forage in?

To answer this question, I extracted the number of pecks at bait performed by each individual, then divided it by the total duration of time of behavior 'head down'. The reasoning for this is that individuals can't peck while having their heads up or while moving, therefore the duration of time spent foraging would better describe the foraging efficiency of the individual.

I do expect that bait will have a significant effect on the foraging efficiency of individuals.

I compiled these results in the dataset called 'DATA'.

```{r Open Data}
DATA<-read.csv("DATA.csv", stringsAsFactors = T)
str(DATA)
```
Next, we'll compile the means of the peck rate by sentinel presence and generalized environment.

```{r Peck Summary, echo = F}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
PECK.MEAN<-summarySE(DATA, "PECK_RATE", c("GENERALIZED_ENVIRONMENT", "SENTINEL_PRESENCE"), na.rm=T)
PECK.MEAN
```

At first glance, it seems like the peck rate when a sentinel is absent in commercial and green areas are similar (56 pecks per minute - 52 pecks per minute, respectively). The peck rate in green areas with a sentinel is somewhat lesser than in the absence of sentinels (47 pecks per minute), yet in the presence of a sentinel and in commercial areas the peck rate increases substantially (73 pecks per minute).

Let's visualize this.

## The Dot Plot

```{r PECK Dot Plot}
PECK.DOT<-ggplot(data = PECK.MEAN
               , aes(x = SENTINEL_PRESENCE
                     , y = PECK_RATE
                     , color = SENTINEL_PRESENCE))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(PECK_RATE-se)
                    , ymax=(PECK_RATE+se))
                , width = 0.2
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Mean Pecks per Minute") +
  scale_colour_manual(values = cbPalette
                      , guide="none") +
  scale_x_discrete(labels = c("Sentinel Absent"
                                  , "Sentinel Present")) +
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  facet_grid(~GENERALIZED_ENVIRONMENT)

PECK.DOT
```

Very nice! Basically shows the results of the summary.

So, it seems like the peck rate in commercial areas and in the presence of a sentinel is higher than in green areas (with or without a sentinel) and in commmercial areas without a sentinel.

Onto the transformations!

```{r Peck Per Minute Histograms Individual, include = F}
Untransformed.PPM<-hist(DATA$PECK_RATE, na.rm=T, breaks = 25)
Transformed.PPM<-hist(log(DATA$PECK_RATE), na.rm=T, breaks = 25)
```

```{r Peck Per Minute Histograms Combined, echo=F}
op <- par(mfrow = c(1,2))
plot(Untransformed.PPM, main="Untransformed", ylab="Frequency", xlab="Pecks per minute")
plot(Transformed.PPM, main="Transformed", ylab="", xlab="Log(Pecks per minute)")
par(op)
```

It looks like the untransformed data is more normal than the log-transformed data. For the subsequent analyses, I will use the **untransformed** data.

Let's fit the models

## The Models

We are testing principally for the effects of the presence of a sentinel and the generalized environment. However, group size and the presence of bait could also affect the peck rate of foragers.

The first model fitted will be a simple model with no interactions.

```{r PECK MODEL 1}
PECK.MOD1<-lmer(PECK_RATE~SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT+BAIT_PRESENCE+GROUP_SIZE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = DATA)

sjPlot::tab_model(PECK.MOD1
                  , show.re.var = T
                  , title = "Peck Rate Model 1 Output"
                  , dv.labels = " Effects on peck rate")
```

The intercept is significant, but none of the factors are. The disturbance frequency is marginally significant.

Surprisingly, the random effects account for much more variance than in the proportion data. For the next model, I'll use the same formula as Model 2 for the bout duration and proportion datasets.

```{r BOUT MODEL 2}
PECK.MOD2<-lmer(PECK_RATE~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = DATA)

sjPlot::tab_model(PECK.MOD2
                  , show.re.var = T
                  , title = "Peck Rate Model 2 Output"
                  , dv.labels = " Effects on peck rate")
```

As in the simple model, none of the factors are significant, yet the disturbance frequency has a marginally significant effect on the peck rate.

Let's try running the robust lmm.

```{r Robust PECK MODEL 2}
PECK.RMOD2<-rlmer(PECK_RATE~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = DATA)

sjPlot::tab_model(PECK.MOD2
                  , show.re.var = T
                  , title = "Robust Peck Rate Model 2 Output"
                  , dv.labels = " Effects on peck rate")
```

Still insignificant, with disturbance frequency being marginally significant. Let's try testing the AICs between the two non-robust models, and a simple lmm (no random effects).

```{r BOUT AIC}
PECK.MOD0<-lm(PECK_RATE~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY, data = DATA)
anova(PECK.MOD1,PECK.MOD2,PECK.MOD0)
```

VERY WEIRD. Model 1 without the interactions is preferred over Model 2. However, the simple MOD0 is preferred. Let's see the outputs of that model.

```{r Robust PECK MODEL 0}
sjPlot::tab_model(PECK.MOD0
                  , show.re.var = T
                  , title = "Robust Peck Rate Model 0 Output"
                  , dv.labels = " Effects on peck rate")
```

Disturbance frequency is now significant! From this, I can infer that the marginally significant effect of disturbance frequency in the LMMs would probably become significant were we to have an increased sample size.

Let's plot these results

## The plot(s) (thicken)

### Disturbance Frequency

```{r PECK Disturb 1}
PK.DF.DOT1<-DATA %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE))+
  geom_point(size = 3) +
  theme_classic() +
  labs(y="Pecks per minute", x="Disturbances per Minute")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom")

PK.DF.DOT1
```

Ok, I see where some issues may be. There are a number of points at 0. In other words, they did not forage. I'll quickly fit another model, this time using individuals with time spent with head down > 0s.

```{r Robust PECK MODEL 4}
PECK.RMOD4<-rlmer(PECK_RATE~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = DATA[which(DATA$BEHAVIOR == "HD" & DATA$BEHAVIOR_DURATION > 0),])

sjPlot::tab_model(PECK.RMOD4
                  , show.re.var = T
                  , title = "Robust Peck Rate Model 4 Output"
                  , dv.labels = " Effects on peck rate")
```

Won't you look at that! When I remove points with time spent foraging is greater than 0, the disturbance frequency becomes significant. However, the variance explained by the random factor is 0.

Let's plot this again.

```{r PECK Disturb 2}
PK.DF.DOT2<-DATA[which(DATA$BEHAVIOR == "HD" & DATA$BEHAVIOR_DURATION > 0),] %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE)) +
  geom_point(size = 3) +
  geom_smooth(method="lm") +
  theme_classic() +
  labs(y="Pecks per minute", x="Disturbances per Minute")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom")

PK.DF.DOT2

```

Very nice! While it is not the prettiest of results, it roughly appears that as the disturbance frequency increases, so does the peck rate. I believe this can be explained by a sense of 'urgency' in higher-risk environments. Greater sample sizes would be required to make conclusive statements.

### Bait Presence

I would have expected there to have been an increase in peck rate in baited sites. To better visualize this, I'll make a dot plot including sentinel presence and environment, as well as peck rate.

```{r BOUT Dot Plot}
PECK.ALL.DOT<-DATA[which(DATA$BEHAVIOR == "HD" & DATA$BEHAVIOR_DURATION > 0),] %>%
  summarySE(measurevar = "PECK_RATE"
            , groupvars = c("SENTINEL_PRESENCE", "GENERALIZED_ENVIRONMENT", "BAIT_PRESENCE")) %>%
  ggplot(aes(x = SENTINEL_PRESENCE
                     , y = PECK_RATE
                     , color = BAIT_PRESENCE
                     , shape = BAIT_PRESENCE))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(PECK_RATE-se)
                    , ymax=(PECK_RATE+se))
                , width = 0.2
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Mean Peck Rate (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none"
                      , labels = c("Bait Absent"
                                   , "Bait Present")) +
  scale_shape_manual(values = c(16,17)
                     , labels = c("Bait Absent"
                                  , "Bait Present")
                     , name = "") +
  scale_x_discrete(labels = c("Sentinel Absent"
                                  , "Sentinel Present"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  facet_grid(~GENERALIZED_ENVIRONMENT)

PECK.ALL.DOT
```

So I have no data for non-baited sites in commercial areas with a sentinel being present. This is unfortunate.
It does seem like there are differences caused by the presence of bait, yet more sampling would have been required.

As for the other factors, there only seems to be a difference in green areas when a sentinel is absent. When bait is present, the peck rate increases. This could be explained as crows needing more time to find food in non-baited sites, yet the difference caused by bait is non-existent in the presence of a sentinel.

### 2023-08-18 Data & Scripts/PTWY.Rmd
Aug 18th
Knitted Scripts, SENT and PTWY done. PCA was archived.

@@
 ---
-```{r Libraries}
-library("lavaan")
-library("lavaanPlot")
+```{r Libraries, include = F}
+library("MANOVA.RM")
+library("dplyr")
+library("ggplot2")
+```
+
+```{r Colorblind-friendly palette, include = F}
+# The palette with black:
+cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
 ```
@@
 ```{r Open DATA.SR}
-PTWY<-read.csv("DATA.SR.csv", stringsAsFactors = T)
-str(PTWY)
-```
-```{r PTWY Analysis}
-PTWY.ENV <- '
-
-# measurement model
-
-ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TOTAL_FREQUENCY_OF_DISTURBANCES
-
-# intercepts
-
-CODED_ENV ~ 1
-CODED_BAIT ~ 1
-DECIMAL_TIME ~ 1
-TOTAL_FREQUENCY_OF_DISTURBANCES ~ 1
-'
-
-PTWY.Fit.ENV <- sem(PTWY.ENV, data = PTWY)
-
-summary(PTWY.Fit.ENV, fit.measures = TRUE, standardized = TRUE)
+PTWY.raw<-read.csv("DATA.SR.csv", stringsAsFactors = T)
+str(PTWY.raw)
 ```
 
-```{r PTWY.SOC Analysis}
-PTWY.Mod.SOC <- '
+Nice, it opened as it should. Let's subset the dataset to only contain what we want.
 
-# measurement model 
-SOCIAL =~ CODED_GROUP_SIZE
+```{r Subsetting PTWY}
+PTWY<- PTWY.raw %>%
+  dplyr::select(VIDEO_ID.,ID., DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_FREQUENCY_OF_DISTURBANCES, DISTURBANCE_FREQUENCY | starts_with("PTWY"))
 
-# intercepts
-CODED_GROUP_SIZE ~ 1
-'
-
-PTWY.Fit.SOC <- sem(PTWY.Mod.SOC, data = PTWY)
-
-summary(PTWY.Fit.SOC, fit.measures = TRUE, standardized = TRUE)
 ```
 
-```{r PTWY Analysis}
-PTWY.Mod <- '
+Next, let's group together the transitions to vigilance, and transitions to vulnerability.
 
-# measurement model
+```{r PTWY group}
+Vulnerable<-cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)
+Vigilant<-cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)
+# This does not work
 
-ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TEMPERATURE + TOTAL_FREQUENCY_OF_DISTURBANCES
+```
 
-VIGILANCE =~ PTWY_HDP.HU + PTWY_HD.HU + PTWY_M.HU
+Seems like it worked. Let's try fitting the model.
 
-FORAGING =~ PTWY_HD.HDP + PTWY_HU.HD + PTWY_M.HD
+## Vulnerable
 
-# regressions
+```{r MANOVA Vulnerable}
+PTWY.vuln<-MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = PTWY, subject = "ID.", iter=1000)
+summary(PTWY.vuln)
+```
 
-VIGILANCE ~ ENVIRONMENT + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE
+Interesting...
+While Sentinel presence and the interaction of sentinel presence and generalized environment had no significant effects, generalized environment had an effect on the frequency of transitions to vulnerability.
 
-FORAGING ~ ENVIRONMENT + VIGILANCE + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE
+**Figure out if you need to recalculate PTWY data to only include those 4 transitions. **
 
-'
+Let's perform a pairwise comparison.
 
-PTWY.Fit <- sem(PTWY.Mod, data = PTWY)
+### Post Hoc
 
-summary(PTWY.Fit, fit.measures = TRUE, standardized = TRUE)
+```{r PTWY CR Vuln}
+simCI(PTWY.vuln, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "GENERALIZED_ENVIRONMENT")
+
 ```
+
+HMMMM... Very weird... Now its not significant... Let's compute univariate comparisons.
+
+```{r PTWY Vulnerable univar}
+PTWY.HD.HDP<-MANOVA.wide(PTWY_HD.HDP~GENERALIZED_ENVIRONMENT, data = PTWY, iter = 1000)
+PTWY.HU.HD<-MANOVA.wide(PTWY_HU.HD~GENERALIZED_ENVIRONMENT, data = PTWY, iter = 1000)
+
+p.adjust(c(PTWY.HD.HDP$resampling[,1], PTWY.HU.HD$resampling[,1]), method = "bonferroni")
+#using the parametric bootstrap Wald-type statistic
+
+```
+
+Very cool!
+Turns out that only the transitions from head up to head down is significantly affected by the generalized environment.
+This is only if we use the Wald-type statistic and not the modified ANOVA-type statistic.
+
+### Plots
+
+```{r PTWY plots}
+source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
+G.ENV.PTWY<-PTWY %>%
+  summarySE(measurevar = "PTWY_HU.HD"
+            , groupvars = "GENERALIZED_ENVIRONMENT") %>%
+    ggplot(aes(x = GENERALIZED_ENVIRONMENT
+                     , y = PTWY_HU.HD
+                     , color = GENERALIZED_ENVIRONMENT))+
+  geom_point(position = position_dodge(width=0.9)
+             , size = 3) +
+  geom_errorbar(aes(ymin=(PTWY_HU.HD-se)
+                    , ymax=(PTWY_HU.HD+se))
+                , width = 0.1
+                , position = position_dodge(width=0.9))+
+  theme_classic() +
+  ylab("Head Up to Head Down (%)") +
+  scale_colour_manual(values = cbPalette
+                      , guide="none")+
+  theme(axis.title.x = element_blank()
+        , legend.position = "bottom"
+        , legend.box="vertical"
+        , legend.margin=margin())
+
+G.ENV.PTWY
+
+```
+
+## Vigilance
+
+```{r MANOVA Vigilance}
+PTWY.vigil<-MANOVA.wide(cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = PTWY, subject = "ID.", iter=1000)
+summary(PTWY.vigil)
+```
+
+The results of this test suggest that no transitions to vigilance are significantly affected by either the presence of a sentinel or the generalized environment.
+
+**Figure out what to do with factors such as bait presence, group size and disturbance frequency**
+
+<!-- # Old code below -->
+<!-- # ```{r PTWY Analysis} -->
+<!-- # PTWY.ENV <- '  -->
+<!-- #  -->
+<!-- # # measurement model  -->
+<!-- #  -->
+<!-- # ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TOTAL_FREQUENCY_OF_DISTURBANCES  -->
+<!-- #  -->
+<!-- # # intercepts  -->
+<!-- #  -->
+<!-- # CODED_ENV ~ 1 -->
+<!-- # CODED_BAIT ~ 1 -->
+<!-- # DECIMAL_TIME ~ 1 -->
+<!-- # TOTAL_FREQUENCY_OF_DISTURBANCES ~ 1 -->
+<!-- # '  -->
+<!-- #  -->
+<!-- # PTWY.Fit.ENV <- sem(PTWY.ENV, data = PTWY)  -->
+<!-- #  -->
+<!-- # summary(PTWY.Fit.ENV, fit.measures = TRUE, standardized = TRUE) -->
+<!-- # ``` -->
+<!-- #  -->
+<!-- # ```{r PTWY.SOC Analysis} -->
+<!-- # PTWY.Mod.SOC <- '  -->
+<!-- #  -->
+<!-- # # measurement model  -->
+<!-- # SOCIAL =~ CODED_GROUP_SIZE  -->
+<!-- #  -->
+<!-- # # intercepts -->
+<!-- # CODED_GROUP_SIZE ~ 1 -->
+<!-- # '  -->
+<!-- #  -->
+<!-- # PTWY.Fit.SOC <- sem(PTWY.Mod.SOC, data = PTWY)  -->
+<!-- #  -->
+<!-- # summary(PTWY.Fit.SOC, fit.measures = TRUE, standardized = TRUE) -->
+<!-- # ``` -->
+<!-- #  -->
+<!-- # ```{r PTWY Analysis} -->
+<!-- # PTWY.Mod <- '  -->
+<!-- #  -->
+<!-- # # measurement model  -->
+<!-- #  -->
+<!-- # ENVIRONMENT =~ CODED_ENV + CODED_BAIT + DECIMAL_TIME + TEMPERATURE + TOTAL_FREQUENCY_OF_DISTURBANCES  -->
+<!-- #  -->
+<!-- # VIGILANCE =~ PTWY_HDP.HU + PTWY_HD.HU + PTWY_M.HU  -->
+<!-- #  -->
+<!-- # FORAGING =~ PTWY_HD.HDP + PTWY_HU.HD + PTWY_M.HD  -->
+<!-- #  -->
+<!-- #   -->
+<!-- #  -->
+<!-- # # regressions  -->
+<!-- #  -->
+<!-- # VIGILANCE ~ ENVIRONMENT + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE -->
+<!-- #  -->
+<!-- # FORAGING ~ ENVIRONMENT + VIGILANCE + CODED_GROUP_SIZE + CODED_SENTINEL_PRESENCE -->
+<!-- #  -->
+<!-- # '  -->
+<!-- #  -->
+<!-- # PTWY.Fit <- sem(PTWY.Mod, data = PTWY)  -->
+<!-- #  -->
+<!-- # summary(PTWY.Fit, fit.measures = TRUE, standardized = TRUE) -->
+<!-- # ``` -->

### 2023-08-18 Data & Scripts/SENT.Rmd
Aug 18th
Knitted Scripts, SENT and PTWY done. PCA was archived.

@@
 ```{r SENT Subset}
-SENT<-SENT[1:81,c(1,2,7,10,11,15,17,22,30,31)]
+SENT<-SENT[1:81,c(1,2,7,10,11,15,17,20,22,30,31)]
 str(SENT)
@@
 
-## Fit the model
+## Chi Square Time
 
-```{r SENT model - Simple}
-SENT.mod1<-glm(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+GROUP_SIZE+TOTAL_FREQUENCY_OF_DISTURBANCES+DECIMAL_TIME+TEMPERATURE
-              , family = binomial()
-              , data = SENT)
+```{r Chi Square}
+library(descr)
+library(MASS)
+CHISQ<-lapply(SENT[,c("GENERALIZED_ENVIRONMENT"
+                                 , "GROUP_SIZE"
+                                 , "DISTURBANCE_FREQUENCY")
+                              ]
+                         , function(x) CrossTable(x,SENT$SENTINEL_PRESENCE
+                                                  , fisher = T
+                                                  , chisq = T
+                                                  , expected = T
+                                                  , prop.c = F
+                                                  , prop.t = F
+                                                  , prop.chisq = F
+                                                  , sresid = T
+                                                  , format = 'SPSS')
+                         )
 
-summary(SENT.mod1)
+CHISQ
 ```
 
-OOF. By removing the duplicates, we now have no significant effects of any variable.
-Next, let's try and see if we can run a glmm and include the video ID as a random effect.
+Hmm... The results of the chi square test suggest that none of the factors, either generalized environment, group size, or disturbance frequency have a significant effect.
 
-```{r SENT GLMM}
-SENT.mod2<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+TOTAL_FREQUENCY_OF_DISTURBANCES+DECIMAL_TIME+TEMPERATURE+(1|VIDEO_ID.)
-                , family = binomial()       
-                , data = SENT
-                       )
+# The continuous variables
 
-summary(SENT.mod2)
+To look more closely at the effects of continuous variables on the presence of a sentinel, I propose to plot the following figures:
+- Count of sentinel presence across time
+- Count of sentinel presence across temperature
+- Count of sentinel presence across disturbance frequency
+- Count of sentinel presence across the number of crows recorded
+
+```{r SENT plots}
+SENT %>% 
+  rename("Temperature (deg. C)" = "TEMPERATURE"
+         , "Decimal Time" = "DECIMAL_TIME"
+         , "Disturbance Frequency (per min)" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
+         , "Number of Crows" = "NUMBER_OF_CROWS_RECORDED") %>%
+  pivot_longer(., cols = c("Temperature (deg. C)"
+                           , "Decimal Time"
+                           , "Number of Crows"
+                           , "Disturbance Frequency (per min)")
+               , names_to = "Var"
+               , values_to = "Val") %>%
+  ggplot(aes(
+    x = Val
+    , colour = SENTINEL_PRESENCE))+
+  geom_freqpoly(linewidth = 1
+                , alpha=I(.6)) +
+  scale_colour_manual(values = cbPalette
+                      , name = "Sentinel Presence") +
+  facet_wrap(~Var, scale = "free_x") +
+  theme_classic() +
+  theme(legend.position = "bottom"
+        , axis.title.x = element_blank())
+
 ```
-Ok, the model is singular. Let's see if we can improve the model by reducing the number of factors.
 
-```{r AIC testing}
-drop1(SENT.mod2, tests = "F")
-```
-From the results of the drop1 function, we see that the removal of temperature and decimal time would improve the model fit. Let's do that next.
+OK!
+There do not appear to be any patterns suggesting that the likelihood of a sentinel being present is affected by either the temperature, time, number of foraging crows or frequency of disturbances.
+In fact, the presence or absence of a sentinel does not appear to follow different patterns at all, in many cases having the same or similar peaks across the continuous variables.
 
-```{r SENT GLMM2}
-SENT.mod3<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+TOTAL_FREQUENCY_OF_DISTURBANCES+(1|VIDEO_ID.)
-                , family = binomial()       
-                , data = SENT
-                       )
+# Results
 
-summary(SENT.mod3)
-```
-Still singular. It seems like my sample size is too small to determine any significant effects. Let's see if we can remove additional variables.
+The likelihood of a sentinel being present is not significantly affected by either the generalized environment, the binned group size, nor the binned disturbance frequency. 
+When we looked at the effects of decimal time, number of foraging crows, temperature, and disturbance frequency (nb.
+disturb/min), we observed strikingly similar patterns between the number of observations with and without a sentinel.
+This suggests that neither variable affects the likelihood of a sentinel being present.
 
-```{r AIC testing2}
-drop1(SENT.mod3, tests = "F")
-```
-Removing Generalized Environment would improve the model. Let's fit the final model.
+<!-- # Old glm stuff below -->
+<!-- # ```{r SENT model - Simple} -->
+<!-- # SENT.mod1<-glm(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+GROUP_SIZE+TOTAL_FREQUENCY_OF_DISTURBANCES+DECIMAL_TIME+TEMPERATURE -->
+<!-- #               , family = binomial() -->
+<!-- #               , data = SENT) -->
+<!-- #  -->
+<!-- # summary(SENT.mod1) -->
+<!-- # ``` -->
 
-```{r SENT GLMM3}
-SENT.mod4<-glmer(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES+(1|TEMPERATURE)
-                , family = binomial()       
-                , data = SENT
-                       )
+<!-- OOF. By removing the duplicates, we now have no significant effects of any variable. -->
+<!-- Next, let's try and see if we can run a glmm and include the video ID as a random effect. -->
 
-summary(SENT.mod4)
-```
-Bummer...
+<!-- ```{r SENT GLMM} -->
+<!-- SENT.mod2<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+TOTAL_FREQUENCY_OF_DISTURBANCES+DECIMAL_TIME+TEMPERATURE+(1|VIDEO_ID.) -->
+<!--                 , family = binomial()        -->
+<!--                 , data = SENT -->
+<!--                        ) -->
+
+<!-- summary(SENT.mod2) -->
+<!-- ``` -->
+<!-- Ok, the model is singular. Let's see if we can improve the model by reducing the number of factors. -->
+
+<!-- ```{r AIC testing} -->
+<!-- drop1(SENT.mod2, tests = "F") -->
+<!-- ``` -->
+<!-- From the results of the drop1 function, we see that the removal of temperature and decimal time would improve the model fit. Let's do that next. -->
+
+<!-- ```{r SENT GLMM2} -->
+<!-- SENT.mod3<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+TOTAL_FREQUENCY_OF_DISTURBANCES+(1|VIDEO_ID.) -->
+<!--                 , family = binomial()        -->
+<!--                 , data = SENT -->
+<!--                        ) -->
+
+<!-- summary(SENT.mod3) -->
+<!-- ``` -->
+<!-- Still singular. It seems like my sample size is too small to determine any significant effects. Let's see if we can remove additional variables. -->
+
+<!-- ```{r AIC testing2} -->
+<!-- drop1(SENT.mod3, tests = "F") -->
+<!-- ``` -->
+<!-- Removing Generalized Environment would improve the model. Let's fit the final model. -->
+
+<!-- ```{r SENT GLMM3} -->
+<!-- SENT.mod4<-glmer(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES+(1|TEMPERATURE) -->
+<!--                 , family = binomial()        -->
+<!--                 , data = SENT -->
+<!--                        ) -->
+
+<!-- summary(SENT.mod4) -->
+<!-- ``` -->
+<!-- Bummer...
 Still singular.
@@
 Next, I will run another 'simple' glm.
+-->
 
-```{r SENT model - Simple 2}
-SENT.mod5<-glm(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES
-              , family = binomial()
-              , data = SENT)
+<!-- ```{r SENT model - Simple 2} -->
+<!-- SENT.mod5<-glm(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES -->
+<!--               , family = binomial() -->
+<!--               , data = SENT) -->
 
-summary(SENT.mod5)
-```
+<!-- summary(SENT.mod5) -->
+<!-- ``` -->
 
-Same results as the glmm. Let's quickly try code I found on R-bloggers (https://www.r-bloggers.com/2015/08/bootstrap-with-logistic-regression/)
+<!-- Same results as the glmm. Let's quickly try code I found on R-bloggers (https://www.r-bloggers.com/2015/08/bootstrap-with-logistic-regression/) -->
 
-```{r Bootstrapped model}
-library(boot)
-logit_test <- function(d,indices) {  
-d <- d[indices,]  
-fit <- glmer(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES+(1|VIDEO_ID.)
-                , family = binomial()       
-                , data = d
-                       ) 
-return(coef(fit))  
-}
-boot_fit <- boot(  
-   data = SENT, 
-   statistic = logit_test, 
-   R = 1e5
-) 
-```
+<!-- ```{r Bootstrapped model} -->
+<!-- library(boot) -->
+<!-- logit_test <- function(d,indices) {   -->
+<!-- d <- d[indices,]   -->
+<!-- fit <- glmer(SENTINEL_PRESENCE~TOTAL_FREQUENCY_OF_DISTURBANCES+(1|VIDEO_ID.) -->
+<!--                 , family = binomial()        -->
+<!--                 , data = d -->
+<!--                        )  -->
+<!-- return(coef(fit))   -->
+<!-- } -->
+<!-- boot_fit <- boot(   -->
+<!--    data = SENT,  -->
+<!--    statistic = logit_test,  -->
+<!--    R = 1e5 -->
+<!-- )  -->
+<!-- ``` -->
 
-Ok, still singular and the bootstrap failed. It is possible that there are too many levels in the frequency of disturbances. Let's try to simplify by binning the disturbance frequencies.
+<!-- Ok, still singular and the bootstrap failed. It is possible that there are too many levels in the frequency of disturbances. Let's try to simplify by binning the disturbance frequencies. -->
 
-```{r SENT GLMM 4}
-SENT.mod6<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+DECIMAL_TIME+TEMPERATURE+(1|VIDEO_ID.)
-                , family = binomial()       
-                , data = SENT
-                       )
+<!-- ```{r SENT GLMM 4} -->
+<!-- SENT.mod6<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+DECIMAL_TIME+TEMPERATURE+(1|VIDEO_ID.) -->
+<!--                 , family = binomial()        -->
+<!--                 , data = SENT -->
+<!--                        ) -->
 
-summary(SENT.mod6)
-```
+<!-- summary(SENT.mod6) -->
+<!-- ``` -->
 
-Still singular. Let's remove some factors using drop 1.
+<!-- Still singular. Let's remove some factors using drop 1. -->
 
-```{r AIC testing3}
-drop1(SENT.mod6, tests = "F")
-```
+<!-- ```{r AIC testing3} -->
+<!-- drop1(SENT.mod6, tests = "F") -->
+<!-- ``` -->
 
-We can remove TEMPERATURE.
+<!-- We can remove TEMPERATURE. -->
 
-```{r SENT GLMM 5}
-SENT.mod7<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+DECIMAL_TIME+(1|VIDEO_ID.)
-                , family = binomial()       
-                , data = SENT
-                       )
+<!-- ```{r SENT GLMM 5} -->
+<!-- SENT.mod7<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+DECIMAL_TIME+(1|VIDEO_ID.) -->
+<!--                 , family = binomial()        -->
+<!--                 , data = SENT -->
+<!--                        ) -->
 
-summary(SENT.mod7)
-drop1(SENT.mod7) #remove decimal time
-```
+<!-- summary(SENT.mod7) -->
+<!-- drop1(SENT.mod7) #remove decimal time -->
+<!-- ``` -->
 
-```{r SENT GLMM 6}
-SENT.mod8<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+(1|VIDEO_ID.)
-                , family = binomial()       
-                , data = SENT
-                       )
+<!-- ```{r SENT GLMM 6} -->
+<!-- SENT.mod8<-glmer(SENTINEL_PRESENCE~GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+(1|VIDEO_ID.) -->
+<!--                 , family = binomial()        -->
+<!--                 , data = SENT -->
+<!--                        ) -->
 
-summary(SENT.mod8)
-drop1(SENT.mod8) #remove environment
-```
+<!-- summary(SENT.mod8) -->
+<!-- drop1(SENT.mod8) #remove environment -->
+<!-- ``` -->
 
-```{r SENT GLMM 7}
-SENT.mod9<-glmer(SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY+(1|VIDEO_ID.)
-                , family = binomial()       
-                , data = SENT
-                       )
+<!-- ```{r SENT GLMM 7} -->
+<!-- SENT.mod9<-glmer(SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY+(1|VIDEO_ID.) -->
+<!--                 , family = binomial()        -->
+<!--                 , data = SENT -->
+<!--                        ) -->
 
-summary(SENT.mod9)
-drop1(SENT.mod9)
-```
+<!-- summary(SENT.mod9) -->
+<!-- drop1(SENT.mod9) -->
+<!-- ``` -->
 
-Still singular. Let's try removing the random effect and running yet another simple model.
+<!-- Still singular. Let's try removing the random effect and running yet another simple model. -->
 
-```{r SENT model - Simple 3}
-SENT.mod10<-glm(SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY
-              , family = binomial()
-              , data = SENT)
+<!-- ```{r SENT model - Simple 3} -->
+<!-- SENT.mod10<-glm(SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY -->
+<!--               , family = binomial() -->
+<!--               , data = SENT) -->
 
-summary(SENT.mod10)
-```
+<!-- summary(SENT.mod10) -->
+<!-- ``` -->
 
-Disturbance frequency is not significant. What about another bootstrap?
+<!-- Disturbance frequency is not significant. What about another bootstrap? -->
 
-```{r Bootstrapped model 2}
-library(boot)
-logit_test <- function(d,indices, formula) {  
-d <- d[indices,]  
-fit <- glm(formula
-           , family = binomial()
-           , data = d
-           ) 
-return(coef(fit))  
-}
+<!-- ```{r Bootstrapped model 2} -->
+<!-- library(boot) -->
+<!-- logit_test <- function(d,indices, formula) {   -->
+<!-- d <- d[indices,]   -->
+<!-- fit <- glm(formula -->
+<!--            , family = binomial() -->
+<!--            , data = d -->
+<!--            )  -->
+<!-- return(coef(fit))   -->
+<!-- } -->
 
-boot_fit <- boot(
-   data = SENT
-   , statistic = logit_test
-   , R = 1000
-   , formula=SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY
-   )
+<!-- boot_fit <- boot( -->
+<!--    data = SENT -->
+<!--    , statistic = logit_test -->
+<!--    , R = 1000 -->
+<!--    , formula=SENTINEL_PRESENCE~DISTURBANCE_FREQUENCY -->
+<!--    ) -->
 
-```
+<!-- ``` -->
 
-RUN A CHI SQUARED!!!
+<!-- RUN A CHI SQUARED!!! -->

### 2023-08-23 Data & Scripts/BOUT.Rmd
Aug 23

@@
 
-As random effects, I've chosen the video and individual ID
+As random effects, I've chosen the individual ID nested within Video ID.
 
@@
 ```{r BOUT MODEL 2}
-BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
+BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)
 

### 2023-08-23 Data & Scripts/PTWY.Rmd
Aug 23

@@
 
+## The other variables
+
+We still have bait presence, group size and disturbance frequency to look at. I want to try and make a quick function to run all the tests.
+
+```{r PTWY Vulnerable function}
+PTWY.vuln.all<-lapply(PTWY[,c("SENTINEL_PRESENCE"
+                              ,"GENERALIZED_ENVIRONMENT"
+                              , "GROUP_SIZE"
+                              , "BAIT_PRESENCE"
+                              , "DISTURBANCE_FREQUENCY")
+                              ]
+                         , function(x) MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~x, data = PTWY, iter = 1000)
+                         )
+PTWY.SENT<-summary(PTWY.vuln.all$SENTINEL_PRESENCE)
+PTWY.GENV<-summary(PTWY.vuln.all$GENERALIZED_ENVIRONMENT)
+PTWY.GS<-summary(PTWY.vuln.all$GROUP_SIZE)
+PTWY.BP<-summary(PTWY.vuln.all$BAIT_PRESENCE)
+PTWY.DF<-summary(PTWY.vuln.all$DISTURBANCE_FREQUENCY)
+```
+
 <!-- # Old code below -->

### 2023-08-23 Data & Scripts/SENT.Rmd
Aug 23

@@
 
-We will be reusing the 'Single Row' dataset for this script.
+We will be reusing the 'Single Row' data set for this script.
 
@@
 ```{r SENT Subset}
-SENT<-SENT[1:81,c(1,2,7,10,11,15,17,20,22,30,31)]
+SENT<-SENT[1:81,c(1,2,7,10,11,15,17,19,20,22,30,31)]
 str(SENT)
@@
                       , "GROUP_SIZE"
-                                 , "DISTURBANCE_FREQUENCY")
+                      , "DISTURBANCE_FREQUENCY"
+                      , "BAIT_PRESENCE")
                               ]

### 2023-08-25 Data & Scripts/BOUT.Rmd
Aug. 25

--- 2023-08-17 Data & Scripts/BOUT.Rmd
@@
 
+```{r Model Labels}
+pl<-c(
+  '(Intercept)' = "Intercept"
+  , BEHAVIORHU = "Behavior"
+  , SENTINEL_PRESENCEYES = "Sentinel Presence"
+  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
+  , GROUP_SIZESMALL = "Group Size"
+  , BAIT_PRESENCEYES = "Bait Presence"
+  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
+  , 'BEHAVIORHU:SENTINEL_PRESENCEYES' = "Behavior: Sentinel Presence"
+  , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
+  , 'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
+  , 'BEHAVIORHU:SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Sentinel Presence: Generalized Environment"
+)
+```
+
 Let's fit a model.
@@
 sjPlot::tab_model(BOUT.MOD1
+                  , pred.labels = pl
+                  , rm.terms = c("BEHAVIORM"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES"
+                                 , "BEHAVIORM:GENERALIZED_ENVIRONMENTGreen Area"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area")
                   , show.re.var = T
@@
 sjPlot::tab_model(BOUT.MOD2
+                  , pred.labels = pl
+                  , rm.terms = c("BEHAVIORM"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES"
+                                 , "BEHAVIORM:GENERALIZED_ENVIRONMENTGreen Area"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area")
                   , show.re.var = T
@@
 The interaction between behavior and group size is significant.
-Lastly, moving and bait presence were significant.
+Lastly, moving and bait presence were significant (removed Moving from output).
 
@@
 sjPlot::tab_model(BOUT.MOD3
+                  , pred.labels = pl
+                  , rm.terms = c("BEHAVIORM"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES"
+                                 , "BEHAVIORM:GENERALIZED_ENVIRONMENTGreen Area"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area")
                   , show.re.var = T
@@
 sjPlot::tab_model(BOUT.RMOD2
+                  , pred.labels = pl
+                  , rm.terms = c("BEHAVIORM"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES"
+                                 , "BEHAVIORM:GENERALIZED_ENVIRONMENTGreen Area"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area")
                   , show.re.var = T

### 2023-09-22 Data & Scripts/GRaD Presentation.Rmd
Sept. 22
Presentation material

---
title: "GRaD Presentation"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
---

# Bookkeeping

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
library("formattable")
library("dplyr")
library("descr")
library("MASS")
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

```{r Model Labels}
labs.prop<-c(
  '(Intercept)' = "Intercept"
  , BEHAVIORHU = "Behavior"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  )

labs.bout.1<-c(
  '(Intercept)' = "Intercept"
  , BEHAVIORHU = "Behavior"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
)

labs.bout.2<-c(
  '(Intercept)' = "Intercept"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  )

labs.peck<-c(
  '(Intercept)' = "Intercept"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'GENERALIZED_ENVIRONMENTGreen Area:DISTURBANCE_FREQUENCY' = "Generalized Environment: Disturbance Frequency"
  )
```

```{r DATA.SR & BOUT}
DATA.SR<-read.csv("DATA.SR.csv", stringsAsFactors = T) %>%
  rename("VIDEO_ID" = "VIDEO_ID."
         , "ID" = "ID.")
BOUT.raw<-read.csv("BOUT.csv", stringsAsFactors = T)
```

# Proportion Data

```{r Open PROP}
PROP<-DATA.SR[,c(1,2,15,17,19,22,30,36,41,46)]%>% 
  subset(.
         , HU_BEHAVIOR_PROPORTION_... != 0 
         & HD_BEHAVIOR_PROPORTION_... != 0 
         & M_BEHAVIOR_PROPORTION_... != 0
         ) %>% #Remove cases where proportion = 1 or 0
  rename("HU" = "HU_BEHAVIOR_PROPORTION_..."
         , "HD" = "HD_BEHAVIOR_PROPORTION_..."
         , "M" = "M_BEHAVIOR_PROPORTION_..."
         , "DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
         ) %>%
  pivot_longer(., cols = c("HU"
                              , "HD"
                              , "M")
               , names_to = "BEHAVIOR"
               , values_to = "PROPORTION")
PROP$ASIN_PROPORTION<-asin(sqrt(PROP$PROPORTION)) #Transform for normality

str(PROP)
```

```{r PROP Summary}
PROP.SUMMARY <- summarySE(data = PROP
                      , measurevar = "PROPORTION"
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "BEHAVIOR"
                                      )
                      )
PROP.SUMMARY
```

## Stacked barplot

```{r PROP Barplot}
PROP.BARPLOT<-ggplot(PROP.SUMMARY
                 , aes(x = GENERALIZED_ENVIRONMENT
                       , y = PROPORTION
                       , fill = BEHAVIOR))+
  geom_bar(stat = 'identity'
           , position = 'stack')+
  geom_text(aes(label = paste0(formattable::digits(PROPORTION*100, dig=2)
                               , "%"))
            , position = position_stack(vjust = 0.5)
            , size = 4) +
  scale_y_continuous(labels = scales::percent) +
  theme_classic()+
  ylab("Proportion of time")+
  scale_x_discrete(labels=c("Commercial Area"
                            , "Green Area"))+
  scale_fill_manual(values = cbPalette, labels = c("Foraging"
                                                   , "Alert"
                                                   , "Moving")
                    , name="")+
  theme(legend.position = "bottom"
        , text = element_text(size = 18)
        , axis.title.x = element_blank())
PROP.BARPLOT
```

## PROP Model

```{r PROP Model}
PROP.MOD<-lm(ASIN_PROPORTION~BEHAVIOR+GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+BAIT_PRESENCE+GROUP_SIZE, data = PROP)

sjPlot::tab_model(PROP.MOD
                  , pred.labels = labs.prop
                  , rm.terms = "BEHAVIORM"
                  , show.re.var = T
                  , title = "Proportion Model 1 Output"
                  , dv.labels = " Effects of on the proportion of behaviors")
```

Cool, nothing is significant (apart from "Moving" being different, but thats to be expected).

Crows allocate similar time to foraging and vigilance.

I tinkered with the models, and the model with the lowest AIC value had no interactions and no random effects.

# Bout Data

```{r Open BOUT}
BOUT<-BOUT.raw %>%
  filter(.
         , DURATION > 0.01) %>% #Remove impossibly small values
  filter(.
         , BEHAVIOR != "M") #Remove "M"
BOUT$LDURATION<-log(BOUT$DURATION) #Transform data for normality
str(BOUT)
str(BOUT.raw)
```

```{r BOUT Summary}
BOUT.SUMMARY <- summarySE(data = BOUT
                      , measurevar = "DURATION"
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "BEHAVIOR"
                                      )
                      ) %>%
  mutate(BEHAVIOR = recode_factor(BEHAVIOR
                                  ,'HD' = 'Foraging'#Rename HD as 'foraging'
                                  , 'HU' = 'Alert') #Rename HU as 'Alert'
                           )

BOUT.SUMMARY
```

## BOUT Dot Plot

```{r BOUT Dot Plot}
BOUT.DOTPLOT<-BOUT.SUMMARY %>%
           ggplot(.
               , aes(x = GENERALIZED_ENVIRONMENT
                     , y = DURATION
                     , colour = GENERALIZED_ENVIRONMENT))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  geom_text(aes(label = round(DURATION, 2)), hjust = -0.2, size = 6) +
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none") +
  scale_x_discrete(labels = c("Commercial Area", "Green Area"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18)) +
  facet_grid(~BEHAVIOR)

BOUT.DOTPLOT
```

> The error bars represent the standard error, while the symbols are the mean duration of bouts.

## BOUT Model - All behaviors

```{r BOUT Model - All behaviors}
BOUT.MOD<-rlmer(LDURATION~BEHAVIOR*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)

sjPlot::tab_model(BOUT.MOD
                  , pred.labels = labs.bout.1
                  , show.re.var = T
                  , title = "Bout Robust Model  Output"
                  , dv.labels = " Effects on duration of bouts of all behaviors")

sjPlot::plot_model(BOUT.MOD
                  , axis.labels = labs.bout.1
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on duration of bouts of all behaviors"
                  )
```

##BOUT model - Head Down

```{r BOUT Model - Head Down}
BOUT.MOD.HD<-rlmer(LDURATION~GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = subset(BOUT, BEHAVIOR == "HD"))

sjPlot::tab_model(BOUT.MOD.HD
                  , pred.labels = labs.bout.2
                  , show.re.var = T
                  , title = "Head Down"
                  , dv.labels = " Effects on foraging bout duration")

sjPlot::plot_model(BOUT.MOD.HD
                  , axis.labels = labs.bout.2
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on foraging bout duration")

```

```{r BOUT Model - Head Up}
BOUT.MOD.HU<-rlmer(LDURATION~GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = subset(BOUT, BEHAVIOR == "HU"))

sjPlot::tab_model(BOUT.MOD.HU
                  , pred.labels = labs.bout.2
                  , show.re.var = T
                  , title = "Head Up"
                  , dv.labels = " Effects on alert bout duration")

sjPlot::plot_model(BOUT.MOD.HU
                  , axis.labels = labs.bout.2
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on alert bout duration")

```

##BOUT Post Hoc

```{r BOUT Post Hoc}
BOUT.DIFF<-emmeans(BOUT.MOD, ~BEHAVIOR+GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
test(pairs(BOUT.DIFF, by= "BEHAVIOR"), adjust="fdr")
```

#Peck Rate

```{r Open PECK}
PECK<-DATA.SR[,c(1,2,15,17,19,22,30,38,48)] %>%
  subset(., HD_BEHAVIOR_DURATION > 0) %>%
  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES")

str(PECK)
```
Next, we'll compile the means of the peck rate by sentinel presence and generalized environment.

```{r PECK Summary, echo = F}
PECK.SUMMARY<-summarySE(PECK, "PECK_RATE"
                     , "GENERALIZED_ENVIRONMENT"
                     , na.rm=T)
PECK.SUMMARY
```

## PECK Dot Plot

```{r PECK Dot Plot}
PECK.DOTPLOT<-ggplot(data = PECK.SUMMARY
               , aes(x = GENERALIZED_ENVIRONMENT
                     , y = PECK_RATE
                     , color = GENERALIZED_ENVIRONMENT))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(PECK_RATE-se)
                    , ymax=(PECK_RATE+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  geom_text(aes(label = round(PECK_RATE, 2)), hjust = -0.2, size = 6) +
  theme_classic() +
  ylab("Mean Peck Rate (per min)") +
  scale_colour_manual(values = cbPalette
                      , guide="none") +
  scale_x_discrete(labels = c("Commercial Area"
                                  , "Green Area")) +
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18))

PECK.DOTPLOT
```

## PECK Model

```{r PECK Model}
PECK.MOD<-rlmer(PECK_RATE~GENERALIZED_ENVIRONMENT*DISTURBANCE_FREQUENCY+GROUP_SIZE+BAIT_PRESENCE+(1|VIDEO_ID/ID), data = PECK)

sjPlot::tab_model(PECK.MOD
                  , pred.labels = labs.peck
                  , show.re.var = T
                  , title = "Robust Peck Rate Model Output"
                  , dv.labels = " Effects on peck rate")
sjPlot::plot_model(PECK.MOD
                  , axis.labels = labs.peck
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on peck rate" )
```

## Disturbance Frequency x Generalized Environment

```{r PECK Disturbance plot}
PECK.DISTURBANCE.DOTPLOT<-PECK %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE
               , color = GENERALIZED_ENVIRONMENT))+
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  theme_classic() +
  labs(y="Peck Rate (per min)", x="Disturbance Frequency (per min)")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom"
        , text = element_text(size = 18))+
  scale_color_manual(values = cbPalette
                    , labels = c("Commercial Area"
                                 , "Green Area")
                    , name="")

PECK.DISTURBANCE.DOTPLOT
```

# Sentinel Likelihood

```{r open SENT}
SENT<-DATA.SR[1:81,c(1,7,10,15,17,19,20,22,30,31)] %>%
  distinct()
str(SENT)
```

```{r SENT Summary}
SENT.SUMMARY <- summarySE(data = SENT
                      , measurevar = "GROUP_SIZE" #Random
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "SENTINEL_PRESENCE"
                                      )
                      )

SENT.SUMMARY
```

##SENT Barplot

```{r SENT Barplot}
SENT.BARPLOT <- ggplot(SENT.SUMMARY, aes(x=GENERALIZED_ENVIRONMENT
                 , fill = SENTINEL_PRESENCE
                 , y = N
                 , group = SENTINEL_PRESENCE)) + 
  geom_bar(stat='identity', width=.5, position = "dodge") +
  geom_text(aes(label = N, group = SENTINEL_PRESENCE), position = position_dodge(width = 0.5), vjust = 1.5, size = 8) +
  ylab("Number of observations") +
  theme_classic() +
  scale_fill_manual(""
                    , values = cbPalette
                    , labels = c("Sentinel Absent", "Sentinel Present")
                      ) +
  theme(axis.title.x = element_blank()
        , text = element_text(size = 18)
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())

SENT.BARPLOT
```

##SENT Chi-Squared

```{r SENT Chi-Squared}
SENT.CHISQ<-lapply(SENT[,c("GENERALIZED_ENVIRONMENT"
                      , "GROUP_SIZE"
                      , "DISTURBANCE_FREQUENCY"
                      )
                              ]
                         , function(x) CrossTable(x,SENT$SENTINEL_PRESENCE
                                                  , fisher = T
                                                  , chisq = T
                                                  , expected = T
                                                  , prop.c = F
                                                  , prop.t = F
                                                  , prop.chisq = F
                                                  , sresid = T
                                                  , format = 'SPSS')
                         )

SENT.CHISQ
```

##SENT Plots - Continuous Variables

```{r SENT Frequency plots}
SENT.FREQ <- SENT %>% 
  rename("Temperature (deg. C)" = "TEMPERATURE"
         , "Decimal Time" = "DECIMAL_TIME"
         , "Disturbance Frequency (per min)" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
         , "Number of Crows" = "NUMBER_OF_CROWS_RECORDED") %>%
  pivot_longer(., cols = c(#"Temperature (deg. C)"
                           # , "Decimal Time"
                           "Number of Crows"
                           , "Disturbance Frequency (per min)")
               , names_to = "Var"
               , values_to = "Val") %>%
  ggplot(aes(
    x = Val
    , colour = SENTINEL_PRESENCE))+
  geom_freqpoly(linewidth = 1
                , alpha=I(.6)) +
  scale_colour_manual(values = cbPalette
                      , name = ""
                      , labels = c("Sentinel Absent", "Sentinel Present")) +
  facet_wrap(~Var, scale = "free_x") +
  theme_classic() +
  theme(legend.position = "bottom"
        , axis.title.x = element_blank()
        , text = element_text(size = 18))

SENT.FREQ
```

### 2023-09-22 Data & Scripts/SENT.Rmd
Sept. 22
Presentation material

@@
 library("emmeans")
+library("descr")
 ```
@@
 ```{r Chi Square}
-library(descr)
-library(MASS)
+library("descr")
+library("MASS")
 CHISQ<-lapply(SENT[,c("GENERALIZED_ENVIRONMENT"
@@
 
+```{r Sent Barplot}
+ggplot(SENT, aes(x=GENERALIZED_ENVIRONMENT
+                 , fill = SENTINEL_PRESENCE)) + 
+  geom_bar(width=.5, position = "dodge") +
+  theme_classic() +
+  xlab("Generalized Environment") +
+  scale_fill_manual("Sentinel Presence"
+                      , values = cbPalette
+                      ) +
+  theme(legend.position = "bottom"
+        , legend.box="vertical"
+        , legend.margin=margin())
+
+```
+
 Hmm... The results of the chi square test suggest that none of the factors, either generalized environment, group size, or disturbance frequency have a significant effect.
@@
          , "Number of Crows" = "NUMBER_OF_CROWS_RECORDED") %>%
-  pivot_longer(., cols = c("Temperature (deg. C)"
-                           , "Decimal Time"
+  pivot_longer(., cols = c(#"Temperature (deg. C)"
+                           # , "Decimal Time"
                            , "Number of Crows"

### 2023-10-20 Data & Scripts/Alex Popescu - Final Stats.Rmd
Final Stats

---
title: "Alex Popescu - Final Stats"
author: "Alex Popescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---

# Bookkeeping

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
library("formattable")
library("dplyr")
library("descr")
library("MASS")
library("jmv")
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
```

```{r Colorblind-friendly palette, include = F}
# The palette with black:
cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7", "#009E73", "#F0E442", "#0072B2", "#D55E00")
```

```{r Labels}
labs.prop<-c(
  '(Intercept)' = "Intercept"
  , SENTINEL_PRESENCEYES = "Sentinel Presence"
  , BEHAVIORHU = "Behavior"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  )

labs.bout.1<-c(
  '(Intercept)' = "Intercept"
  , SENTINEL_PRESENCEYES = "Sentinel Presence"
  , BEHAVIORHU = "Behavior"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
  , 'BEHAVIORHU:SENTINEL_PRESENCEYES' = "Behavior: Sentinel Presence"
  , 'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
)

labs.bout.2<-c(
  '(Intercept)' = "Intercept"
  , SENTINEL_PRESENCEYES = "Sentinel Presence"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  ,'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
  )

labs.peck<-c(
  '(Intercept)' = "Intercept"
  , SENTINEL_PRESENCEYES = "Sentinel Presence"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'GENERALIZED_ENVIRONMENTGreen Area:DISTURBANCE_FREQUENCY' = "Generalized Environment: Disturbance Frequency"
  , 'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
  )

labs.ptwy<-c('(Intercept)' = "Intercept"
  , SENTINEL_PRESENCEYES = "Sentinel Presence"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
  )
```

```{r DATA.SR & BOUT}
DATA.SR<-read.csv("DATA.SR.csv", stringsAsFactors = T) %>%
  rename("VIDEO_ID" = "VIDEO_ID."
         , "ID" = "ID.")
BOUT.raw<-read.csv("BOUT.csv", stringsAsFactors = T)
```

# Proportion Data

```{r Open PROP}
PROP<-DATA.SR[,c(1,2,15,17,19,22,30,36,41,46)] %>% 
  subset(.
         , HU_BEHAVIOR_PROPORTION_... != 0 
         & HD_BEHAVIOR_PROPORTION_... != 0 
         ) %>% #Remove cases where proportion = 1 or 0
  rename("HU" = "HU_BEHAVIOR_PROPORTION_..."
         , "HD" = "HD_BEHAVIOR_PROPORTION_..."
         , "M" = "M_BEHAVIOR_PROPORTION_..."
         , "DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
         ) %>%
  pivot_longer(., cols = c("HU"
                              , "HD"
                              , "M"
                           )
               , names_to = "BEHAVIOR"
               , values_to = "PROPORTION")

str(PROP)
```

For every individual, I divided the time spent performing each behavior
by the total time the individual was recorded. To remove outliers, I
removed all observations with no time spent performing one or more
behaviors (4 observations).

```{r PROP Summary}
PROP.SUMMARY <- summarySE(data = PROP
            , measurevar = "PROPORTION"
            , groupvars = c("GENERALIZED_ENVIRONMENT"
                            , "BEHAVIOR"
                            , "SENTINEL_PRESENCE"
                            )
            )
PROP.SUMMARY$SENTINEL_PRESENCE<-as.factor(replace(as.character(PROP.SUMMARY$SENTINEL_PRESENCE)
                                                  , PROP.SUMMARY$SENTINEL_PRESENCE=="NO"
                                                  , "Sentinel Absent"
                                                  )
                                          )

PROP.SUMMARY$SENTINEL_PRESENCE<-as.factor(replace(as.character(PROP.SUMMARY$SENTINEL_PRESENCE)
                                                  , PROP.SUMMARY$SENTINEL_PRESENCE=="YES"
                                                  , "Sentinel Present"
                                                  )
                                          )

PROP.SUMMARY
```

## PROP Stacked barplot

```{r PROP Barplot}
PROP.BARPLOT<-ggplot(PROP.SUMMARY
                 , aes(x = GENERALIZED_ENVIRONMENT
                       , y = PROPORTION
                       , fill = BEHAVIOR))+
  geom_bar(stat = 'identity'
           , position = 'stack')+
  geom_text(aes(label = paste0(formattable::digits(PROPORTION*100, dig=2)
                               , "%"))
            , position = position_stack(vjust = 0.5)
            , size = 4) +
  scale_y_continuous(labels = scales::percent) +
  theme_classic()+
  ylab("Proportion of time")+
  scale_x_discrete(labels=c("Commercial Area"
                            , "Green Area"))+
  scale_fill_manual(values = cbPalette, labels = c("Foraging"
                                                   , "Alert"
                                                   , "Moving")
                    , name="")+
  theme(legend.position = "bottom"
        , text = element_text(size = 18)
        , axis.title.x = element_blank()) +
  facet_grid(~SENTINEL_PRESENCE)
PROP.BARPLOT
```

To better visualize the effects of generalized environment and sentinel
presence, I calculated the mean proportion for each behavior and plotted
a stacked barplot for each combination of factors.

## PROP Model

```{r PROP Model}
PROP.MOD<-lm(PROPORTION~BEHAVIOR+SENTINEL_PRESENCE+GENERALIZED_ENVIRONMENT, data = subset(PROP, PROP$BEHAVIOR != "M"))

sjPlot::tab_model(PROP.MOD
                  , pred.labels = labs.prop
                  , show.stat = T
                  , show.se = T
                  , digits = 4
                  , show.re.var = T
                  , title = "Proportion Model Output"
                  , dv.labels = " Effects on the proportion of behaviors")
```

A linear model was performed. While all behaviors were included in the
model, "Moving" behavior was omitted from analysis, as not all bouts of
movement were recorded in their entirety.

Results: Crows allocate similar time to foraging and vigilance (Estimate
= 0.0263, SE = 0.0236, t-stat = 1.16, p = 0.248), and neither the
presence of a sentinel (Estimate = -0.0335, SE = 0.0234, t-stat =
-1.4314, p = 0.154) or the generalized environment (Estimate = 0.0336,
SE = 0.0230, t-stat = 1.4625, p = 0.146 have an effect on the
proportion of time allocated to either alert or foraging behavior.

# Bout Data

```{r Open BOUT}
BOUT<-BOUT.raw %>%
  filter(.
         , DURATION > 0.01) %>% #Remove impossibly small values
  filter(.
         , BEHAVIOR != "M") #Remove "M"
BOUT$LDURATION<-log(BOUT$DURATION) #Transform data for normality
str(BOUT)
```

In order to analyze the effects of generalized environment and the
presence of a sentinel on the duration of bouts, I first removed all
bouts of a duration shorter than 0.01s to remove impossibly small
outliers (likely artifacts from video coding in BORIS, 21 observations
removed), and removed all bouts where the individuals were moving as
they were not always recorded in their entirety (1173 observations
removed).

```{r BOUT Summary}
BOUT.SUMMARY <- summarySE(data = BOUT
                      , measurevar = "DURATION"
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "BEHAVIOR"
                                      , "SENTINEL_PRESENCE"
                                      )
                      ) %>%
  mutate(BEHAVIOR = recode_factor(BEHAVIOR
                                  ,'HD' = 'Foraging'#Rename HD as 'foraging'
                                  , 'HU' = 'Alert') #Rename HU as 'Alert'
                           )

BOUT.SUMMARY
```

## BOUT Dot Plot

```{r BOUT Dot Plot}
BOUT.DOTPLOT<-BOUT.SUMMARY %>%
           ggplot(.
               , aes(x = GENERALIZED_ENVIRONMENT
                     , y = DURATION
                     , colour = SENTINEL_PRESENCE))+
  geom_point(position = position_dodge(width = 0.1)
             , size = 3
             ) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.1))+
  geom_text(aes(label = round(DURATION, 2))
            , hjust = -0.2
            , size = 6
            , show.legend = F) +
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , labels = c("Sentinel Absent", "Sentinel Present")
                      , name = "") +
  scale_x_discrete(labels = c("Commercial Area", "Green Area"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18)
        ) +
  facet_grid(~BEHAVIOR)

BOUT.DOTPLOT
```

> The error bars represent the standard error, while the symbols are the
> mean duration of bouts.

## BOUT Model - All behaviors

```{r BOUT Model - All behaviors}
BOUT.MOD<-rlmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|ID), data = BOUT)

sjPlot::tab_model(BOUT.MOD
                  , pred.labels = labs.bout.1
                  , show.re.var = T
                  , show.se = T
                  , show.stat = T
                  , digits = 4
                  , title = "Bout Robust Model  Output"
                  , dv.labels = " Effects on duration of bouts of all behaviors")

sjPlot::plot_model(BOUT.MOD
                  , axis.labels = labs.bout.1
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on duration of bouts of all behaviors"
                  )
```

I then visualized the data and ran a robust linear mixed model on the
bouts of all behaviors using the following structure:

\$Log(Bout Duration) \~ Behavior X Sentinel Presence + Behavior X
Generalized Environment + Sentinel Presence X Generalized Environment +
Group Size + Bait Presence + Disturbance Frequency + ( 1 \| Individual
ID ) \$

Where the behavior, presence of a sentinel, generalized environment,
group size, bait presence, and disturbance frequency are fixed effects,
while the individual ID is the random effect.

I subsequently ran robust linear mixed models on each behavior
separately, using a similar formula as described above.

RESULTS: Bouts of alertness and foraging were significantly different
(Estimate = -0.2557, SE = 0.0511, t-stat = -5.002, p = \<0.001), with
bouts of alertness being significantly shorter than bouts of vigilance.

Sentinel presence increased the duration of all bouts significantly
(Estimate = 0.1974, SE = 0.0720, t-stat = 2.7406, p = 0.006). All bouts
in green areas were significantly longer than those in commercial areas
(Estimate = 0.3534, SE = 0.0873, t-stat = 4.0482, p = \<0.001). The
interaction between generalized environment and sentinel presence had a
significant effect (Estimate = -0.2524, SE = 0.0882, t-stat = -2.8630, p
= 0.004).

The interaction between behavior type and generalized environment was
also significant (Estimate = -0.2023, SE = 0.0537, t-stat = -3.7690, p =
\<0.001).

Lastly, disturbance frequency had a significant effect on the duration
of all bouts (Estimate = -0.0878, SE = 0.0295, t-stat = -2.9748, p =
0.003), with bout duration decreasing as disturbance frequency
increased.

```{r BOUT Post Hoc - All Behaviors}
BOUT.DIFF.ALL_BEHAVIORS<-emmeans(BOUT.MOD, ~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
test(pairs(BOUT.DIFF.ALL_BEHAVIORS), adjust="fdr")
```

RESULTS Post-hoc pairwise testing revealed significant differences in
the duration of all bouts. The following results are averaged over the
levels of bait, group size and bait presence, and the p-values are
adjusted using the 'FDR' method for 6 tests. In commercial areas, the
presence of a sentinel increased the duration of foraging bouts
(Estimate = -0.157, SE = 0.0653, z-ratio = -2.402, p = 0.0489). In the
absence of a sentinel, foragers in green areas had longer foraging bouts
(Estimate = -0.252, SE = 0.0821, z-ratio = -3.074, p = 0.0127). Foragers
in commercial areas and in the absence of a sentinel had a marginally
significant effect, shortening foraging bouts in green areas and in the
presence of a sentinel (Estimate = -0.157, SE = 0.0720, z-ratio =
-2.117, p = 0.0589). All other comparisons were not significant (p \<
0.3430)

##BOUT Model - Head Down

```{r BOUT Model - Head Down}
BOUT.MOD.HD<-rlmer(LDURATION~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|ID), data = subset(BOUT, BEHAVIOR == "HD"))

sjPlot::tab_model(BOUT.MOD.HD
                  , pred.labels = labs.bout.2
                  , show.re.var = T
                  , show.se = T
                  , show.stat = T
                  , digits = 4
                  , title = "Head Down"
                  , dv.labels = " Effects on foraging bout duration")

sjPlot::plot_model(BOUT.MOD.HD
                  , axis.labels = labs.bout.2
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on foraging bout duration")

```

RESULTS: Sentinel presence had no significant effect on the duration of
bouts of foraging (Estimate = 0.0919, SE = 0.0718, t-stat = 1.2799, p =
0.201).

Generalized environment had a significant effect on the duration of
bouts of foraging, with bouts being longer in green areas (Estimate =
0.3826, SE = 0.0778, t-stat = 4.9194, p = \<0.001). The interaction
between generalized environment and sentinel presence was also
significant (Estimate = -0.2272, SE = 0.0914, t-stat = -2.4849, p =
0.013).

Increasing group size significantly increased the duration of bouts of
foraging behavior (Estimate = -0.1519, SE = 0.0684, t-stat = -2.2205, p
= 0.026).

The presence of bait decreased the duration of foraging bouts (Estimate
= -0.1385, SE = 0.0697, t-stat = -1.9885, p = 0.047).

Increasing disturbance frequency significantly decreased the duration of
foraging bouts (Estimate = -0.1075, SE = 0.0301, t-stat = -3.5664, p =
\<0.001).

```{r BOUT HD - Group Size Figure}
BOUT.DOTPLOT.GROUP_SIZE<-BOUT %>%
  filter(., BEHAVIOR == "HD") %>%
  summarySE(., measurevar = "DURATION", groupvar = "GROUP_SIZE") %>%
           ggplot(.
               , aes(x = GROUP_SIZE
                     , y = DURATION
                     , colour = GROUP_SIZE
                     ))+
  geom_point(position = position_dodge(width = 0.1)
             , size = 3
             ) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.1))+
  geom_text(aes(label = round(DURATION, 2))
            , hjust = -0.2
            , size = 6
            , show.legend = F) +
  theme_classic() +
  xlab("Group Size") +
  ylab("Mean Bout Duration (s)") +
  scale_color_manual(values = cbPalette) +
  scale_x_discrete(labels = c("Small Group"
                                  , "Large Group")
                   , limits = c("SMALL", "LARGE")) +
  theme(axis.title.x = element_blank()
        , legend.position = "none"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18)
        )

BOUT.DOTPLOT.GROUP_SIZE
```

```{r BOUT HD - BAIT_PRESENCE Figure}
BOUT.DOTPLOT.BAIT<-BOUT %>%
  filter(., BEHAVIOR == "HD") %>%
  summarySE(., measurevar = "DURATION", groupvar = "BAIT_PRESENCE") %>%
           ggplot(.
               , aes(x = BAIT_PRESENCE
                     , y = DURATION
                     , colour = BAIT_PRESENCE
                     ))+
  geom_point(position = position_dodge(width = 0.1)
             , size = 3
             ) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.1))+
  geom_text(aes(label = round(DURATION, 2))
            , hjust = -0.2
            , size = 6
            , show.legend = F) +
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_color_manual(values = cbPalette) +
  scale_x_discrete(labels = c("Bait Absent"
                                  , "Bait Present")) +
  theme(axis.title.x = element_blank()
        , legend.position = "none"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18)
        )

BOUT.DOTPLOT.BAIT
```

```{r BOUT HD - DISTURBANCES Figure}
BOUT.DOTPLOT.DISTURBANCES<-BOUT %>%
  filter(., BEHAVIOR == "HD") %>%
           ggplot(.
               , aes(x = DISTURBANCE_FREQUENCY
                     , y = DURATION
                     ))+
  geom_point(size = 2
             , colour = cbPalette[1]
             , alpha = 0.5
             )+
  geom_smooth(method = "lm"
              , se = F
              , colour = cbPalette [2]) +
  theme_classic() +
  xlab("Disturbance Frequency (per minute)") +
  ylab("Mean Bout Duration (s)") +
  theme(legend.position = "none"
        , text = element_text(size = 18)
        )

BOUT.DOTPLOT.DISTURBANCES
```

## BOUT Model - Head Up

```{r BOUT Model - Head Up}
BOUT.MOD.HU<-rlmer(LDURATION~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|ID), data = subset(BOUT, BEHAVIOR == "HU"))

sjPlot::tab_model(BOUT.MOD.HU
                  , pred.labels = labs.bout.2
                  , show.re.var = T
                  , show.se = T
                  , show.stat = T
                  , digits = 4
                  , title = "Head Up"
                  , dv.labels = " Effects on alert bout duration")

sjPlot::plot_model(BOUT.MOD.HU
                  , axis.labels = labs.bout.2
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on alert bout duration")

```

RESULTS: In contrast to foraging behavior, generalized environment,
group size, bait presence and disturbance frequency had no significant
effect on the duration of bouts of alert behavior (p \< 0.141).

Sentinel behavior did not have an effect on the duration of alert
behavior.

However, the interaction between sentinel behavior and generalized
environment was significant (Estimate = -0.2736, SE = 0.1352, t-stat =
-2.0243, p = 0.043).

#Peck Rate

```{r Open PECK}
PECK<-DATA.SR[,c(1,2,15,17,19,22,30,38,48)] %>%
  subset(., HD_BEHAVIOR_DURATION > 0) %>%
  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES")

str(PECK)
```

The peck rate was calculated for every individual by dividing the total
number of pecks at food performed by the total duration of foraging (or
'Head Down') behavior. Individuals that did not forage were excluded.

```{r PECK Summary, echo = F}
PECK.SUMMARY<-summarySE(PECK, "PECK_RATE"
                     , c("GENERALIZED_ENVIRONMENT"
                         , "SENTINEL_PRESENCE")
                     , na.rm=T)
PECK.SUMMARY
```

## PECK Dot Plot

```{r PECK Dot Plot}
PECK.DOTPLOT<-ggplot(data = PECK.SUMMARY
               , aes(x = GENERALIZED_ENVIRONMENT
                     , y = PECK_RATE
                     , color = SENTINEL_PRESENCE))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3
             ) +
  geom_errorbar(aes(ymin=(PECK_RATE-se)
                    , ymax=(PECK_RATE+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  geom_text(aes(label = round(PECK_RATE, 2))
            , position = position_dodge(width = 0.4)
            , size = 6
            , show.legend = F) +
  theme_classic() +
  ylab("Mean Peck Rate (per min)") +
  scale_colour_manual(values = cbPalette
                      , labels = c("Sentinel Absent", "Sentinel Present")
                      , name = "")+
  scale_x_discrete(labels = c("Commercial Area"
                                  , "Green Area")) +
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18))

PECK.DOTPLOT
```

## PECK Model

I then ran another robust linear mixed model using a similar formula as
for the bout duration. Sentinel presence, generalized environment, group
size, bait presence and the total frequency of disturbances per minute
were used as fixed effects, with the individual ID used as a random
factor.

```{r PECK Model}
PECK.MOD<-rlmer(PECK_RATE~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GENERALIZED_ENVIRONMENT*DISTURBANCE_FREQUENCY+GROUP_SIZE+BAIT_PRESENCE+(1|ID), data = PECK)

sjPlot::tab_model(PECK.MOD
                  , pred.labels = labs.peck
                  , show.re.var = T
                  , show.se = T
                  , show.stat = T
                  , digits = 4
                  , title = "Robust Peck Rate Model Output"
                  , dv.labels = " Effects on peck rate")
sjPlot::plot_model(PECK.MOD
                  , axis.labels = labs.peck
                  , show.values=T
                  , show.p=T
                  , value.offset = 0.4
                  , value.size = 3.5
                  , wrap.title = 48
                  , show.re.var = T
                  , title = "Effects on peck rate" )
```

RESULTS Neither the presence of a sentinel nor the generalized
environment alone had a significant effect on the peck rate of foragers
(p \> 0.702).

Peck rate increases significantly as disturbance frequency increases
(Estimate = 5.29, t = 2.312, p = 0.021), with the interaction between
generalized environment and disturbance frequency also significantly
affecting peck rate (Estimate = 16.15, t = 3.046, p = 0.002).

Lastly, the presence of bait significantly increased the peck rate of
foragers (Estimate = 13.99, t = 2.231, p = 0.020).

## Disturbance Frequency x Generalized Environment

```{r PECK Disturbance plot}
PECK.DOTPLOT.DISTURBANCE<-PECK %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE
               , color = GENERALIZED_ENVIRONMENT))+
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  theme_classic() +
  labs(y="Peck Rate (per min)", x="Disturbance Frequency (per min)")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom"
        , text = element_text(size = 18))+
  scale_color_manual(values = cbPalette
                    , labels = c("Commercial Area"
                                 , "Green Area")
                    , name="")

PECK.DOTPLOT.DISTURBANCE
```

RESULTS Green areas appear to have much fewer disturbances than
commercial areas.

```{r PECK Dot Plot - Bait Presence}
PECK.DOTPLOT.BAIT<-summarySE(PECK
                             , measurevar = "PECK_RATE"
                             , groupvars = "BAIT_PRESENCE") %>%
  ggplot(., aes(x = BAIT_PRESENCE
                     , y = PECK_RATE
                     , color = BAIT_PRESENCE))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3
             ) +
  geom_errorbar(aes(ymin=(PECK_RATE-se)
                    , ymax=(PECK_RATE+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  geom_text(aes(label = round(PECK_RATE, 2))
            , position = position_dodge(width = 0.9)
            , size = 6
            , show.legend = F) +
  theme_classic() +
  ylab("Mean Peck Rate (per min)") +
  scale_x_discrete(labels = c("Baited Absent"
                                  , "Bait Present")) +
  scale_color_manual(values = cbPalette) +
  theme(axis.title.x = element_blank()
        , legend.position = "none"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18))

PECK.DOTPLOT.BAIT
```

# Sentinel Likelihood

```{r open SENT}
SENT<-DATA.SR[1:81,c(1,7,10,15,17,19,20,22,30,31)] %>%
  distinct()
str(SENT)
```

A chi-squared test was performed to determine if there were
environmental or social factors that affected the likelihood of a
sentinel being present in our videos.

```{r SENT Summary}
SENT.SUMMARY <- summarySE(data = SENT
                      , measurevar = "GROUP_SIZE" #Random
                      , groupvars = c("GENERALIZED_ENVIRONMENT"
                                      , "SENTINEL_PRESENCE"
                                      )
                      )

SENT.SUMMARY
```

##SENT Barplot

```{r SENT Barplot}
SENT.BARPLOT <- ggplot(SENT.SUMMARY, aes(x=GENERALIZED_ENVIRONMENT
                 , fill = SENTINEL_PRESENCE
                 , y = N
                 , group = SENTINEL_PRESENCE)) + 
  geom_bar(stat='identity', width=.5, position = "dodge") +
  geom_text(aes(label = N, group = SENTINEL_PRESENCE), position = position_dodge(width = 0.5), vjust = 1.5, size = 8) +
  ylab("Number of observations") +
  theme_classic() +
  scale_fill_manual(""
                    , values = cbPalette
                    , labels = c("Sentinel Absent", "Sentinel Present")
                      ) +
  scale_x_discrete(labels = c("Commercial Area"
                              , "Green Area"))
  theme(axis.title.x = element_blank()
        , text = element_text(size = 18)
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())

SENT.BARPLOT
```

##SENT Chi-Squared

```{r SENT Chi-Squared}
SENT.CHISQ<-lapply(SENT[,c("GENERALIZED_ENVIRONMENT"
                      , "GROUP_SIZE"
                      , "DISTURBANCE_FREQUENCY"
                      )
                              ]
                         , function(x) CrossTable(x,SENT$SENTINEL_PRESENCE
                                                  , fisher = T
                                                  , chisq = T
                                                  , expected = T
                                                  , prop.c = F
                                                  , prop.t = F
                                                  , prop.chisq = F
                                                  , sresid = T
                                                  , format = 'SPSS'
                                                  )
                         )

SENT.CHISQ
```

RESULTS Neither the generalized environment (Chi\^2 = 0.1221515, df = 1,
p = 0.727), group size (Chi\^2 = 0.2481203, df = 1, p = 0.618 ), or the
disturbance frequency (Chi\^2 = 2.032678, df = 2, p = 0.362)
significantly affected the likelihood of a sentinel being present.

##SENT Plots - Continuous Variables

```{r SENT Frequency plots}
SENT.FREQ <- SENT %>% 
  rename("Temperature (deg. C)" = "TEMPERATURE"
         , "Decimal Time" = "DECIMAL_TIME"
         , "Disturbance Frequency (per min)" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
         , "Number of Crows" = "NUMBER_OF_CROWS_RECORDED") %>%
  pivot_longer(., cols = c(#"Temperature (deg. C)"
                           # , "Decimal Time"
                           "Number of Crows"
                           , "Disturbance Frequency (per min)")
               , names_to = "Var"
               , values_to = "Val") %>%
  ggplot(aes(
    x = Val
    , colour = SENTINEL_PRESENCE))+
  geom_freqpoly(linewidth = 1
                , alpha=I(.6)) +
  scale_colour_manual(values = cbPalette
                      , name = ""
                      , labels = c("Sentinel Absent"
                                   , "Sentinel Present")) +
  facet_wrap(~Var, scale = "free_x") +
  theme_classic() +
  theme(legend.position = "bottom"
        , axis.title.x = element_blank()
        , text = element_text(size = 16))

SENT.FREQ
```

# Pathway Analysis

## Pathway

I calculated the likelihood of one behavior following another by
dividing the number of transitions from one behavior to another by the
total number of transitions from that behavior.

```{r Subsetting PTWY}
PTWY<- DATA.SR %>%
  dplyr::select(VIDEO_ID,ID, DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_NB_TRANSITIONS, TOTAL_FREQUENCY_OF_DISTURBANCES | starts_with("NB_")) %>%
  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES")

```

### PTWY Plots

```{r PTWY Plots}
PTWY.SUMMARY<-PTWY %>%
  rename("Head Down to Peck" = "NB_HD.HDP"
         , "Head Up to Head Down" = "NB_HU.HD"
         , "Head Down to Head Up" = "NB_HD.HU"
         , "Peck to Head Up" = "NB_HDP.HU") %>%
  pivot_longer(., cols = c("Head Down to Peck"
                           , "Head Up to Head Down"
                           , "Head Down to Head Up"
                           , "Peck to Head Up"
                           )
               , names_to = "TRANSITION"
               , values_to = "PTWY") %>%
  summarySE(.
            , measurevar = "PTWY"
            , groupvars = c("TRANSITION"
                            , "SENTINEL_PRESENCE"
                            , "GENERALIZED_ENVIRONMENT"))

PTWY.BOXPLOT<-PTWY %>%
  rename("Head Down to Peck" = "NB_HD.HDP"
         , "Head Up to Head Down" = "NB_HU.HD"
         , "Head Down to Head Up" = "NB_HD.HU"
         , "Peck to Head Up" = "NB_HDP.HU") %>%
  pivot_longer(., cols = c("Head Down to Peck"
                           , "Head Up to Head Down"
                           , "Head Down to Head Up"
                           , "Peck to Head Up"
                           )
               , names_to = "TRANSITION"
               , values_to = "PTWY") %>%
    ggplot(.
           , aes(x = GENERALIZED_ENVIRONMENT
                 , y = PTWY
                 , colour = SENTINEL_PRESENCE)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_dodge(width = 0.9)) +
  geom_point(data = PTWY.SUMMARY
             , aes( x = GENERALIZED_ENVIRONMENT
                    , y = PTWY
                    , group = SENTINEL_PRESENCE)
             , size = 2
             , colour = "black"
             , position = position_dodge(width = 0.9)) +
  geom_errorbar(data = PTWY.SUMMARY
                , aes(ymin=(PTWY-se)
                    , ymax=(PTWY+se)
                    , group = SENTINEL_PRESENCE
                    , x = GENERALIZED_ENVIRONMENT)
                , width = 0.2
                , colour = "black"
                , position = position_dodge(width=0.9)) +
  theme_classic() +
  ylab("Number of Transitions") +
  scale_colour_manual(values = cbPalette
                      , labels = c("Sentinel Absent"
                                   , "Sentinel Present")
                      , name = "") +
  scale_x_discrete(labels = c("Commercial", "Green")
                   )+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 12)
        ) +
  facet_grid(~TRANSITION)

PTWY.BOXPLOT
```

### GLMM

A generalized linear mixed model with a poisson distribution was
performed to assess changes in the frequency of transitions to each
behavior. Sentinel presence, generalized environment, disturbance
frequency and bait presence were used as fixed effects, with the total
number of transitions as a random effect.

```{r PTWY GLMM}
PTWY.MODELS<-list()
DV<-list("NB_HD.HU"
         , "NB_HD.HDP"
         , "NB_HU.HD"
         , "NB_HDP.HU"
         )

for (y in DV){
  formula<-formula(paste(y,"~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+BAIT_PRESENCE+(1|TOTAL_NB_TRANSITIONS)"))
  PTWY.MODELS[[y]]<-glmer(formula, data = PTWY, family = "poisson", control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=100000)))
}

tab_model(PTWY.MODELS
          , pred.labels = labs.ptwy
                  , show.re.var = T
                  , show.se = T
                  , show.stat = T
                  , digits = 4
                  , title = "Transitions GLMM output"
                  , dv.labels = c("Head Down to Head Up"
                                  , "Head Down to Peck"
                                  , "Head Up to Head Down"
                                  , "Peck to Head Up"))
```

```{r PTWY Post Hoc}
PTWY.DIFF<-emmeans(PTWY.MODELS$NB_HD.HU, ~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT)
test(pairs(PTWY.DIFF), adjust="fdr")
```

RESULTS Transitions from head down to pecking were significantly
affected by the presence of bait, with more transitions occuring in the
presence of bait (IRR = 1.7096, SE = 0.3842, z-stat = 2.3858, p =
0.017). All other factors did not affect this transition.

Transitions from pecking to head up were similarly affected by the
presence of bait, significantly increasing when bait was present (IRR =
2.2037, SE = 0.5378, z-stat = 3.2378, p = 0.001).

Transitions from head up to head down were not affected by any factors,
but bait presence had a marginally significant effect (IRR = 1.5134, SE
= 0.3506, z-stat - 1.7888, p = 0.074). In the presence of bait, the
number of transitions from head up to head down appears to increase.

Transitions from head down to head up, from vulnerability to vigilance,
were significantly affected by generalized environment, disturbance
frequency, and the interaction between generalized environment and
sentinel presence.

In the absence of a sentinel, individuals exhibited marginally more
transitions from foraging to alert behavior in commercial areas
(Estimate = 0.864, SE = 0.372, z-ratio = 2.321, p = 0.0608). However, in
the presence of a sentinel, individuals exhibited marginally more of the
same transitions in green areas (Estimate = -0.749, SE = 0.363, z-ratio
= -2.062, p = 0.0785).

In green areas, individuals performed more transitions from foraging to
alert behavior when in the presence of a sentinel (Estimate = -1.124, SE
= 0.346, z-ratio = -3.250, p = 0.0069).

```{r PTWY Figures - Bait Presence}
PTWY.DOTPLOT.BAIT<-PTWY %>%
  rename("Head Down to Peck" = "NB_HD.HDP"
         , "Head Up to Head Down" = "NB_HU.HD"
         , "Head Down to Head Up" = "NB_HD.HU"
         , "Peck to Head Up" = "NB_HDP.HU"
         ) %>%
  pivot_longer(., cols = c("Head Down to Peck"
                           , "Head Up to Head Down"
                           , "Head Down to Head Up"
                           , "Peck to Head Up"
                           )
               , names_to = "TRANSITION"
               , values_to = "PTWY"
               ) %>%
  summarySE(.
            , measurevar = "PTWY"
            , groupvars = c("BAIT_PRESENCE"
                            , "TRANSITION" 
                            )
            ) %>%
  ggplot(.
           , aes(x = BAIT_PRESENCE
                 , y = PTWY
                 , colour = BAIT_PRESENCE)
         ) +
  geom_point(position = position_dodge(width = 0.9)
             ) +
  geom_errorbar(aes(ymin=(PTWY-se)
                    , ymax=(PTWY+se)
                    )
                , width = 0.2
                , position = position_dodge(width=0.9)
                ) +
    geom_text(aes(label = round(PTWY, 2))
            , position = position_dodge2(width = 0.9)
            , show.legend = F) +
  theme_classic() +
  ylab("Number of Transitions") +
  scale_color_manual(values = cbPalette) +
  scale_x_discrete(labels = c("Bait Absent"
                              , "Bait Present")) +
  theme(axis.title.x = element_blank()
        , legend.position = "none"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 12)
        ) +
  facet_grid(~TRANSITION)

PTWY.DOTPLOT.BAIT
```

```{r PTWY Figures - Disturbance Frequency}
PTWY.DOTPLOT.DISTURBANCE<- PTWY %>%
  ggplot(.
         , aes(x = DISTURBANCE_FREQUENCY
               , y = NB_HD.HU
               )
         ) +
  geom_point(colour = cbPalette[1]
             ) +
  geom_smooth(method = "glm"
              , method.args = list(family = poisson)
              , colour = cbPalette[2]
              ) +
  theme_classic() +
  xlab("Disturbance Frequency (per min)") +
  ylab("Number of Transitions from Foraging to Alert")

PTWY.DOTPLOT.DISTURBANCE
```

### 2023-11-13 Data & Scripts/Alex Popescu - Final Stats.Rmd
Nov. 13
More writing!

@@
 library("ggplot2")
-library("ggpattern")
 library("lmerTest")
@@
 library("effects")
-library("glmm")
 library("doBy")
@@
 library("jmv")
+library("webshot")
+library("kableExtra")
 source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
@@
 
-```{r Labels}
+```{r Labels, include = F}
 labs.prop<-c(
@@
   '(Intercept)' = "Intercept"
-  , SENTINEL_PRESENCEYES = "Sentinel Presence"
+  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
   , BEHAVIORHU = "Behavior"
@@
   , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
-  , 'BEHAVIORHU:SENTINEL_PRESENCEYES' = "Behavior: Sentinel Presence"
-  , 'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
+  , 'BEHAVIORHU:SENTINEL_PRESENCESentinel Absent' = "Behavior: Sentinel Presence"
+  , 'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
 )
@@
   '(Intercept)' = "Intercept"
-  , SENTINEL_PRESENCEYES = "Sentinel Presence"
+  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
   , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
@@
   , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
-  ,'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
+  ,'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
   )
@@
 labs.ptwy<-c('(Intercept)' = "Intercept"
-  , SENTINEL_PRESENCEYES = "Sentinel Presence"
+  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
   , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
@@
   , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
-  , 'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
+  , 'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
   )
@@
   ylab("Proportion of time")+
-  scale_x_discrete(labels=c("Commercial Area"
-                            , "Green Area"))+
+  scale_x_discrete(labels=c("Commercial"
+                            , "Green"))+
   scale_fill_manual(values = cbPalette, labels = c("Foraging"
@@
                   , show.re.var = T
-                  , title = "Proportion Model Output"
-                  , dv.labels = " Effects on the proportion of behaviors")
+                  , title = ""
+                  , dv.labels = " Effects on the proportion of behaviors"
+                  , file = "Proportion_Table.html")
+```
+```{r Save PROP Table, include = F}
+webshot("Proportion_Table.html", "Proportion_Table.png")
 ```
@@
   filter(.
-         , BEHAVIOR != "M") #Remove "M"
+         , BEHAVIOR != "M") %>% #Remove "M"
+  mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
+                         , 'YES' = "Sentinel Present"
+                         , 'NO' = "Sentinel Absent"))
 BOUT$LDURATION<-log(BOUT$DURATION) #Transform data for normality
@@
                       , name = "") +
-  scale_x_discrete(labels = c("Commercial Area", "Green Area"))+
+  scale_x_discrete(labels = c("Commercial", "Green"))+
   theme(axis.title.x = element_blank()
@@
 BOUT.DIFF.ALL_BEHAVIORS<-emmeans(BOUT.MOD, ~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
-test(pairs(BOUT.DIFF.ALL_BEHAVIORS), adjust="fdr")
+kbl(test(pairs(BOUT.DIFF.ALL_BEHAVIORS), by = NULL, adjust="fdr"), format = "html", digits = 4, caption = "Bout post hoc - All behaviors", col.names = c("Contrast", "Estimate", "SE", "df", "z-ratio", "p"), align = "lccccc" )%>%
+  kable_classic(html_font = "Times New Roman") %>%
+  save_kable(file = "Bout_PH_All Behaviors.html", self_contained = T)
+```
+```{r Save Bout Post Hoc - All Behaviors, include = F}
+webshot("Bout_PH_All Behaviors.html", "Bout_PH_All Behaviors.png")
 ```
@@
 
+```{r BOUT - DISTURBANCES Figure}
+BOUT.DOTPLOT.DISTURBANCES<-BOUT %>%
+           ggplot(.
+               , aes(x = DISTURBANCE_FREQUENCY
+                     , y = DURATION
+                     ))+
+  geom_point(size = 2
+             , colour = cbPalette[1]
+             , alpha = 0.5
+             )+
+  geom_smooth(method = "lm"
+              , se = F
+              , colour = cbPalette [2]) +
+  theme_classic() +
+  xlab("Disturbance Frequency (per minute)") +
+  ylab("Bout Duration (s)") +
+  theme(legend.position = "none"
+        , text = element_text(size = 18)
+        )
+
+BOUT.DOTPLOT.DISTURBANCES
+```
+
 ##BOUT Model - Head Down
@@
 \<0.001).
+
+```{r BOUT Post Hoc - Foraging}
+BOUT.DIFF.HD<-emmeans(BOUT.MOD.HD, ~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
+kbl(test(pairs(BOUT.DIFF.HD), by = NULL, adjust="fdr"), format = "html", digits = 4, caption = "Bout post hoc - Foraging", col.names = c("Contrast", "Estimate", "SE", "df", "z-ratio", "p"), align = "lccccc" )%>%
+  kable_classic(html_font = "Times New Roman") %>%
+  save_kable(file = "Bout_PH_HD.html", self_contained = T)
+```
+```{r Save Bout Post Hoc - Foraging, include = F}
+webshot("Bout_PH_HD.html", "Bout_PH_HD.png")
+```
 
@@
   xlab("Group Size") +
-  ylab("Mean Bout Duration (s)") +
+  ylab("Mean Foraging Bout Duration (s)") +
   scale_color_manual(values = cbPalette) +
@@
   theme_classic() +
-  ylab("Mean Bout Duration (s)") +
+  ylab("Mean Foraging Bout Duration (s)") +
   scale_color_manual(values = cbPalette) +
@@
 ```{r BOUT HD - DISTURBANCES Figure}
-BOUT.DOTPLOT.DISTURBANCES<-BOUT %>%
+BOUT.DOTPLOT.DISTURBANCES.HD<-BOUT %>%
   filter(., BEHAVIOR == "HD") %>%
@@
   xlab("Disturbance Frequency (per minute)") +
-  ylab("Mean Bout Duration (s)") +
+  ylab("Foraging Bout Duration (s)") +
   theme(legend.position = "none"
@@
 
-BOUT.DOTPLOT.DISTURBANCES
+BOUT.DOTPLOT.DISTURBANCES.HD
 ```
@@
 
+```{r Save BOUT Tables, include = F}
+sjPlot::tab_model(c(BOUT.MOD, BOUT.MOD.HD, BOUT.MOD.HU)
+                  , pred.labels = labs.bout.1
+                  , show.re.var = T
+                  , show.se = T
+                  , show.stat = T
+                  , digits = 4
+                  , order.terms = c(1,3,4,5,6,7,10,2,8,9)
+                  , title = ""
+                  , dv.labels = c("Bouts of all behaviors"
+                                  , "Bouts of foraging behavior"
+                                  , "Bouts of alert behavior")
+                  , file = "Bout_Table.html"
+                  )
+webshot("Bout_Table.html", "Bout_Table.png")
+
+```
+
 #Peck Rate
@@
                 , position = position_dodge(width=0.9))+
-  geom_text(aes(label = round(PECK_RATE, 2))
+  geom_text(aes(label = scales::label_number(accuracy=0.01)(PECK_RATE))
             , position = position_dodge(width = 0.4)
             , size = 6
-            , show.legend = F) +
+            , show.legend = F
+            ) +
   theme_classic() +
@@
                       , name = "")+
-  scale_x_discrete(labels = c("Commercial Area"
-                                  , "Green Area")) +
+  scale_x_discrete(labels = c("Commercial"
+                                  , "Green")) +
   theme(axis.title.x = element_blank()
@@
                   , digits = 4
-                  , title = "Robust Peck Rate Model Output"
-                  , dv.labels = " Effects on peck rate")
+                  , title = ""
+                  , dv.labels = " Effects on peck rate (per min)"
+                  , file = "Peck_Table.html")
+
 sjPlot::plot_model(PECK.MOD
@@
                   , show.re.var = T
-                  , title = "Effects on peck rate" )
+                  , title = "Effects on peck rate (per min)" )
+```
+```{r Save Peck Table, include = F}
+webshot("Peck_Table.html", "Peck_Table.png")
 ```
@@
 ```{r PECK Dot Plot - Bait Presence}
-PECK.DOTPLOT.BAIT<-summarySE(PECK
+PECK.DOTPLOT.BAIT<-PECK %>%
+  summarySE(.
             , measurevar = "PECK_RATE"
             , groupvars = "BAIT_PRESENCE") %>%
-  ggplot(., aes(x = BAIT_PRESENCE
+  ggplot(.
+         , aes(x = BAIT_PRESENCE
                      , y = PECK_RATE
-                     , color = BAIT_PRESENCE))+
-  geom_point(position = position_dodge(width = 0.9)
+                     , colour = BAIT_PRESENCE
+               ))+
+  geom_point(position = position_dodge(width = 0.1)
              , size = 3
@@
                 , width = 0.1
-                , position = position_dodge(width=0.9))+
+                , position = position_dodge(width=0.1))+
   geom_text(aes(label = round(PECK_RATE, 2))
-            , position = position_dodge(width = 0.9)
+            , hjust = -0.2
             , size = 6
@@
                       ) +
-  scale_x_discrete(labels = c("Commercial Area"
-                              , "Green Area"))
+  scale_x_discrete(labels = c("Commercial"
+                              , "Green")) +
   theme(axis.title.x = element_blank()
@@
   dplyr::select(VIDEO_ID,ID, DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_NB_TRANSITIONS, TOTAL_FREQUENCY_OF_DISTURBANCES | starts_with("NB_")) %>%
-  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES")
+  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES") %>%
+  mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
+                         , 'YES' = "Sentinel Present"
+                         , 'NO' = "Sentinel Absent"))
 
@@
 PTWY.SUMMARY<-PTWY %>%
-  rename("Head Down to Peck" = "NB_HD.HDP"
-         , "Head Up to Head Down" = "NB_HU.HD"
-         , "Head Down to Head Up" = "NB_HD.HU"
-         , "Peck to Head Up" = "NB_HDP.HU") %>%
-  pivot_longer(., cols = c("Head Down to Peck"
-                           , "Head Up to Head Down"
-                           , "Head Down to Head Up"
-                           , "Peck to Head Up"
+  rename("Foraging to Peck" = "NB_HD.HDP"
+         , "Alert to Foraging" = "NB_HU.HD"
+         , "Foraging to Alert" = "NB_HD.HU"
+         , "Peck to Alert" = "NB_HDP.HU"
+         ) %>%
+  pivot_longer(., cols = c("Foraging to Peck"
+                           , "Alert to Foraging"
+                           , "Foraging to Alert"
+                           , "Peck to Alert"
                            )
@@
 PTWY.BOXPLOT<-PTWY %>%
-  rename("Head Down to Peck" = "NB_HD.HDP"
-         , "Head Up to Head Down" = "NB_HU.HD"
-         , "Head Down to Head Up" = "NB_HD.HU"
-         , "Peck to Head Up" = "NB_HDP.HU") %>%
-  pivot_longer(., cols = c("Head Down to Peck"
-                           , "Head Up to Head Down"
-                           , "Head Down to Head Up"
-                           , "Peck to Head Up"
+  rename("Foraging to Peck" = "NB_HD.HDP"
+         , "Alert to Foraging" = "NB_HU.HD"
+         , "Foraging to Alert" = "NB_HD.HU"
+         , "Peck to Alert" = "NB_HDP.HU"
+         ) %>%
+  pivot_longer(., cols = c("Foraging to Peck"
+                           , "Alert to Foraging"
+                           , "Foraging to Alert"
+                           , "Peck to Alert"
                            )
@@
   geom_boxplot(outlier.shape = NA) +
-  geom_jitter(position = position_dodge(width = 0.9)) +
+  geom_jitter(aes(x = GENERALIZED_ENVIRONMENT
+                  , group = SENTINEL_PRESENCE)
+              , position = position_jitterdodge(jitter.width = 0.5)
+              , alpha = 0.25
+              ) +
   geom_point(data = PTWY.SUMMARY
@@
              , colour = "black"
-             , position = position_dodge(width = 0.9)) +
+             , position = position_dodge(width = 0.75)) +
   geom_errorbar(data = PTWY.SUMMARY
@@
                 , colour = "black"
-                , position = position_dodge(width=0.9)) +
+                , position = position_dodge(width=0.75
+                                            )) +
   theme_classic() +
@@
                   , digits = 4
-                  , title = "Transitions GLMM output"
-                  , dv.labels = c("Head Down to Head Up"
-                                  , "Head Down to Peck"
-                                  , "Head Up to Head Down"
-                                  , "Peck to Head Up"))
+                  , title = ""
+                  , dv.labels = c("Foraging to Alert"
+                                  , "Foraging to Peck"
+                                  , "Alert to Foraging"
+                                  , "Peck to Alert")
+          , file = "PTWY_Table.html")
+
+```
+```{r Save PTWY Table, include = F}
+webshot("PTWY_Table.html", "PTWY_Table.png")
 ```
@@
 PTWY.DIFF<-emmeans(PTWY.MODELS$NB_HD.HU, ~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT)
-test(pairs(PTWY.DIFF), adjust="fdr")
+kbl(test(pairs(PTWY.DIFF), by = NULL, adjust="fdr"), format = "html", digits = 4, caption = "Pathway analysis post hoc", col.names = c("Contrast", "Estimate", "SE", "df", "z-ratio", "p"), align = "lccccc" )%>%
+  kable_classic(html_font = "Times New Roman") %>%
+  save_kable(file = "PTWY_PH.html", self_contained = T)
+```
+```{r Save PTWY} Post Hoc - Foraging, include = F}
+webshot("PTWY_PH.html", "PTWY_PH.png")
 ```
@@
 PTWY.DOTPLOT.BAIT<-PTWY %>%
-  rename("Head Down to Peck" = "NB_HD.HDP"
-         , "Head Up to Head Down" = "NB_HU.HD"
-         , "Head Down to Head Up" = "NB_HD.HU"
-         , "Peck to Head Up" = "NB_HDP.HU"
+  rename("Foraging to Peck" = "NB_HD.HDP"
+         , "Alert to Foraging" = "NB_HU.HD"
+         , "Foraging to Alert" = "NB_HD.HU"
+         , "Peck to Alert" = "NB_HDP.HU"
          ) %>%
-  pivot_longer(., cols = c("Head Down to Peck"
-                           , "Head Up to Head Down"
-                           , "Head Down to Head Up"
-                           , "Peck to Head Up"
+  pivot_longer(., cols = c("Foraging to Peck"
+                           , "Alert to Foraging"
+                           , "Foraging to Alert"
+                           , "Peck to Alert"
                            )
@@
                 , width = 0.2
-                , position = position_dodge(width=0.9)
-                ) +
-    geom_text(aes(label = round(PTWY, 2))
-            , position = position_dodge2(width = 0.9)
+                ) +
+    geom_text(aes(label = scales::label_number(accuracy = 0.01)(PTWY)
+                  , x = BAIT_PRESENCE
+                  , colour = BAIT_PRESENCE)
+            , hjust = -0.1
             , show.legend = F) +
@@
   ylab("Number of Transitions") +
+  xlab("Bait Presence") +
   scale_color_manual(values = cbPalette) +
-  scale_x_discrete(labels = c("Bait Absent"
-                              , "Bait Present")) +
-  theme(axis.title.x = element_blank()
-        , legend.position = "none"
+  scale_x_discrete(labels = c("Absent"
+                              , "Present")) +
+  theme(legend.position = "none"
         , legend.box="vertical"

### 2023-11-13 Data & Scripts/Old Stats/BOUT.Rmd
Nov. 13
More writing!

--- 2023-08-25 Data & Scripts/BOUT.Rmd
@@
 )
+
+pl2<-c(
+  '(Intercept)' = "Intercept"
+  , SENTINEL_PRESENCEYES = "Sentinel Presence"
+  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
+  , GROUP_SIZESMALL = "Group Size"
+  , BAIT_PRESENCEYES = "Bait Presence"
+  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
+  , 'SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
+)
 ```
@@
                   , dv.labels = " Effects on duration of bouts of behaviors")
+sjPlot::plot_model(BOUT.RMOD2
+                  , axis.labels = pl
+                  , rm.terms = c("BEHAVIORM"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES"
+                                 , "BEHAVIORM:GENERALIZED_ENVIRONMENTGreen Area"
+                                 ,"BEHAVIORM:SENTINEL_PRESENCEYES:GENERALIZED_ENVIRONMENTGreen Area")
+                  , show.values=T
+                  , show.p=T
+                  , value.offset = 0.4
+                  , value.size = 3.5
+                  , wrap.title = 48
+                  , show.re.var = T
+                  , title = "Bout Robust Model 2 Output"
+                  , dv.labels = " Effects on duration of bouts of behaviors")
 ```
@@
 
+#### Separate behaviors
+
+I want to determine the effects of my variables within each behavior (excluding 'Moving', since it is not of use to us) to better ascertain the direction of the effects.
+I will reuse Model 2's formula, removing the behavior from the formula.
+
+```{r BOUT.Mod2 - Separate behavior}
+BOUT.MOD2.HD<-rlmer(LDURATION~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = subset(BOUT, BEHAVIOR == "HD"))
+
+BOUT.MOD2.HU<-rlmer(LDURATION~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = subset(BOUT, BEHAVIOR == "HU"))
+
+sjPlot::tab_model(BOUT.MOD2.HD
+                  , pred.labels = pl2
+                  , show.re.var = T
+                  , title = ""
+                  , dv.labels = " Effects on duration of bouts with head down")
+sjPlot::tab_model(BOUT.MOD2.HU
+                  , pred.labels = pl2
+                  , show.re.var = T
+                  , title = ""
+                  , dv.labels = " Effects on duration of bouts with head up")
+
+```
+
+OK! Very interesting! When I separate the behaviors (i.e. not looking at the fixed effects on all behaviors), we get different results.
+
+For head down:
+- Sentinel presence is not significant.
+- Generalized environment is significant, with increased duration of bouts with head down in green areas.
+- Group size is significant, with shorter duration in smaller groups
+- The presence of bait increased duration of bouts with head down
+- The increased disturbance frequency decreased the duration of bouts with head down
+- The interaction between sentinel presence and generalized environment decreased the duration of bouts with head down
+
+For head up:
+- All fixed effects not significant
+- The interaction between sentinel presence and generalized environment significantly decreased the duration of bouts with head up.
+
+These results differ from those in BOUT.MOD2. This is to be expected since we are now looking at the effects on the separate behaviors.
+
+Let's plot the effect sizes:
+
+```{r Effect sizes - Separate behaviors}
+sjPlot::plot_model(BOUT.MOD2.HD
+                  , axis.labels = pl2
+                  , show.values=T
+                  , show.p=T
+                  , value.offset = 0.4
+                  , value.size = 3.5
+                  , wrap.title = 48
+                  , show.re.var = T
+                  , title = "Effects on head down bout duration")
+
+sjPlot::plot_model(BOUT.MOD2.HU
+                  , axis.labels = pl2
+                  , show.values=T
+                  , show.p=T
+                  , value.offset = 0.4
+                  , value.size = 3.5
+                  , wrap.title = 48
+                  , show.re.var = T
+                  , title = "Effects on head up bout duration")
+```
 #### Post Hoc testing
@@
 ```{r BOUT Post Hoc}
-BOUT.DIFF<-emmeans(BOUT.MOD2, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
+BOUT.DIFF<-emmeans(BOUT.MOD2, ~BEHAVIOR+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
 test(pairs(BOUT.DIFF, by= "BEHAVIOR"), adjust="fdr")
@@
 There are differences caused by the presence of a sentinel in commercial areas.
-There are significant differences between bout durations in commercial areas without a sentinel and bout durations in green areas, in both the presence and absence of a sentinel.
+There are significant differences between bout duration in commercial areas without a sentinel and bout duration in green areas, in both the presence and absence of a sentinel.
 
@@
 
-```{r Pairwise comparisons}
-BOUT.DOT
+```{r BOUT Dot Plot2}
+BOUT.DOT2<-ggplot(data = BOUT.MEAN
+               , aes(x = GENERALIZED_ENVIRONMENT
+                     , y = DURATION
+                     , color = GENERALIZED_ENVIRONMENT
+                     , shape = SENTINEL_PRESENCE))+
+  geom_point(position = position_dodge(width = 0.9)
+             , size = 3) +
+  geom_errorbar(aes(ymin=(DURATION-se)
+                    , ymax=(DURATION+se))
+                , width = 0.2
+                , position = position_dodge(width=0.9))+
+  theme_classic() +
+  ylab("Mean Bout Duration (s)") +
+  scale_colour_manual(values = cbPalette
+                      , guide="none") +
+  scale_shape_manual(values = c(16,17)
+                     , labels = c("Sentinel Absent"
+                                  , "Sentinel Present")
+                     , name = "") +
+  scale_x_discrete(labels = c("Commercial", "Green Area"))+
+  theme(axis.title.x = element_blank()
+        , legend.position = "bottom"
+        , legend.box="vertical"
+        , legend.margin=margin()) +
+  facet_grid(~BEHAVIOR)
+
+BOUT.DOT2
 ```
-
-Huh... I would have expected that there would be significant differences in bouts of 'head up' behavior.
 
@@
 -   Behavior, generalized environment and group size have significant effects. The former is to be expected, while the latter two are interesting, but not very informative (alone).
--   The interaction between behavior and generalized environment are significant. The effect size appears negative for both "head up" and "moving" behavior in green areas.
--   There is a significant interaction of sentinel presence and generalized environment.
--   There is a significant interaction of behavior and group size
--   There is a significant interaction of "moving" and bait presence.
+
+When separating by behavior type, we get more information about the effects of generalized environment and sentinel presence
+- Sentinel presence had no effect on the duration of bouts of head down and head up, this is unexpected!
+- Generalized environment increased the duration of bout of head down behavior, but not head up.
+Also unexpected but explainable.
+Green areas require more time spent with head down to look for and manipulate food, especially when in grass.
+- Duration of bouts of head up were not significantly affected by any fixed effects, apart from the interaction between sentinel presence and generalized environment.
+This decreased the duration of bouts of head up.
+- Bouts of head down behavior were also affected by the interaction of sentinel behavior and generalized environment (which decreased the duration), but also group size (smaller = shorter duration), bait presence (yes = shorter duration), and disturbance frequency (higher = shorter duration)
 
@@
   summarySE(measurevar = "DURATION"
-            , groupvars = c("SENTINEL_PRESENCE", "GENERALIZED_ENVIRONMENT")) %>%
+            , groupvars = c("SENTINEL_PRESENCE", "GENERALIZED_ENVIRONMENT", "BEHAVIOR")) %>%
     ggplot(aes(x = SENTINEL_PRESENCE
                      , y = DURATION
-                     , color = SENTINEL_PRESENCE))+
+                     , color = GENERALIZED_ENVIRONMENT))+
   geom_point(position = position_dodge(width = 0.9)
@@
   scale_colour_manual(values = cbPalette
-                      , guide="none")+
-  scale_x_discrete(labels = c("Sentinel Absent", "Sentinel Present"))+
-  theme(axis.title.x = element_blank()
-        , legend.position = "bottom"
+                      , labels = c("Commercial", "Green")
+                      , "Generalized Environment")+
+  xlab("Sentinel Presence") +
+  scale_x_discrete(labels = c("Absent", "Present")
+                   )+
+  theme(legend.position = "bottom"
         , legend.box="vertical"
         , legend.margin=margin()) +
-  facet_grid(~GENERALIZED_ENVIRONMENT)
+  facet_grid(~BEHAVIOR)
 
@@
 ```{r Disturb}
-DF.DOT<-BOUT %>%
+DF.DOT<-BOUT[which(BOUT$BEHAVIOR == "HD"),] %>%
     ggplot(aes(x = DISTURBANCE_FREQUENCY
                , y = DURATION
-               , color = BEHAVIOR))+
-  geom_point(size = 3) +
+               ))+
+  geom_point(size = 2, colour = 'red') +
   theme_classic() +
-  labs(y="Mean Bout Duration (s)", x="Disturbances per Minute", color= "Behavior")+
-  scale_colour_manual(values = cbPalette, labels = c("Foraging", "Alert", "Moving"))+
+  labs(y="Mean Head Down Bout Duration (s)", x="Disturbances per Minute")+
   scale_x_continuous(n.breaks=14)+

### 2023-11-13 Data & Scripts/Old Stats/PTWY.Rmd
Nov. 13
More writing!

--- 2023-08-18 Data & Scripts/PTWY.Rmd
@@
 
-Seems like it worked. Let's try fitting the model.
+Let's try fitting the model.
 
@@
 
-```{r MANOVA Vulnerable}
+```{r PTWY MANOVA Vulnerable}
 PTWY.vuln<-MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = PTWY, subject = "ID.", iter=1000)
@@
 
-```{r PTWY CR Vuln}
+```{r PTWY CR Vuln GENV}
 simCI(PTWY.vuln, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "GENERALIZED_ENVIRONMENT")
@@
 **Figure out what to do with factors such as bait presence, group size and disturbance frequency**
+
+## The other variables - Vulnerability
+
+We still have bait presence, group size and disturbance frequency to look at. I want to try and make a quick function to run all the tests.
+
+```{r PTWY Vulnerable function}
+PTWY.vuln.all<-lapply(PTWY[,c("SENTINEL_PRESENCE"
+                              ,"GENERALIZED_ENVIRONMENT"
+                              , "GROUP_SIZE"
+                              , "BAIT_PRESENCE"
+                              , "DISTURBANCE_FREQUENCY")
+                              ]
+                         , function(x) MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~x, data = PTWY, iter = 1000)
+                         )
+```
+```{r PTWY Vulnerable summaries, echo = F}
+PTWY.vuln.SENT<-summary(PTWY.vuln.all$SENTINEL_PRESENCE)
+PTWY.vuln.GENV<-summary(PTWY.vuln.all$GENERALIZED_ENVIRONMENT)
+PTWY.vuln.GS<-summary(PTWY.vuln.all$GROUP_SIZE)
+PTWY.vuln.BP<-summary(PTWY.vuln.all$BAIT_PRESENCE)
+PTWY.vuln.DF<-summary(PTWY.vuln.all$DISTURBANCE_FREQUENCY)
+```
+```{r PTWY Vulnerable Results}
+Factor<-c("Sentinel Presence", "Generalized Environment", "Group Size", "Bait Presence", "Disturbance Frequency")
+Results.vuln<-cbind(Factor, rbind(PTWY.vuln.SENT, PTWY.vuln.GENV, PTWY.vuln.GS, PTWY.vuln.BP, PTWY.vuln.DF))
+Results.vuln
+```
+
+As seen above, Gen.
+env has a  significant effect (p-value = 0.007).
+Interestingly,  group size (p-value ~ 0.072), bait presence (p-value ~ 0.06), and disturbance frequency (p-value ~ 0.01) are have either significant or marginally significant effects on the frequency of pathways to vulnerability.
+
+### Post Hoc - Disturbance Frequency
+
+```{r PTWY CR Vuln DF}
+PTWY.DF.Vuln<-MANOVA.wide(cbind(PTWY$PTWY_HD.HDP, PTWY$PTWY_HU.HD)~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
+simCI(PTWY.DF.Vuln, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "DISTURBANCE_FREQUENCY")
+
+```
+
+HMMMMM... Not significant... Univariate analysis time.
+
+```{r PTWY.DF Vulnerable univar}
+PTWY.DF.HD.HDP<-MANOVA.wide(PTWY_HD.HDP~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
+PTWY.DF.HU.HD<-MANOVA.wide(PTWY_HU.HD~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
+
+p.adjust(c(PTWY.DF.HD.HDP$resampling[,1], PTWY.DF.HU.HD$resampling[,1]), method = "bonferroni")
+#using the parametric bootstrap Wald-type statistic
+
+```
+
+Curious... the frequency of the pathway HD>HDPeck is marginally significant. Let's quickly plot it.
+
+### Plots
+
+```{r PTWY.DF.vuln plots}
+source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
+DF.PTWY.vuln<-PTWY %>%
+  summarySE(measurevar = "PTWY_HD.HDP"
+            , groupvars = "DISTURBANCE_FREQUENCY") %>%
+    ggplot(aes(x = DISTURBANCE_FREQUENCY
+                     , y = PTWY_HD.HDP
+                     , color = DISTURBANCE_FREQUENCY))+
+  geom_point(position = position_dodge(width=0.9)
+             , size = 3) +
+  geom_errorbar(aes(ymin=(PTWY_HD.HDP-se)
+                    , ymax=(PTWY_HD.HDP+se))
+                , width = 0.1
+                , position = position_dodge(width=0.9))+
+  theme_classic() +
+  ylab("Head Down to Head Down + Peck (%)") +
+  scale_colour_manual(values = cbPalette
+                      , guide="none")+
+  theme(axis.title.x = element_blank()
+        , legend.position = "bottom"
+        , legend.box="vertical"
+        , legend.margin=margin()) +
+  scale_x_discrete(limits = c("LOW", "MEDIUM", "HIGH"))
+
+DF.PTWY.vuln
+
+```
+
+THATS UNEXPECTED...
+well..
+not that unexpected.
+It mirrors the results of the peck data!
+As disturbance frequency increases, so does the peck rate, and therefore the frequency of head down to head down + peck.
+It could be a sense of urgency.
+
+## The other variables - Vulnerability
+
+We still have bait presence, group size and disturbance frequency to look at. I want to try and make a quick function to run all the tests.
+
+```{r PTWY Vigilant function}
+PTWY.vigil.all<-lapply(PTWY[,c("SENTINEL_PRESENCE"
+                              ,"GENERALIZED_ENVIRONMENT"
+                              , "GROUP_SIZE"
+                              , "BAIT_PRESENCE"
+                              , "DISTURBANCE_FREQUENCY")
+                              ]
+                         , function(x) MANOVA.wide(cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)~x, data = PTWY, iter = 1000)
+                         )
+```
+```{r PTWY Vigilant summaries, echo = F}
+PTWY.vigil.SENT<-summary(PTWY.vigil.all$SENTINEL_PRESENCE)
+PTWY.vigil.GENV<-summary(PTWY.vigil.all$GENERALIZED_ENVIRONMENT)
+PTWY.vigil.GS<-summary(PTWY.vigil.all$GROUP_SIZE)
+PTWY.vigil.BP<-summary(PTWY.vigil.all$BAIT_PRESENCE)
+PTWY.vigil.DF<-summary(PTWY.vigil.all$DISTURBANCE_FREQUENCY)
+```
+```{r PTWY Vigilant Results}
+Results.vigil<-cbind(Factor, rbind(PTWY.vigil.SENT, PTWY.vigil.GENV, PTWY.vigil.GS, PTWY.vigil.BP, PTWY.vigil.DF))
+Results.vigil
+```
+
+Another interesting result! Disturbance frequency is marginally significant for transitions to vigilance. Same workflow as above.
+
+```{r PTWY DF CR Vigil}
+PTWY.DF.Vigil<-MANOVA.wide(cbind(PTWY$PTWY_HD.HU, PTWY$PTWY_HDP.HU)~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
+simCI(PTWY.DF.Vigil, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "DISTURBANCE_FREQUENCY")
+
+```
+
+Pairwise comparisons for medium - high is marginally significant.
+
+```{r PTWY.DF Vigil univar}
+PTWY.DF.HD.HU<-MANOVA.wide(PTWY_HD.HU~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
+PTWY.DF.HDP.HU<-MANOVA.wide(PTWY_HDP.HU~DISTURBANCE_FREQUENCY, data = PTWY, iter = 1000)
+
+p.adjust(c(PTWY.DF.HD.HU$resampling[,1], PTWY.DF.HDP.HU$resampling[,1]), method = "bonferroni")
+#using the parametric bootstrap Wald-type statistic
+
+```
+
+Disturbance frequency has a marginally significant effect on transitions from Head Down + Peck to Head Up. Plot time!
+
+```{r PTWY.DF.vigil plots}
+source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
+DF.PTWY.vigil<-PTWY %>%
+  summarySE(measurevar = "PTWY_HDP.HU"
+            , groupvars = "DISTURBANCE_FREQUENCY") %>%
+    ggplot(aes(x = DISTURBANCE_FREQUENCY
+                     , y = PTWY_HDP.HU
+                     , color = DISTURBANCE_FREQUENCY))+
+  geom_point(position = position_dodge(width=0.9)
+             , size = 3) +
+  geom_errorbar(aes(ymin=(PTWY_HDP.HU-se)
+                    , ymax=(PTWY_HDP.HU+se))
+                , width = 0.1
+                , position = position_dodge(width=0.9))+
+  theme_classic() +
+  ylab("Head Down + Peck to Head Up  (%)") +
+  scale_colour_manual(values = cbPalette
+                      , guide="none")+
+  theme(axis.title.x = element_blank()
+        , legend.position = "bottom"
+        , legend.box="vertical"
+        , legend.margin=margin()) +
+  scale_x_discrete(limits = c("LOW", "MEDIUM", "HIGH"))
+
+DF.PTWY.vigil
+
+```
+
+# Frequency
+
+We'll be fitting another manova on the frequency data.
+
+```{r Subsetting FREQ}
+FREQ<- PTWY.raw %>%
+  dplyr::select(VIDEO_ID.,ID., DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_FREQUENCY_OF_DISTURBANCES, DISTURBANCE_FREQUENCY | starts_with("FREQ"))
+
+```
+
+Let's try fitting the model.
+
+## Vulnerable
+
+```{r FREQ MANOVA Vulnerable}
+FREQ.vuln<-MANOVA.wide(cbind(FREQ$FREQ_HD.HDP, FREQ$FREQ_HU.HD)~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, data = FREQ, subject = "ID.", iter=1000)
+summary(FREQ.vuln)
+```
+
+Same as with the PTWY data, only the generalized environment had a significant effect, and that only with the Wald-type statistic. Weird...
+
+### Post Hoc
+
+```{r FREQ CR Vuln}
+simCI(FREQ.vuln, contrast = "pairwise", type = "Tukey", interaction  = F, factor = "GENERALIZED_ENVIRONMENT")
+
+```
+
+Womp womp. Not significant. :/
+However, this still follows the same pattern as the PTWY analysis. Let's perform a univariate analysis to confirm.
+
+```{r FREQ Vulnerable univar}
+FREQ.HD.HDP<-MANOVA.wide(FREQ_HD.HDP~GENERALIZED_ENVIRONMENT, data = FREQ, iter = 1000)
+FREQ.HU.HD<-MANOVA.wide(FREQ_HU.HD~GENERALIZED_ENVIRONMENT, data = FREQ, iter = 1000)
+
+p.adjust(c(FREQ.HD.HDP$resampling[,1], FREQ.HU.HD$resampling[,1]), method = "bonferroni")
+#using the parametric bootstrap Wald-type statistic
+
+```
+
+Cool! HU>HD is marginally insignificant with a p-value of ~0.08.
+Onto vigilance!
 

### 2023-11-18 Data & Scripts/Alex Popescu - Final Stats.Rmd
More edits to methods ands results

@@
             , position = position_stack(vjust = 0.5)
-            , size = 4) +
+            , size = 4
+            , family = "serif") +
   scale_y_continuous(labels = scales::percent) +
@@
   theme(legend.position = "bottom"
-        , text = element_text(size = 18)
+        , text = element_text(size = 18, family = "serif")
         , axis.title.x = element_blank()) +
@@
                   , title = ""
-                  , dv.labels = " Effects on the proportion of behaviors"
+                  , dv.labels = ""
                   , file = "Proportion_Table.html")
@@
             , size = 6
-            , show.legend = F) +
+            , show.legend = F
+            , family = "serif") +
   theme_classic() +
@@
         , legend.margin=margin()
-        , text = element_text(size = 18)
+        , text = element_text(size = 18, family = "serif")
         ) +
@@
                   , digits = 4
-                  , title = "Bout Robust Model  Output"
-                  , dv.labels = " Effects on duration of bouts of all behaviors")
+                  , title = ""
+                  , dv.labels = "")
 
@@
                   , show.re.var = T
-                  , title = "Effects on duration of bouts of all behaviors"
-                  )
+                  , title = ""
+                  ) +
+  theme_sjplot(base_family = "serif" )
 ```
@@
   theme(legend.position = "none"
-        , text = element_text(size = 18)
+        , text = element_text(size = 18, family = "serif")
         )
@@
                   , digits = 4
-                  , title = "Head Down"
-                  , dv.labels = " Effects on foraging bout duration")
+                  , title = ""
+                  , dv.labels = "")
 
@@
                   , show.re.var = T
-                  , title = "Effects on foraging bout duration")
+                  , title = "") +
+  theme_sjplot(base_family = "serif")
 
@@
             , size = 6
-            , show.legend = F) +
+            , show.legend = F
+            , family = "serif"
+            ) +
   theme_classic() +
@@
         , legend.margin=margin()
-        , text = element_text(size = 18)
+        , text = element_text(size = 18, family = "serif")
         )
@@
                   , title = "Head Up"
-                  , dv.labels = " Effects on alert bout duration")
+                  , dv.labels = " Effects on alert bout duration")+
+  theme_sjplot(base_family = "serif")
 
@@
                   , show.re.var = T
-                  , title = "Effects on alert bout duration")
+                  , title = "Effects on alert bout duration") +
+  theme_sjplot(base_family = "serif")
 
@@
             , show.legend = F
+            , family = "serif"
             ) +
@@
         , legend.margin=margin()
-        , text = element_text(size = 18))
+        , text = element_text(size = 18, family = "serif"))
 
@@
                   , title = ""
-                  , dv.labels = " Effects on peck rate (per min)"
+                  , dv.labels = ""
                   , file = "Peck_Table.html")
@@
                   , show.re.var = T
-                  , title = "Effects on peck rate (per min)" )
+                  , title = "" ) +
+  theme_sjplot(base_family = "serif")
 ```
@@
   theme(legend.position = "bottom"
-        , text = element_text(size = 18))+
+        , text = element_text(size = 18, family = "serif"))+
   scale_color_manual(values = cbPalette
@@
 
-##SENT Barplot
+## SENT Barplot
 
@@
   geom_bar(stat='identity', width=.5, position = "dodge") +
-  geom_text(aes(label = N, group = SENTINEL_PRESENCE), position = position_dodge(width = 0.5), vjust = 1.5, size = 8) +
+  geom_text(aes(label = N, group = SENTINEL_PRESENCE)
+            , position = position_dodge(width = 0.5)
+            , vjust = 1.5
+            , size = 8
+            , family = "serif") +
   ylab("Number of observations") +
@@
   theme(axis.title.x = element_blank()
-        , text = element_text(size = 18)
+        , text = element_text(size = 18, family = "serif")
         , legend.position = "bottom"
@@
 
-##SENT Chi-Squared
+## SENT Chi-Squared
 
@@
         , axis.title.x = element_blank()
-        , text = element_text(size = 16))
+        , text = element_text(size = 18, family = "serif"))
 
@@
 # Pathway Analysis
-
-## Pathway
 
@@
   dplyr::select(VIDEO_ID,ID, DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_NB_TRANSITIONS, TOTAL_FREQUENCY_OF_DISTURBANCES | starts_with("NB_")) %>%
-  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES") %>%
+  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
+         , "Transitions" = "TOTAL_NB_TRANSITIONS") %>%
   mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
@@
         , legend.margin=margin()
-        , text = element_text(size = 12)
+        , text = element_text(size = 14, family = "serif")
         ) +
@@
 for (y in DV){
-  formula<-formula(paste(y,"~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+BAIT_PRESENCE+(1|TOTAL_NB_TRANSITIONS)"))
+  formula<-formula(paste(y,"~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+DISTURBANCE_FREQUENCY+BAIT_PRESENCE+(1|Transitions)"))
   PTWY.MODELS[[y]]<-glmer(formula, data = PTWY, family = "poisson", control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=100000)))
@@
                                   , "Peck to Alert")
-          , file = "PTWY_Table.html")
-
+          )+
+  theme_sjplot(base_family = "serif")
 ```
@@
             , hjust = -0.1
-            , show.legend = F) +
+            , show.legend = F
+            , family = "serif") +
   theme_classic() +
@@
         , legend.margin=margin()
-        , text = element_text(size = 12)
+        , text = element_text(size = 18, family = "serif")
         ) +
@@
   xlab("Disturbance Frequency (per min)") +
-  ylab("Number of Transitions from Foraging to Alert")
+  ylab("Transitions from Foraging to Alert")+
+  theme(text = element_text(size = 18, family = "serif"))
 

### 2024-01-25 Data & Scripts/Alex Popescu - Final Stats.Rmd
Jan 25
Go over the script with a fine tooth comb:
Make sure the right labels are right
Change "behavior" to "behaviour"
Continue working on the Results and methods section, address comments
WITH THE NEW CHANGES, outline the discussion again

@@
   '(Intercept)' = "Intercept"
-  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
+  , 'SENTINEL_PRESENCESentinel Present' = "Sentinel Presence"
   , BEHAVIORHU = "Behavior"
@@
   , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
-  , 'BEHAVIORHU:SENTINEL_PRESENCESentinel Absent' = "Behavior: Sentinel Presence"
-  , 'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
+  , 'BEHAVIORHU:SENTINEL_PRESENCESentinel Present' = "Behavior: Sentinel Presence"
+  , 'SENTINEL_PRESENCESentinel Present:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
 )
@@
   '(Intercept)' = "Intercept"
-  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
+  , 'SENTINEL_PRESENCESentinel Present' = "Sentinel Presence"
   , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
@@
   , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
-  ,'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
+  ,'SENTINEL_PRESENCESentinel Present:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
   )
@@
 labs.ptwy<-c('(Intercept)' = "Intercept"
-  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
+  , 'SENTINEL_PRESENCESentinel Present' = "Sentinel Presence"
   , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
@@
   , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
-  , 'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
+  , 'SENTINEL_PRESENCESentinel Present:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
   )
@@
   mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
+                                           , 'NO' = "Sentinel Absent"
                          , 'YES' = "Sentinel Present"
-                         , 'NO' = "Sentinel Absent"))
+                         ))
 BOUT$LDURATION<-log(BOUT$DURATION) #Transform data for normality
@@
   scale_colour_manual(values = cbPalette
-                      , labels = c("Sentinel Absent", "Sentinel Present")
                       , name = "") +
@@
   mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
-                         , 'YES' = "Sentinel Present"
-                         , 'NO' = "Sentinel Absent"))
+                                           , 'NO' = "Sentinel Absent"
+                                           , 'YES' = "Sentinel Present"))
 
@@
   ylab("Number of Transitions") +
+  ylim(0,100)+
   scale_colour_manual(values = cbPalette
-                      , labels = c("Sentinel Absent"
-                                   , "Sentinel Present")
                       , name = "") +

### 2024-01-25 Data & Scripts/PECK.Rmd
Jan 25
Go over the script with a fine tooth comb:
Make sure the right labels are right
Change "behavior" to "behaviour"
Continue working on the Results and methods section, address comments
WITH THE NEW CHANGES, outline the discussion again

@@
 ```{r Peck Summary, echo = F}
-source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
+### Summarize
 PECK.MEAN<-summarySE(DATA, "PECK_RATE", c("GENERALIZED_ENVIRONMENT", "SENTINEL_PRESENCE"), na.rm=T)
@@
 
-VERY WEIRD. Model 1 without the interactions is preferred over Model 2. However, the simple MOD0 is preferred. Let's see the outputs of that model.
+VERY WEIRD. Model 1 without the interactions is perferred over Model 2. However, the simple MOD0 is preferred. Let's see the outputs of that model.
 
-```{r Robust PECK MODEL 0}
+```{r Robust PECK MODEL 2}
 sjPlot::tab_model(PECK.MOD0
@@
 
-```{r PECK Disturb 1}
+```{r PECK Disturb}
 PK.DF.DOT1<-DATA %>%
@@
 
-```{r PECK Disturb 2}
+```{r PECK Disturb}
 PK.DF.DOT2<-DATA[which(DATA$BEHAVIOR == "HD" & DATA$BEHAVIOR_DURATION > 0),] %>%
@@
 
-Very nice!
-While it is not the prettiest of results, it roughly appears that as the disturbance frequency increases, so does the peck rate.
-I believe this can be explained by a sense of 'urgency' in higher-risk environments.
-Greater sample sizes would be required to make conclusive statements.
+Very nice! We can see that as the disturbance frequency increases, so does the peck rate. I believe this can be explained by a sense of 'urgency' in higher-risk environments.
 

### 2024-01-30 Data & Scripts/Alex Popescu - Final Stats.Rmd
Jan 30th
Script reviewed: all good now
ES Methods and Results NEEDS to be completed tonight

@@
   , SENTINEL_PRESENCEYES = "Sentinel Presence"
-  , BEHAVIORHU = "Behavior"
+  , BEHAVIORHU = "Behaviour"
   , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
@@
   , 'SENTINEL_PRESENCESentinel Present' = "Sentinel Presence"
-  , BEHAVIORHU = "Behavior"
+  , BEHAVIORHU = "Behaviour"
   , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
@@
   , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
-  , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
-  , 'BEHAVIORHU:SENTINEL_PRESENCESentinel Present' = "Behavior: Sentinel Presence"
+  , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behaviour: Generalized Environment"
+  , 'BEHAVIORHU:SENTINEL_PRESENCESentinel Present' = "Behaviour: Sentinel Presence"
   , 'SENTINEL_PRESENCESentinel Present:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
@@
                 , position = position_dodge(width=0.1))+
-  geom_text(aes(label = round(DURATION, 2))
+  geom_text(aes(label =  scales::label_number(accuracy=0.01)(DURATION))
             , hjust = -0.2

### 2024-01-31 Data & Scripts/Alex Popescu - Final Stats.Rmd
Jan 31
Gotta go to work

@@
 
-##BOUT Model - Head Down
+## BOUT Model - Head Down
 
@@
                   , title = "Head Up"
-                  , dv.labels = " Effects on alert bout duration")+
-  theme_sjplot(base_family = "serif")
+                  , dv.labels = " Effects on alert bout duration")
 
@@
                                   , "Peck to Alert")
-          )+
-  theme_sjplot(base_family = "serif")
+          )
 ```

### 2024-02-01 Data & Scripts/Alex Popescu - Final Stats.Rmd
Feb 1
Eyyyyy Methods and results sent in again. Work on the discussion outline.

@@
                                   , "Peck to Alert")
+          , file = "PTWY_Table.html"
           )

### 2024-09-25 Data & Scripts/Alex Popescu - Final Stats.Rmd
Habituation test completed

@@
 # Bookkeeping
-
 ```{r Packages, include = F}
@@
 ```
+
+## Habituation
+```{r plot habituation}
+Habit<-BOUT.raw %>%
+   filter(., BEHAVIOR == "HU", LATITUDE == "43.18262") %>%
+           ggplot(.
+               , aes(x = JULIAN_DATE
+                     , y = DURATION
+                     )) +
+  geom_point() +
+  geom_smooth(method="lm") +
+  theme_classic() +
+  xlab("Sampling Date") +
+  ylab("Duration of bouts of alertness (s)") +
+  theme(text = element_text(size = 18, family = "serif"))
+Habit
+
+```
+
+```{r Habituation test - opportunistic sampling}
+Habit.MOD.opp<-lm(DURATION~JULIAN_DATE, data = subset(BOUT.raw, BOUT.raw$BEHAVIOR == "HU" & BOUT.raw$LATITUDE != "43.18262" & BOUT.raw$BAIT_PRESENCE == "YES"))
+
+sjPlot::tab_model(Habit.MOD.opp
+                  , show.stat = T
+                  , show.se = T
+                  , digits = 4
+                  , show.re.var = T
+                  , title = ""
+                  , dv.labels = ""
+)
+```
+
+```{r Habituation test - recurrent sampling}
+Habit.MOD.rec<-lm(DURATION~JULIAN_DATE, data = subset(BOUT.raw, BOUT.raw$BEHAVIOR == "HU" & BOUT.raw$LATITUDE == "43.18262" & BOUT.raw$BAIT_PRESENCE == "YES"))
+
+sjPlot::tab_model(Habit.MOD.rec
+                  , show.stat = T
+                  , show.se = T
+                  , digits = 4
+                  , show.re.var = T
+                  , title = ""
+                  , dv.labels = ""
+)
+```
